### 2016-08-21 camera_ready/paper.tex
first commit

\documentclass[runningheads,a4paper]{llncs}
%\usepackage[top=1.3in, bottom=1.3in, left=1.3in, right=1.3in]{geometry}
\usepackage{floatrow}
\usepackage{graphicx} % more modern
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{subfigure} 
%\usepackage{natbib}
\usepackage[footnotesize]{caption}
\usepackage{multirow}
\usepackage{makeidx}

%%%%%%%%%%%%%%%%%%%%%%%%%%
\newfloatcommand{capbtabbox}{table}[][11cm]

%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{url}
\urldef{\mailsa}\path|mkshirs@us.ibm.com|
\urldef{\mailsb}\path|{jgc,judithks,kmuruges}@cs.cmu.edu|
%%%%%%%%%%%%%%%%%%%%%%%%%%

\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\argmin}{\operatornamewithlimits{argmin}}

%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{tocdepth}{3}

%\setcounter{secnumdepth}{0}

\begin{document}
\pagestyle{headings}

\mainmatter

\title{Multitask matrix completion for learning protein interactions across diseases}

\titlerunning{Multitask matrix completion}

\author{Meghana Kshirsagar$^{1}$%
\and Jaime G. Carbonell$^{2}$
\and Judith Klein-Seetharaman$^{3}$
\and Keerthiram Murugesan$^{2}$}

\authorrunning{Multitask matrix completion}
% (feature abused for this document to repeat the title also on left hand pages)

\institute{$^{1}$IBM T. J. Watson Research, Yorktown Heights, NY 10598, USA \\
$^{2}$Language Technologies Institute, Carnegie Mellon University, 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
$^{3}$Metabolic \& Vascular Health, Warwick Medical School, University of Warwick, Coventry, UK\\
\mailsa \quad \mailsb
}

\toctitle{Matrix completion for transfer learning in host-pathogen PPI networks}
\tocauthor{Meghana Kshirsagar, Jaime G. Carbonell, Judith Klein-Seetharaman and Keerthiram Murugesan}
\maketitle

\begin{abstract}

Disease causing pathogens such as viruses, introduce their proteins into the host cells 
where they interact with the host's proteins enabling the virus to replicate inside the host.
These interactions between pathogen and host proteins are key to understanding infectious diseases.
Often multiple diseases involve phylogenetically related or biologically similar pathogens. Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}. 
Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
We obtain upto 30 \% improvement in predictive performance over prior state-of-the-art models. 
We show how our model's parameters can be interpreted to reveal both general and specific
interaction-relevant characteristics of the viruses. Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}
\keywords{host-pathogen protein interactions, multitask learning, matrix completion}
\end{abstract}

\section{Introduction}

Infectious diseases such as H1N1 influenza, the recent Ebola outbreak and bacterial infections, such as the recurrent \textit{Salmonella} and \textit{E. coli} outbreaks
are a major health concern worldwide, causing millions of illnesses and many deaths each year. 
Key to the infection process are host-pathogen interactions at the molecular level, where pathogen proteins physically bind to human proteins
to manipulate important biological processes in the host cell, to evade the host's immune response and to multiply within the host.
Very little is known about these protein-protein interactions (PPIs) between pathogen and host proteins for any individual disease.
However, such PPI data is widely available across several diseases, and the central question in this
paper is: \emph{Can we model host-pathogen PPIs better by leveraging
data across multiple diseases?} This is of particular interest for 
lesser known or recently evolved diseases where the data is particularly scarce. Furthermore, it
allows us to learn models that generalize better across diseases by
modeling global phenomena related to infection.

An elegant way to formulate the interaction prediction problem is via a graph completion based framework, where we have several bipartite 
graphs over multiple hosts and pathogens.% as illustrated in Figure \ref{fig:hpppigraph}.
Nodes in the graphs represent host and pathogen proteins, with edges between them representing interactions 
(host protein $\frac{interacts}{}$ pathogen protein).
Given some observed edges (interactions obtained from laboratory based experiments), we wish to predict the other edges in the graphs.
Such bipartite graphs arise in a plethora of problems including: recommendation systems (user $\frac{prefers}{}$ movie), 
citation networks (author $\frac{cites}{}$ paper), disease-gene networks (gene $\frac{influences}{}$ disease) etc. 
In our problem, each bipartite graph $\mathcal{G}$ can be represented using a matrix $M$, where the rows correspond to pathogen proteins 
and columns correspond to host proteins.
The matrix entry $M_{ij}$ encodes the edge between pathogen protein $i$ and host protein $j$ from the graph, with $M_{ij}=1$ for the observed interactions. 
Thus, the graph completion problem can be mathematically modeled as a matrix completion problem \cite{candes08}. 
%Traditional approaches towards matrix completion rely on the assumption that the underlying function that generated the matrix can be decomposed into a small number of `latent' factors. The solution based on this assumption involves finding a low-rank matrix factorization for $M$, mathematically expressed as: finding $U$ and $V$ such that $M \approx UV^T$. Here the parameters $U$ and $V$ are called the factor matrices and represent the latent properties of the host and pathogen proteins respectively. In the recommendation systems example, the `low rank' structure suggests that movies can be grouped into a small number of latent `genres'. In the case of proteins these latent properties could correspond to various \textit{biological functions} of proteins or encode \textit{interaction propensities}.

Most of the prior work on host-pathogen PPI prediction has modeled each bipartite graph separately, and hence cannot exploit the similarities 
in the edges across the various graphs. Here we present a \textit{multitask} matrix completion method that \textit{jointly models} 
several bipartite graphs by sharing information across them. 
From the multitask perspective, a \textit{task} is the graph between one host and one pathogen (can also be seen as interactions relevant to one disease). 
We focus on the setting where we have a single host species (human) and several related viruses, where we hope to gain from the 
fact that similar viruses will have similar strategies to infect and hijack biological processes in the human body.
%Such opportunities for sharing arise in other applications as well: for instance, predicting user preferences in movies may inform preferences in selection of books, or vice-versa, as movies and books are semantically related. Multitask learning based models that incorporate and exploit these correlations should benefit from the additional information.
% and result in a better prediction performance. 
Our model is motivated by the following biological intuition governing protein interactions across diseases.
\begin{enumerate}
\item An interaction depends on the structural properties of the proteins, which are conserved across similar viruses as they have evolved from common ancestors. We use a component to capture these latent similarities, that is \textit{shared} across tasks.
%This latent structural similarity can be expressed via latent factors that project the proteins into a shared lower dimensional subspace that is common to all tasks.
\item In addition to the shared properties discussed above, each pathogen has also evolved specialized mechanisms to target host proteins. These are unique to the pathogen and can be expressed using a \textit{task-specific} parameter.
\end{enumerate}

This leads us to the following model that incorporates the above ideas. The interactions matrix $M_t$ of task $t$ can be written as: $M_t = \mu_t *\; \textrm{(shared component)} + (1-\mu_t) *\; \textrm{(specific component)} $, with hyperparameter $\mu_t$ allowing each task to customize its' amount of shared and specific components.

\begin{table}
\caption{Tasks and their sizes. Each column corresponds to one bipartite graph between human proteins and the pathogen indicated in the column header. All pathogens are single stranded RNA viruses. The interactions and the protein count both includes data across various strains of each pathogen}
% (see Sec \ref{sec:datasets} for details). The last row shows that each of our graphs is extremely sparse.} 
\label{datasets}
\begin{footnotesize}
\begin{tabular}{crrr} \toprule
Task $\rightarrow$ & \textbf{\textit{Influenza A}} & \textbf{\textit{Hepatitis C}} & \textbf{\textit{Ebola}}\\ \midrule 
number of HP PPIs (positives) & 848& 981 & 90 \\\hline
\# of distinct virus proteins in PPIs & 54 & 151 & 2 \\\hline
{\# of distinct human proteins in PPIs} & 362 & 385 & 88 \\\hline
{total \# of virus proteins across strains} & 542 & 163 & 150 \\\hline
{number of negatives} & 84800 & 98100 & 9000 \\\hline
{density of observed graph$^\ddagger$ (as \%)} & .15 & .60 & .06 \\\hline
\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
\multicolumn{4}{l}{\scriptsize{$\ddagger$: considering all proteins from the two tasks involved. }}\\
\multicolumn{4}{l}{\scriptsize{Note: considering the total number of human proteins to be $\approx$ 100,000.}}\\
\bottomrule
\end{tabular}
\end{footnotesize}
\end{table}

To incorporate the above ideas, we assume that the interactions matrix $M$ is generated from two components. 
The first component has low-rank latent factors over the human and virus proteins,
with these latent factors jointly learned over all tasks. The second component involves a task specific parameter, on which we
additionally impose a sparsity constraint as we do not want this parameter to overfit the data. Section \ref{ourmodel} discusses
our model in detail. We trade-off the relative importance of the two components using task-specific hyperparameters.
We can thus learn what is conserved and what is different across pathogens, rather than having to specify it manually.

%One of the key challenges in inducing such a model is: in addition to the graph structure itself, we want to incorporate information about the nodes available in the form of features. 
The applications that we consider involve extremely sparse graphs with a large number of nodes and very few observed edges. There will be nodes i.e proteins that are not involved in any known interactions -- the model should be able to predict links between such prior `unseen' node pairs (called the \textit{cold start problem} in the recommendation systems community). 
For instance, the host-pathogen PPI network of human-Ebola virus (column-3, Table \ref{datasets}) has $\approx$ 90 observed edges (equivalent to 0.06\% of the possible edges) which involve only 2 distinct virus proteins. Any biologist studying virus-human interactions will be more interested in the virus proteins which have yet unknown interactions.
The main contributions of this work are:\\
1. We extend a prior matrix completion model \cite{abernethy} to the multitask setting.\\
2. We leverage node features which allows us to predict on unseen nodes.\\
3. We apply the model to an important problem -- prediction of interactions in disease-relevant host-pathogen protein networks, for multiple related diseases and demonstrate improved performance over prior state-of-the-art multitask models.\\
4. We use unlabeled data to initialize the parameters of our model, which gives us a modest boost in prediction performance.

\subsection{Background: Host-pathogen protein interactions}
%An important part of our sustenance depends on how we counter infectious diseases -- those that are caused by external agents like pathogens. To understand how these microbial organisms reside and disrupt the host processes we need to comprehend the cross-organism protein interactions between the host species such as mammals, and the causative pathogen species such as viruses and bacteria. %These interactions  pathogen proteins physically bind with human proteins to manipulate important biological processes in the host cell, evade the host immune response and multiply within the host. 
The experimental discovery of host-pathogen protein interactions involves biochemical and biophysical methods such as co-immunoprecipitation (co-IP), yeast two-hybrid (Y2H) assays, co-crystalization. % These databases are valuable sources of information to bioinformaticians for developing models. \\
The most reliable experimental methods are often very time-consuming and expensive, making it hard to investigate the prohibitively large set of possible host-pathogen interactions -- e.g. the bacterium \textit{Bacillus anthracis} with about 2321 proteins when coupled with the 100,000 human proteins gives $\approx$232 million protein pairs to validate. Computational techniques complement laboratory-based methods by predicting highly probable PPIs. Supervised machine learning based methods use the few known interactions as training data and formulate the interaction prediction problem in a classification setting%, with target classes: ``interacting'' or ``non-interacting''. Features are derived using various attributes of the two proteins. %such as: protein sequences from Uniprot \cite{uniprot}, protein structure from PDB, gene ontology from GO database \cite{GO}. 
%In our work we use protein-sequence based features.

\subsection{Prior work}
%\label{prior_work}
Most of the prior work in PPI prediction has focussed on building models separately for individual 
organisms \cite{struct2net,qi06} or on building a model specific to a disease in the case of host-pathogen PPI 
prediction \cite{tastan09,dyer07,me12}. There has been limited work on combining PPI datasets to learn joint models. \cite{qi2010} proposed a semi-supervised multi-task framework to predict PPIs from partially labeled reference sets. \cite{me_ismb_2013} develop a task regularization based
framework that incorporates the similarity in biological pathways targeted by various diseases.
%  to couple multiple tasks together. %Matrix factorization based protein-protein interaction (PPI) prediction has seen very little work, mainly due to the extremely sparse nature of these datasets which makes it very difficult to get reliable predictors. 
\cite{ppi_cmf} uses a collective matrix factorization based approach in a multi-task learning setting for within species PPI prediction. The methods used 
in all prior work on PPI prediction do not explicitly model the features of the proteins and cannot be applied on proteins which have no known interactions available. Our work addresses both these issues.
%in using a formulation of the matrix completion problem originally proposed in the work by \cite{abernethy}.
%A majority of the prior work in the relevant areas of collaborative filtering and link prediction includes single relation models that use neighbourhood based prediction \cite{sarwar01}, matrix factorization based approaches \cite{koren09,menon_ecml11} and bayesian approaches using graphical models \cite{jin02,phung09}. 
%The matrix factorization based methods have been more popular than the other methods. 
%There have also been multitask approaches on link prediction \cite{zhang12,cao_icml2010,li09,cmf}. 
%\cite{li11} presents a survey on multitask/transfer methods in collaborative filtering. 
%The multi-relational learning literature \cite{xu09} and the work on link prediction in heterogenous networks is not relevant as their setting involves \textit{different types of relationships} between the \textit{same} set of nodes.
%The prior work on incorporating side information in the form of features of the nodes (users/movies) and edges
%has been proposed for single-graph link prediction problems. 
%\cite{menon_ecml11} propose a single-graph model that combines linear and bilinear features, latent parameters on the nodes and several other parameters into a function that minimizes a ranking loss. 
%Probabilistic models to combine network structure with node features have also been proposed. 
We extend the framework from \cite{abernethy} to the multitask setting. 
%There has been a lot of work on other low-rank models for multitask learning \cite{ando2005,ji2009,chen2012,chen2013}. 
% that try to capture the task relationship via a shared low-rank structure over the model parameters. Chen et al \cite{chen2012} have developed efficient algorithms for sparse and low-rank multitask learning models for classification.

%\subsection{Background: Host-pathogen protein interactions}
%The biochemical and biophysical methods that study protein-protein interactions (PPI) are very time-consuming and expensive, making it hard to investigate the 
%prohibitively large set of possible host-pathogen interactions\footnote{The bacterium \textit{Bacillus anthracis} which causes anthrax has $\approx$ 2321 proteins, the human body has $\approx$ 100,000 giving $\approx$232 million protein pairs to test}. 
%Computational techniques complement these laboratory-based methods by predicting probable PPIs allowing biologists to focus their resources and time on the more promising candidates. Supervised learning methods use the few experimentally-discovered interactions as training data and solve the problem in a classification setting, with classes: ``interacting'' or ``non-interacting''. These use features derived using protein properties such as: protein sequences, protein structure, gene function etc.

%\noindent{\textbf{Multi-task learning for host-pathogen PPI prediction}}: Pathogens that have similar DNA sequences produce proteins with similar functions and will have similar strategies of infecting their hosts. Knowledge can thus be shared across diseases to better understand various global phenomena related to infection and to improve predictive models. Multitask approaches thus have several advantages, especially in dealing with data scarcity issues that hamper the construction of disease-specific models for less studied diseases. In our setting, each disease can be considered as a `task' i.e the PPIs involving the host and the causative pathogen. Figure \ref{fig:hpppigraph} shows the various possible multitask settings for this problem.

\section{Bilinear low-rank matrix decomposition}
In this section, we present the matrix decomposition model that we extend for the multitask scenario. In the context of our problem, at a high level, this model states that -- protein interactions can be 
expressed as dot products of features in a lower dimensional subspace.

\label{MCintro}
Let $\mathcal{G}_{t}$ be a bipartite graph connecting nodes of type $\upsilon$ with nodes of type $\varsigma$. Let there be $m_t$ nodes of type $\upsilon$ and $n_t$ nodes of type $\varsigma$. We denote by $M \in \mathbb{R}^{m_t \times n_t}$, the matrix representing the edges in $\mathcal{G}_{t}$. Let the set of observed edges be $\Omega$. 
Let $\mathcal{X}$ and $\mathcal{Y}$ be the feature spaces for the node types $\upsilon$ and $\varsigma$ respectively. For the sake of notational convenience we assume that the two feature spaces have the same dimension $d_t$ \footnote{the dimensions being different does not influence the method or the optimization in any way}. Let $\mathbf{x}_i \in \mathcal{X}$ denote the feature vector for a node $i$ of type $\upsilon$ and $\mathbf{y}_j \in \mathcal{Y}$ be the feature vector for node $j$ of type $\varsigma$. The goal of the general matrix completion problem is to learn a function $f : \mathcal{X} \times \mathcal{Y} \rightarrow \mathbb{R}$ that also explains the observed entries in the matrix $M$. We assume that the function $f$ is bilinear on $\mathcal{X} \times \mathcal{Y}$. This bilinear form was first introduced by \cite{abernethy} and takes the following form:
\begin{equation}
f(\mathbf{x}_i, \mathbf{y}_j) = \mathbf{x}_i^\intercal H \mathbf{y}_j = \mathbf{x}_i^\intercal U V^\intercal \mathbf{y}_j
\end{equation}

The factor $H \in \mathbb{R}^{d_t \times d_t}$ maps the two feature spaces $\mathcal{X}$ and $\mathcal{Y}$. This model assumes that $H$ has a low-rank factorization given by $H = U V^\intercal$, where $U \in \mathbb{R}^{d_t \times k}$ and $V \in \mathbb{R}^{d_t \times k}$. The factors $U$ and $V$ project the two feature spaces to a common lower-dimensional subspace of dimension $k$. While the dimensionality of the feature spaces $\mathcal{X}$ and $\mathcal{Y}$ may be very large, the latent lower dimensional subspace is sufficient to capture all the information pertinent to interactions. To predict whether two new nodes (i.e nodes with no observed edges) with features $\mathbf{p}_i$ and $\mathbf{q}_j$ interact, we simply need to compute the product: $\mathbf{p}_i U V^\intercal \mathbf{q}_j$. This enables the model to avoid the cold start problem that many prior models suffer from. The objective function to learn the parameters of this model has two main terms: (1) a data-fitting term, which imposes a penalty for deviating from the observed entries in $\Omega$ and (2) a low-rank enforcing term on the matrix $H$.

The first term can be any loss function such as squared error, logistic-loss, hinge loss. We tried both squared error and logistic-loss and found the performance to be similar. The squared error function has the advantage of being amenable to adaptive step-size based optimization which results in a much faster convergence. The low-rank constraint on $H$ (mentioned in (2) above) is NP-hard to solve and it is standard practice to replace it with either the trace-norm or the nuclear norm. Minimizing the trace norm (i.e. sum of singular values) of $H = UV^\intercal$, is equivalent to minimizing $\|U\|^2_F + \|V\|^2_F$. This choice makes the overall function easier to optimize:
\begin{equation}
\label{objective}
\begin{array}{ll}
\mathcal{L}(U,V) & =\displaystyle{\sum_{(i,j) \in \Omega}} c_{ij} \; \ell(M_{ij}, \mathbf{x}_i^\intercal U V^\intercal \mathbf{y}_j) + \lambda ( \|U\|^2_F + \|V\|^2_F ) \\
& \textrm{ where } \ell(a, b) = (a - b)^2
\end{array}
\end{equation}

The constant $c_{ij}$ is the weight/cost associated with the edge $(i,j)$ which allows us to penalize the error on individual instances independently. The parameter $\lambda$ controls the trade-off between the loss term and the regularizer. 
%The function in equation(\ref{objective}) is non-convex. To optimize this function a common procedure called alternating minimization (AltMin) is used, which fixes one of the two parameters and optimizes w.r.t the other.

%\noindent\textbf{Biological intuition}: The projection allows each feature to take on several different weights and hence allows for more flexibility as compared to a linear model. In the context of the host-pathogen PPI prediction, one can interpret this new subspace to represent latent biological characteristics of the features or binding propensities. In the Results section, we will analyze these parameters $U$ and $V$ and show how they can be used to understand the PPIs data better.

\section{The bilinear sparse low-rank multitask model (BSL-MTL)} 
\label{ourmodel}
In the previous section, we described the bilinear low-rank model for matrix completion. 
Note that in order to capture linear functions over the features, we introduce a constant feature for every protein 
(i.e $[\mathbf{x}_i 1]$). We now discuss the multitask extensions that we propose. Let $\{\mathcal{G}_t\}$ where $t=1 \ldots T$ be the set of $T$ bipartite graphs and the corresponding matrices be $\{M_t\}$. Each matrix $M_t$ has rows corresponding to node type $\upsilon_t$ and columns corresponding to the node type $\varsigma_t$. The feature vectors for individual nodes of the two types be represented by $\mathbf{x}_{ti}$ and $\mathbf{y}_{tj}$ respectively. Let $\Omega_t$ be the set of observed links (and non-links) in the graph $\mathcal{G}_t$. Our goal is to learn individual link prediction functions $f_t$ for each graph. In order to exploit the relatedness of the $T$ bipartite graphs, we make some assumptions on how they share information. We assume that each matrix $M_t$ has a low-rank decomposition that is shared across all graphs and a sparse component that is specific to the task $t$. That is,
\begin{equation}
f_t(\mathbf{x}_{ti}, \mathbf{y}_{tj}) = \mathbf{x}_{ti}^\intercal H_t \mathbf{y}_{tj}, \textrm{ where } H_t = \mu_t U V^\intercal + (1-\mu_t)  S_t
\end{equation}

As before, the shared factors $U$ and $V$ are both $\mathbb{R}^{d_t \times k}$ (where the common dimensionality $d_t$ of the two node types is assumed for convenience). The matrix $S_t \in \mathbb{R}^{d_t \times d_t}$ is a sparse matrix. The objective function for the multitask model is given by:
\begin{equation}
\label{mtl_objective}
\begin{array}{ll}
\mathcal{L}(U,V,\{S_t\}) & = \displaystyle{ \frac{1}{N} \sum_{t=1}^T \sum_{(i,j) \in \Omega_t}} c^t_{ij} \; \ell \big( M_{t_{ij}}, \mathbf{x}_{ti}^\intercal H_t \mathbf{y}_{tj} \big)^2 + \lambda ( \|U\|^2_F + \|V\|^2_F ) + \displaystyle{\sum_{t=1}^T \sigma_t \|S_t\|_1} 
\end{array}
\end{equation}

Here $N = \sum_t | \Omega_t |$, is the total number of training examples (links and non-links included) from all tasks. To enforce the sparsity of $S_t$ we apply an $\ell_1$ norm. In our experiments, we tried both $\ell_1$ and $\ell_2$ norms and found that the $\ell_1$ norm works better.

%\input{altmin_algo.tex}

%\begin{algorithm}[h]   \caption{Alternating Least Squares}
%   \label{alg}
%   \begin{small}
%\begin{algorithmic}[1]
%\STATE \textbf{Input}: \\
%$k$ : number of latent factors\\
%$\Gamma$: pairs of entities for initialization\\
%For every task $t$,\\
%${\{\mathbf{x}_{ti}\}, \{\mathbf{y}_{tj}\}}$: feature vectors for pathogen and host proteins resp.\\
%$\Omega_t$: the observed entries of the matrix $M_t$\\
%\STATE \textbf{Initialization}: \\
%\STATE An iteration $r$, let $\mathbf{S}^r$ represent $\{S^r_t\}_{t=1}^T$
%\STATE $\mathbf{S}^0 \leftarrow 0$
%\STATE $U^0 \leftarrow$ top-$k$ left singular vectors and $V^0 \leftarrow$ top-$k$ right singular vectors from the SVD of $\displaystyle{\sum_{(i,j) \in \Gamma}} \mathbf{x}_i \mathbf{y}_j^\intercal$\\
%\STATE $\mathcal{L}^0$ : initial loss
%\REPEAT
%  \STATE $U^{r+1} \leftarrow \displaystyle{\argmin_{U}} \; \mathcal{L}(U, V^r, \mathbf{S}^r)$  \STATE $V^{r+1} \leftarrow \displaystyle{\argmin_{V}} \; \mathcal{L}(U^{r+1}, V, \mathbf{S}^r) $  \STATE For each task $t$\\
%  \hspace{0.5cm} $S^{r+1}_t \leftarrow \displaystyle{\argmin_{S_t}} \; \mathcal{L}(U^{r+1}, V^{r+1}, \mathbf{S}^r_{-t}) $
%  \STATE Compute $\mathcal{L}^{r+1}$ and let $\delta \leftarrow (\mathcal{L}^r - \mathcal{L}^{r+1})/\mathcal{L}^r$
%\UNTIL {$\delta < \tau$}%\While
%\end{algorithmic}\end{small}\end{algorithm}

\noindent \textbf{Optimization}:
The function $\mathcal{L}(U, V, \{S_t\})$ is non-convex. However, it is convex in every one of the parameters (i.e when the other parameters are fixed) 
and a block coordinate descent method called alternating least squares (ALS) is commonly used to optimize such functions. 
To speed up convergence we use an adaptive step size. %The detailed optimization procedure is shown in Algorithm \ref{alg}.

\noindent \textbf{Convergence:} The ALS algorithm is guaranteed to converge only to a local minimum. There is work showing convergence guarantees to global optima for related simpler problems, however the assumptions on the matrix and the parameter structure are not very practical and 
it is difficult to verify whether they hold for our setting. \\
\noindent \textbf{Initialization of $U$ and $V$}: We tried random initialization (where we randomly set the values to lie in the range [0 1]), and also 
the following strategies that initialize: $U^0 \leftarrow$ top-$k$ left singular vectors, and $V^0 \leftarrow$ top-$k$ right singular vectors from 
the SVD of $\displaystyle{\sum_{(i,j) \in \Gamma}} m_{ij} \mathbf{x}_i \mathbf{y}_j^\intercal$. We set $\Gamma$ to 
(a) training examples from all tasks, or (b) a random sample of 10000 unlabeled data from all tasks. We found that using the unlabeled data for initialization gives us a better performance.

%(a) \textit{Training data}: $\Gamma$ is set to all the training examples from all tasks. This initialization from Jain et al. \cite{prateek} uses the labels of the examples.\\
%(b) \textit{Unlabeled data}: $\Gamma$ is set to . Note that there are no labels $m_{ij}$ available here. In our experiments we found that using the unlabeled data for initialization gives us a better performance than using the training data or random initialization.

\subsection{Handling the `curse of missing negatives'}
\label{negs}
For the MC algorithm to work in practice the matrix entries $M_{ij}$ should represent interaction scores (range [0 1]) 
or take binary values (1s for positives and 0s for negatives). Our experiments with PPI probabilities (obtained using the MINT-scoring algorithm)
gave bad models. The binary matrix setting requires some observed 0s. However non-interactions are not available as they cannot be verified 
experimentally for various reasons. Hence we derived a set of `probable negatives' using a heuristic often used in PPI prediction 
work \cite{qi06,qi09,dyer11,me_ismb_2013}. We pair up all virus proteins with all human proteins and sample a random set to be negatives. 
This heuristic works in practice as the interaction ratio (i.e number of positives in a large random set of protein pairs) is expected to be very low:
$\approx$ 1/100 to 1/500. That is, the probability that our negatives contain true positives is negligible. \\
\textbf{High class imbalance}: We incorporate the prior on the interaction ratio by setting the size of our randomly sampled negatives set equal to
100 times the number of gold standard positives.

\section{Experimental setup}
\subsection{Datasets and features}
\label{sec:datasets}
We use three human-virus PPI datasets from the PHISTO \cite{phisto} database (version from 2014), the characteristics of which are summarized in Table \ref{datasets}. 
The \textit{Influenza A} task includes various strains of flu: H1N1, H3N2, H5N1, H7N3. 
Similarly, the \textit{Hepatitis} task includes various strains of the virus. \footnote{Since we use data from several strains for each task, the PPI data contains some interactions that are interologs. Please see the supplementary for details} 
All three are single-strand RNA viruses, with \textit{Hepatitis} being a positive-strand ssRNA whereas \textit{Influenza} and \textit{Ebola} are negative-strand viruses. %Note that we use UniprotKB identifiers to refer to the proteins; hence if an identical protein appears in two strains it will be treated differently as it has a different UniprotKB id. 
The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (last row in Table \ref{datasets}).\\
\noindent\textbf{Features}: 
Since the sequence of a protein determines its structure and consequently its function, it may be possible to predict PPIs using the amino acid sequence of a protein pair. \cite{shen:2007} introduced the ``conjoint triad model'' for predicting PPIs using only amino acid sequences. They partitioned the twenty amino acids into seven classes based on their electrostatic and water affinities.\footnote{For details of these classes, please refer to the supplementary or the original paper}
A protein's amino acid sequence is first transformed to a class-sequence (by replacing each amino acid by its class).
For $k$=3, they count the number of times each distinct tri-mer (set of three consecutive amino acids) occurred in the sequence.
Since there are 343 ($7^3$) possible tri-mers (with an alphabet of size 7), the feature vector containing the tri-mer frequency counts will have 343 elements.
To account for protein size, they normalized the counts by linearly transforming them to lie between 0 and 1.
Thus the value of each feature in the feature vector is the normalized count for each of the possible amino acid three-mers. We use di-, tri- and four-mers thus leading to a total of 2793 features ($7^2 + 7^3 + 7^4$).
Such features have been successfully applied in prior work \cite{dyer07,me_ismb_2013}.

\subsection{Competing methods}
The results upon comparing our model to various single and multitask methods are shown in Table \ref{resultsTable}: including other recent low-rank and sparse models, conventional multitask methods and prior work on HP PPI prediction. Wherever appropriate, we concatenated the features of the two node types into a single feature vector. Let $W \in \mathbb{R}^{T \times d_t}$ be the matrix with the task-specific weight vectors $\mathbf{w}_t$. For a uniform comparison we used least squared loss in all the methods. The MALSAR package was used for some of the baselines.

\noindent\textbf{Single task (STL)}: Ridge regression with $\ell_2$ regularization (which performed better than $\ell_1$). \\
\noindent\textbf{MMTL}: The mean regularized multitask learning model from \cite{pontil04}. \\
%\noindent\textbf{Dirty model} \cite{jalali2010}: This model assumes that $W = P + Q$, where $P$ enforces group sparsity and $Q$ is to control for element-wise sparsity. It uses the regularizer $\rho_1 \|P\|_{1,\infty} + \rho_2 \| Q \|_1$. \\
\noindent\textbf{Low rank model (TraceNorm)}: A low-rank structure is enforced on $W$ by minimizing the nuclear norm $\|W\|_*$. \\
\noindent\textbf{Sparse + low-rank (SpLowRank)} \cite{chen2012}: $W$ is assumed to have the decomposition: $W = P + Q$, where $P$ is sparse and $Q$ has a low-rank structure. \\
\noindent\textbf{IMC} \cite{prateek,nagarajan}: This is the link-prediction model from Section \ref{MCintro}, where data from all tasks is combined without incorporating any task relationships (comparable to the `union' setting from \cite{widmer}). $U$ and $V$ are shared by all tasks. We use the same initialization for this method as we do for our model. A comparison to this model tells us how much we gain from the task-specific sparsity component $S_t$. \\
\noindent\textbf{MTPL} \cite{me_ismb_2013}: A biologically inspired regularizer is used to capture task similarity.\\ \noindent\textbf{BSL-MTL}: This work: Bilinear sparse low-rank multitask learning

\begin{table*}[t]\caption{Area Under the Precision-Recall curve for each task in the two settings. X\% training indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data. We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100). The standard deviation is also shown. The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold. The first row is the only single-task method and all others are multitask models.}
\label{resultsTable}
\begin{small}
\begin{center}
\def\arraystretch{1.2}
\begin{tabular}{c|ccccccc}
\toprule
& \multicolumn{3}{c}{\textbf{10\% training}} && \multicolumn{3}{c}{\textbf{30\% training}} \\ \cline{2-4}\cline{6-8}
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.220$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\
MMTL \cite{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\
Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\
SpLowRank \cite{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\
%Dirty model \cite{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\
MTPL \cite{me_ismb_2013} & \textbf{0.217}$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && \textbf{0.260}$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\
IMC \cite{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
\end{tabular}
\end{center}
\end{small}
\end{table*}

\subsection{Evaluation setup}
We first compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data. For the first setting we randomly split the labeled data from each task into 10\% training and 90\% test, such that the class-skew of 1:100 is maintained in both splits (i.e stratified splits). The second setting uses a 30\% training, 70\% test split (we use identical splits for all algorithms). 
In each setting we generate ten random splits and average the performance over the ten runs.
Next, we do a standard 10-fold cross validation (CV) experiment (8 folds to train, 1 fold as held-out, 1 fold as test data). In this setting, each algorithm has access to a much larger training set but a significantly smaller test set. The two prior settings (10\% and 30\%) portray a more realistic multitask scenario where we have access to little training data from each task.

We report the area under the precision recall curve (AUC-PR) along with the standard deviation. 
AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \cite{davis2006} such as ours. Please refer to the supplementary for parameter tuning.

\section{Results}
Table \ref{resultsTable} has the AUC-PR for all methods. 
Note that the AUC-PR of a random classifier model is $\approx 0.01$.
The first row (STL) is the single-task baseline and all others are multitask models. In general, we notice that multitask learning benefits
all tasks. The first three columns show the results in the 10\% setting. 
%The number of training positive examples from each task are 8 for Ebola, 85 for Influenza and 98 for Hepatitis-C. 
Our model (last row) has significant gains for Influenza (1.4 times better than the next best) and modest improvements for the other tasks.
The variance in the performance is high for the Ebola task (column 1) owing to the small number of positives 
in the training splits (8 positives).
The most benefits for our model are seen in the 30\% setting for all tasks, with 
improvements of 39, 3 and 12 percentage points on the \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} tasks, respectively. \textit{Ebola}, the data-poorest task, benefits the most.
10 fold CV results are in the supplementary.

\subsection{Biological significance of the model}
\label{bioanalysis}
The model parameters $U$, $V$ and $S$ are a source of rich information which can be used to further understand host-pathogen
interactions. Note that our features are derived from the amino acid sequences of the proteins which provide
opportunities to interpret the parameters.

\subsubsection{Clustering proteins based on interaction propensities}
We analyze the proteins by projecting them using the model parameters $U$ and $V$ into a lower dimensional subspace 
(i.e computing $X U^\intercal$ and $Y V^\intercal$ to get projections of the virus and human proteins respectively).  
The principal component analysis (PCA) of this lower dimensional representation is compared with PCA in the original feature space (protein sequence features) in Fig \ref{fig:pca}. 
Firstly, the projected data has a much better separation than the original data.
Secondly, Fig \ref{fig:pca} (right) tells us that Hepatitis-C and Influenza have many proteins with similar binding tendencies, and that these behave differently than most Ebola virus proteins. This observation is not obvious in the PCA of the original feature space (Fig \ref{fig:pca} left), where proteins with similar sequences cluster together.
These clusters of proteins can be analyzed further for enrichment of Gene Ontology (GO) annotations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
\begin{floatrow}
\ffigbox[0.65\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.27, trim = 0 0.8cm 0 0]{figures/fig2-eps-converted-to.pdf}
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/fig3-eps-converted-to.pdf}
}{%
\caption{Principal component analysis (PCA) of virus proteins in the original feature space (left) and the projected subspace (right). The first two principal components are shown. Shape of the points indicates which virus that protein comes from.}%
\label{fig:pca}
}
\ffigbox[0.3\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.2, trim = 0 0.8cm 0 0]{figures/fig4-eps-converted-to.pdf}
}{%
\caption{Top tri-mer sequence motifs from virus proteins across all three viruses.}%
\label{fig:virus_3mer_motifs}
}
\end{floatrow}
\end{figure}

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.18, trim = 0 0 0 0]{figures/fig5-eps-converted-to.pdf}
}{%
\caption{Sequence motif constructed from the top four-mer features of virus proteins across all three viruses.}%
\label{fig:virus_4mer_motifs}
}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.18, trim = 0 0 0 0]{figures/fig6-eps-converted-to.pdf}
}{%
\caption{Sequence motif constructed from the top four-mer features of human proteins}%
\label{fig:human_motifs}
}
\end{floatrow}
\end{figure}

\subsubsection{Sequence motifs from virus proteins}
In Figures \ref{fig:virus_3mer_motifs}-\ref{fig:human_motifs}, we show sequence motifs derived from the top $k$-mers that contribute to interactions. The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to 
compute these motifs. The top positive-valued entries from the product $U V^T$ indicate which pairs of features: ($(f_v, f_h)$: virus protein feature, human protein feature) are important for interactions across all the virus-human PPI tasks.
Analogously, the entries from $S_t$ give us pairs of features important to a particular virus-human task `$t$'.
We find that most of the top entries from $U V^T$ correspond to linear virus features, whereas those from the various $S_t$
involve bilinear features. We analyze the $k$-mers corresponding to the top 20 features from each of the matrices.

Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}): 
the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged). Given the set of top virus features we can obtain the corresponding set of 
amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the 
training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
be the set of amino-acid k-mers that appear in the training data. Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction. 
To summarize $I_v$, we use a popular tool Seq2Logo \cite{seq2logo} to generate a sequence motif. 
The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
analysis). Figures \ref{fig:virus_3mer_motifs} and \ref{fig:virus_4mer_motifs} show the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.

This procedure described above is used to analyze the most significant human protein features, obtained from
the matrix $U V^T$. These motifs are shown in Figure \ref{fig:human_motifs}. We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_3mer_motifs} is dominated by 
hydrophilic amino acids (T, K, R, D, E). All other motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G)
though S and T do appear in some motifs as well.
%The human protein motifs are shown in \ref{fig:human_motifs}; the 3-mer motif show less of a pattern, in particular the first residue does not seem to have any dominant amino-acids.
%However the 4-mer seems to be dominated by hydrophobic amino-acids L, P, I, F in the first three positions and hydrophilic residues S, T, Y in the fourth.
The human protein motifs are shown in Fig \ref{fig:human_motifs}. In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
Further analysis is in the supplementary.

\subsubsection{Novel interactions with Ebola proteins}
The top four Ebola-human PPI are all predictions for the Ebola envelope glycoprotein (GP) with four different human proteins (Note: GP is not in the gold standard PPIs). There is abundant evidence in the published literature \cite{nanbo2010} for the critical role played by GP in virus docking and fusion with the host cell.

\section{Conclusions and future extensions}
%Multitask link prediction is an important area with many applications ranging from recommendation systems to biomedical host-pathogen interactions.  
This work developed and tested a new method based on low-rank matrix completion for sharing information across tasks
%and showed significant increases in prediction performance. 
The method was evaluated in the host-pathogen protein interaction domain for three pathogens (three tasks) and exhibited significant increases in prediction accuracy. The model parameters provide several avenues for
studying host-pathogen interactions for biologists that can lead to interesting observations and insights. Finally, the model we present is general enough to be applicable on other problems such as: gene-disease relevance prediction across organisms or disease conditions.

\scriptsize{
\bibliographystyle{plain}
%\bibliographystyle{natbib}
%\bibliographystyle{agsm}
\bibliography{references}}
%\end{scriptsize}

\end{document}

### 2016-08-21 camera_ready/supplementary.tex
first commit

\documentclass[a4paper,11pt]{article}

\usepackage{subfigure} 
\usepackage{floatrow}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{url}

\begin{document}

\title{Supplementary Information}

\section{Parameter tuning}
We tune the hyper parameters using a 3 fold cross-validation (CV) on the training split. For all baseline regularization parameters %(or a grid search based on the method): 
we tried the range: [$100, 50, 10, 1, 0.1, 0.01, 0.05, 10^{-3}$]. To address the class-skew we 
assign a higher weight to the positives. For BSL-MTL, to tune the rank parameter '$k$' we tried: [1, 5, 10, 25, 40, 60, 100] and the regularization parameter controlling the norm of $U$ and $V$ was tuned over the range $\lambda$=$\{1 ... 10^{-3}\}$. For each task $t$, $\sigma_t$ was varied over the values $[10^{-3}, 10^{-4}, 10^{-5}, 10^{-6}]$ and
$\mu_t$ was varied over $\{0, 0.25, 0.5, 0.75, 1\}$.
The optimal setting was:  $k=10$. %, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.

\section{10 fold cross validation}
In the 10 fold cross-validation (CV) experiments, a much larger training set is available. So the single task baseline becomes much harder to beat as it can independently tune parameters for each task. All methods have a higher variance in their performance (we think this is due to the smaller and hence more variable test data in a 10 fold CV setting). Our method improves on only the \textit{Hep-C} task.

\begin{table*}[h]\caption{Area Under the Precision-Recall curve for 10 fold CV. The first row is the only single-task method and all others are multitask models.}
\label{resultsTable}
\begin{center}
\def\arraystretch{1.2}
\begin{tabular}{c|ccc}
\toprule
& \multicolumn{3}{c}{\textbf{10 fold CV}} \\
& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
 STL     & 0.28$\pm$.11 & 0.74$\pm$.05 & 0.66$\pm$.08  \\ 
 MMTL    & 0.28$\pm$.11 & 0.68$\pm$.04 & 0.40$\pm$.03  \\ 
 Norm    & 0.20$\pm$.09 & 0.70$\pm$.06 & 0.44$\pm$.05  \\ 
 Rank    & 0.18$\pm$.08 & 0.70$\pm$.06 & 0.45$\pm$.04  \\ 
 MTPL    & 0.27$\pm$.08 & 0.67$\pm$.05 & 0.48$\pm$.06  \\ 
 IMC     & 0.16$\pm$.08 & 0.72$\pm$.06 & 0.45$\pm$.06  \\ 
 BSL-MTL & 0.28$\pm$.11 & 0.82$\pm$.05 & 0.62$\pm$.02  \\ 
\bottomrule
\end{tabular}
\end{center}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Homologous interactions}
Since we use data from several strains for each task, the PPI data contains some interactions that are interologs.
We observed this for two of the tasks: \textit{Hepatitis-C} and \textit{Influenza A}. Note that we did not find any interologs \textit{across} tasks. These homologous interactions in the various strains are reported by different experimental studies and we believe their presence suggests the confidence of such interactions. Since we use only PPI derived from experimental methods (and not electronically inferred), we do not exclude any homologous interactions from our training. \\
\textbf{Removing homologs from the evaluation}:\\
\noindent The number of distinct PPI for which interologs or homologs exist is very small ($\approx$20), but there are several homologs for each such PPI. Note that any benefits achieved from the presence of homologous interactions will be available to all the methods (since the models were built using identical train/test data). Here we present results on the 10\% setting by removing all homologs from the test data (note: the \textit{Ebola} task did not have any homologs). For lack of space we only mention the trend for our method (BSL-MTL) - \textit{Hep-C}: 0.85 and \textit{Influenza:} 0.45. The other methods show a similar trend; our method continues to outperform by significant margins.

\section{Biological significance of interactions} 
 \noindent\emph{Phosphorylation sites}: 
 We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
 related to the amino acids being frequent targets of phosphorylation.
 Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \cite{phosphosite} are
  repositories for known sites in human proteins. Since these are sites in human proteins, we searched for the patterns
	from the 4-mer motif in Fig \ref{fig:human_motifs} and found several to be flanking known phosphorylation sites in human proteins:
	\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site). This observation also supports the notion that the motifs predictive
	of interaction are biologically significant.

\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \cite{iedb}.
The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:hepc_motifs} is found in 53 epitopes.
From the \textit{Ebola} motifs in Fig \ref{fig:ebola_motifs}, we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes. \texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
is a further validation of its performance.
Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0.2cm 0 0 0]{figures/fig7-eps-converted-to.pdf}
}{%
\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
\label{fig:ebola_motifs}
}

\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/fig8-eps-converted-to.pdf}
}{%
\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}

\end{floatrow}
\end{figure}

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/fig9-eps-converted-to.pdf}
}{%
\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
\label{fig:flu_motifs}
}
\end{floatrow}
\end{figure}

\newpage

\section{Tri-mers}

Below, we show the sequence motifs from the tri-mers found to be highly relevant to predicting interactions between human and viral proteins.

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top10ebola_3mers_seq2logo.eps}
}{%
\caption{Sequence motifs specific to \textit{Ebola} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}

\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top10flavi_3mers_seq2logo-eps-converted-to.pdf}
}{%
\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
\label{fig:hepc_motifs}
}
\end{floatrow}
\end{figure}

\begin{figure}[h]
\begin{floatrow}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top10flu_3mers_seq2logo-eps-converted-to.pdf}
}{%
\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
\label{fig:flu_motifs}
}
\ffigbox[0.4\columnwidth]{%
%trim option's parameter order: left bottom right top
\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top20human_3mers_seq2logo-eps-converted-to.pdf}
}{%
\caption{Sequence motif constructed from the top tri-mer features of human proteins}%
\label{fig:human_motifs}
}
\end{floatrow}
\end{figure}

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2016-08-21 paper.tex
first commit

--- 2016-08-21 camera_ready/paper.tex
@@
 \documentclass[runningheads,a4paper]{llncs}
-%\usepackage[top=1.3in, bottom=1.3in, left=1.3in, right=1.3in]{geometry}
+\usepackage[top=1.3in, bottom=1.3in, left=1in, right=1in]{geometry}
 \usepackage{floatrow}
@@
 \usepackage{subfigure} 
-%\usepackage{natbib}
+\usepackage{natbib}
 \usepackage[footnotesize]{caption}
 \usepackage{multirow}
-\usepackage{makeidx}
 
@@
 
-\newcommand{\specialcell}[2][c]{%
-  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \usepackage{url}
-\urldef{\mailsa}\path|mkshirs@us.ibm.com|
-\urldef{\mailsb}\path|{jgc,judithks,kmuruges}@cs.cmu.edu|
+\urldef{\mailsa}\path|{mkshirsa,jgc,judithks,kmuruges}@cs.cmu.edu|
 %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%\setcounter{secnumdepth}{0}
+\setcounter{secnumdepth}{0}
 
 \begin{document}
-\pagestyle{headings}
 
@@
 \author{Meghana Kshirsagar$^{1}$%
-\and Jaime G. Carbonell$^{2}$
-\and Judith Klein-Seetharaman$^{3}$
-\and Keerthiram Murugesan$^{2}$}
+\and Jaime G. Carbonell$^{1}$
+\and Judith Klein-Seetharaman$^{2}$
+\and Keerthiram Murugesan$^{1}$}
 
@@
 
-\institute{$^{1}$IBM T. J. Watson Research, Yorktown Heights, NY 10598, USA \\
-$^{2}$Language Technologies Institute, Carnegie Mellon University, 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
-$^{3}$Metabolic \& Vascular Health, Warwick Medical School, University of Warwick, Coventry, UK\\
-\mailsa \quad \mailsb
+\institute{$^{1}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
+$^{2}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK\\
+\mailsa
 }
@@
 Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
-We obtain upto 30 \% improvement in predictive performance over prior state-of-the-art models. 
+We obtain between 7 to 39 \% points improvement in predictive performance over prior state-of-the-art models. 
 We show how our model's parameters can be interpreted to reveal both general and specific
@@
 An elegant way to formulate the interaction prediction problem is via a graph completion based framework, where we have several bipartite 
-graphs over multiple hosts and pathogens.% as illustrated in Figure \ref{fig:hpppigraph}.
-Nodes in the graphs represent host and pathogen proteins, with edges between them representing interactions 
+graphs over multiple hosts and pathogens as illustrated in Figure \ref{fig:hpppigraph}.
+Nodes in the graphs represent host proteins (circles) and pathogen proteins (triangles), with edges between them representing interactions 
 (host protein $\frac{interacts}{}$ pathogen protein).
@@
 The matrix entry $M_{ij}$ encodes the edge between pathogen protein $i$ and host protein $j$ from the graph, with $M_{ij}=1$ for the observed interactions. 
-Thus, the graph completion problem can be mathematically modeled as a matrix completion problem \cite{candes08}. 
+Thus, the graph completion problem can be mathematically modeled as a matrix completion problem \citep{candes08}. 
 %Traditional approaches towards matrix completion rely on the assumption that the underlying function that generated the matrix can be decomposed into a small number of `latent' factors.
@@
 fact that similar viruses will have similar strategies to infect and hijack biological processes in the human body.
-%Such opportunities for sharing arise in other applications as well: for instance, predicting user preferences in movies may inform preferences in selection of books, or vice-versa, as movies and books are semantically related.
-Multitask learning based models that incorporate and exploit these correlations should benefit from the additional information.
-% and result in a better prediction performance. 
-Our model is motivated by the following biological intuition governing protein interactions across diseases.
+Such opportunities for sharing arise in other applications as well: for instance, predicting user preferences in movies may inform preferences in selection of books, 
+or vice-versa, as movies and books are semantically related. Multitask learning based models that incorporate and exploit these correlations should benefit from 
+the additional information.% and result in a better prediction performance.
+
+Our multitask matrix completion based model is motivated by the following biological intuition governing protein interactions across diseases.
 \begin{enumerate}
 \item An interaction depends on the structural properties of the proteins, which are conserved across similar viruses as they have evolved from common ancestors.
-We use a component to capture these latent similarities, that is \textit{shared} across tasks.
+Our model thus needs a component to capture these latent similarities, which is \textit{shared} across tasks.
 %This latent structural similarity can be expressed via latent factors that project the proteins into a shared lower dimensional subspace that is common to all tasks.
 \item In addition to the shared properties discussed above, each pathogen has also evolved specialized mechanisms to target host proteins.
-These are unique to the pathogen and can be expressed using a \textit{task-specific} parameter.
+These are unique to the pathogen and can be expressed using a \textit{task-specific} parameter in the model.
 \end{enumerate}
@@
 
-\begin{table}
-\caption{Tasks and their sizes.
-Each column corresponds to one bipartite graph between human proteins and the pathogen indicated in the column header.
-All pathogens are single stranded RNA viruses.
-The interactions and the protein count both includes data across various strains of each pathogen}
-% (see Sec \ref{sec:datasets} for details). The last row shows that each of our graphs is extremely sparse.} 
-\label{datasets}
+\begin{figure*}
+\begin{floatrow}
+\ffigbox[0.29\textwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.33, trim = 0.2 9.2cm 4cm 0]{fig1.eps}
+}
+{
+\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
+\label{fig:hpppigraph}
+}
+\capbtabbox[0.69\textwidth]{%
 \begin{footnotesize}
@@
 \multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
-\multicolumn{4}{l}{\scriptsize{$\ddagger$: considering all proteins from the two tasks involved. }}\\
+\multicolumn{4}{l}{\scriptsize{$\ddagger$: considering all proteins in the two organisms involved. }}\\
 \multicolumn{4}{l}{\scriptsize{Note: considering the total number of human proteins to be $\approx$ 100,000.}}\\
@@
 \end{footnotesize}
-\end{table}
+}
+{
+\caption{Tasks and their sizes.
+Each column corresponds to one bipartite graph between human proteins and the pathogen indicated in the column header.
+All pathogens are single stranded RNA viruses.
+Row 4 shows that each of our graphs is extremely sparse.}
+\label{datasets}
+}
+\end{floatrow}
+\end{figure*}
 
@@
 our model in detail. We trade-off the relative importance of the two components using task-specific hyperparameters.
-We can thus learn what is conserved and what is different across pathogens, rather than having to specify it manually.
+Our model can thus learn what is conserved and what is different across pathogens, rather than having to specify it manually.
 
-%One of the key challenges in inducing such a model is: in addition to the graph structure itself, we want to incorporate information about the nodes available in the form of features. 
-The applications that we consider involve extremely sparse graphs with a large number of nodes and very few observed edges.
-There will be nodes i.e proteins that are not involved in any known interactions -- the model should be able to predict links between such prior `unseen' node pairs (called the \textit{cold start problem} in the recommendation systems community).
-For instance, the host-pathogen PPI network of human-Ebola virus (column-3, Table \ref{datasets}) has $\approx$ 90 observed edges (equivalent to 0.06\% of the possible edges) which involve only 2 distinct virus proteins.
-Any biologist studying virus-human interactions will be more interested in the virus proteins which have yet unknown interactions.
-The main contributions of this work are:\\
-1. We extend a prior matrix completion model \cite{abernethy} to the multitask setting.\\
-2. We leverage node features which allows us to predict on unseen nodes.\\
-3.
-We apply the model to an important problem -- prediction of interactions in disease-relevant host-pathogen protein networks, for multiple related diseases and demonstrate improved performance over prior state-of-the-art multitask models.\\
-4. We use unlabeled data to initialize the parameters of our model, which gives us a modest boost in prediction performance.
+The key challenges in inducing such a model are: (1) In addition to the interactions from each graph, it should exploit information available in the form of features. 
+(2) Exploiting features is particularly crucial since the graph $\mathcal{G}$ is often extremely sparse, i.e there are a large number of nodes and very few edges are observed. 
+There will be proteins (i.e nodes) that are not involved in any known interactions -- called the \textit{cold start problem} in the recommendation systems community. 
+The model should be able to predict the existence of links (or their absence) between such prior `unseen' node pairs. 
+This is of particular significance in graphs that capture biological phenomena. For instance, the host-pathogen PPI network
+of human-Ebola virus (column-3, Table \ref{datasets}) has $\approx$ 90 observed edges (equivalent to 0.06\% of the possible edges) which involve only 2 distinct virus proteins.
+Any biologist studying virus-human interactions will be interested in knowing more about the hundreds of other virus proteins (which have yet unknown interactions).
+(3) A side-effect of having scarce data is the availability of a large number of unlabeled examples, 
+i.e pairs of nodes with no edge between them. 
+These unlabeled examples can contain information about the graph as a whole, and a good model should be able to use them.
+%(4) ---- class skew ----
+
+The main contributions of this work are:
+\begin{itemize}
+\item We extend a prior matrix completion model \citep{abernethy} to the multitask setting. This extension is new.
+\item Unlike most prior approaches, our model exploits node-based features which allows us to deal with the `cold start' problem (generating predictions on unseen nodes).
+\item We apply the model to an important, real-world problem -- prediction of interactions in disease-relevant host-pathogen protein networks, for multiple related diseases.
+We demonstrate the superior performance of our model over prior state-of-the-art multitask models.
+\item We use unlabeled data to initialize the parameters of our model, which serves as a prior. This gives us a modest boost in prediction performance.
+\end{itemize}
 
@@
 The experimental discovery of host-pathogen protein interactions involves biochemical and biophysical methods such as co-immunoprecipitation (co-IP), yeast two-hybrid (Y2H) assays, co-crystalization.
-% These databases are valuable sources of information to bioinformaticians for developing models.
+The host-pathogen protein interactions from several small-scale and high throughput experiments are aggregated by databases such as virus-MINT\cite{virusmint}, HPIDB \cite{hpidb}, PHISTO \cite{phisto} etc by literature curation.
+These databases are valuable sources of information to bioinformaticians for developing models.
 \\
-The most reliable experimental methods are often very time-consuming and expensive, making it hard to investigate the prohibitively large set of possible host-pathogen interactions -- e.g.
-the bacterium \textit{Bacillus anthracis} with about 2321 proteins when coupled with the 100,000 human proteins gives $\approx$232 million protein pairs to validate.
+
+\noindent\textbf{Prediction of host-pathogen PPIs}: The most reliable experimental methods for studying protein-protein interactions (PPI) are often very time-consuming and expensive, making it hard to investigate the prohibitively large set of possible host-pathogen interactions -- for example, the bacterium \textit{Bacillus anthracis} which causes anthrax has about 2321 proteins which when coupled with the 100,000 or so human proteins gives $\approx$232 million protein pairs to test, experimentally.
 Computational techniques complement laboratory-based methods by predicting highly probable PPIs.
-Supervised machine learning based methods use the few known interactions as training data and formulate the interaction prediction problem in a classification setting%, with target classes: ``interacting'' or ``non-interacting''.
-Features are derived using various attributes of the two proteins.
-%such as: protein sequences from Uniprot \cite{uniprot}, protein structure from PDB, gene ontology from GO database \cite{GO}.
-%In our work we use protein-sequence based features.
+These techniques use the known interactions data from previous experiments and predict the most plausible new interactions.
+In particular, supervised machine learning based methods use the few known interactions as training data and formulate the interaction prediction problem in a classification setting, with target classes: ``interacting'' or ``non-interacting''.
+Features are derived using various attributes of the two proteins such as: protein sequences from Uniprot \cite{uniprot}, protein structure from PDB, gene ontology from GO database \cite{GO}.
 
@@
 Most of the prior work in PPI prediction has focussed on building models separately for individual 
-organisms \cite{struct2net,qi06} or on building a model specific to a disease in the case of host-pathogen PPI 
-prediction \cite{tastan09,dyer07,me12}.
-There has been limited work on combining PPI datasets to learn joint models.
+organisms \citep{chen,yeast,struct2net,qi06} or on building a model specific to a disease in the case of host-pathogen PPI 
+prediction \citep{tastan09,qi09,dyer07,me12}. There has been little work on combining PPI datasets with the goal of improving 
+prediction performance for multiple organisms.
 \cite{qi2010} proposed a semi-supervised multi-task framework to predict PPIs from partially labeled reference sets.
 \cite{me_ismb_2013} develop a task regularization based
-framework that incorporates the similarity in biological pathways targeted by various diseases.
-%  to couple multiple tasks together.
-%Matrix factorization based protein-protein interaction (PPI) prediction has seen very little work, mainly due to the extremely sparse nature of these datasets which makes it very difficult to get reliable predictors.
-\cite{ppi_cmf} uses a collective matrix factorization based approach in a multi-task learning setting for within species PPI prediction. The methods used 
+framework called MTPL that incorporates the similarity in biological pathways targeted by various diseases to 
+couple multiple tasks together.
+Matrix factorization based protein-protein interaction (PPI) prediction has seen very little work, mainly due to the extremely sparse nature of these datasets which makes it very difficult to get reliable predictors.
+\cite{ppi_cmf} use a CMF-based approach in a multi-task learning setting for within species PPI prediction.
+The methods used
 in all prior work on PPI prediction do not explicitly model the features of the proteins and cannot be applied on proteins which have no known interactions available.
@@
 %in using a formulation of the matrix completion problem originally proposed in the work by \cite{abernethy}.
-%A majority of the prior work in the relevant areas of collaborative filtering and link prediction includes single relation models that use neighbourhood based prediction \cite{sarwar01}, matrix factorization based approaches \cite{koren09,menon_ecml11} and bayesian approaches using graphical models \cite{jin02,phung09}.
+
+A majority of the prior work in the relevant areas of collaborative filtering and link prediction includes single relation models that use neighbourhood based prediction \citep{sarwar01}, matrix factorization based approaches \citep{koren09,menon_ecml11} and bayesian approaches using graphical models \citep{jin02,phung09}.
 %The matrix factorization based methods have been more popular than the other methods. 
-%There have also been multitask approaches on link prediction \cite{zhang12,cao_icml2010,li09,cmf}. 
+There have also been multitask approaches on link prediction \citep{zhang12,cao_icml2010,li09,cmf}. 
 %\cite{li11} presents a survey on multitask/transfer methods in collaborative filtering. 
@@
 %has been proposed for single-graph link prediction problems. 
-%\cite{menon_ecml11} propose a single-graph model that combines linear and bilinear features, latent parameters on the nodes and several other parameters into a function that minimizes a ranking loss.
+\cite{menon_ecml11} propose a single-graph model that combines linear 
+and bilinear features, latent parameters on the nodes and several other parameters into a function that minimizes a ranking loss. 
 %Probabilistic models to combine network structure with node features have also been proposed. 
-We extend the framework from \cite{abernethy} to the multitask setting. 
-%There has been a lot of work on other low-rank models for multitask learning \cite{ando2005,ji2009,chen2012,chen2013}. 
+\cite{abernethy} cast the problem of matrix completion in terms of the abstract problem of learning linear operators.
+Their framework allows the incorporation of features and kernels.
+We extend their bilinear model for the multitask setting.
+There has been a lot of work on other low-rank models for multitask learning \citep{ando2005,ji2009,chen2012,chen2013}.
 % that try to capture the task relationship via a shared low-rank structure over the model parameters.
@@
 experimentally for various reasons. Hence we derived a set of `probable negatives' using a heuristic often used in PPI prediction 
-work \cite{qi06,qi09,dyer11,me_ismb_2013}. We pair up all virus proteins with all human proteins and sample a random set to be negatives. 
+work \citep{qi06,qi09,dyer11,me_ismb_2013}. We pair up all virus proteins with all human proteins and sample a random set to be negatives. 
 This heuristic works in practice as the interaction ratio (i.e number of positives in a large random set of protein pairs) is expected to be very low:
@@
 \label{sec:datasets}
-We use three human-virus PPI datasets from the PHISTO \cite{phisto} database (version from 2014), the characteristics of which are summarized in Table \ref{datasets}. 
+We use three human-virus PPI datasets from the PHISTO \citep{phisto} database (version from 2014), the characteristics of which are summarized in Table \ref{datasets}. 
 The \textit{Influenza A} task includes various strains of flu: H1N1, H3N2, H5N1, H7N3. 
 Similarly, the \textit{Hepatitis} task includes various strains of the virus. 
-\footnote{Since we use data from several strains for each task, the PPI data contains some interactions that are interologs.
-Please see the supplementary for details}
 All three are single-strand RNA viruses, with \textit{Hepatitis} being a positive-strand ssRNA whereas \textit{Influenza} and \textit{Ebola} are negative-strand viruses.
-%Note that we use UniprotKB identifiers to refer to the proteins; hence if an identical protein appears in two strains it will be treated differently as it has a different UniprotKB id.
-The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (last row in Table \ref{datasets}).\\
+Note that we use UniprotKB identifiers to refer to
+the proteins; hence if an identical protein appears in two strains it will be treated differently as it has a different
+UniprotKB id. The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (row-4 in Table \ref{datasets}). \\
 \noindent\textbf{Features}: 
@@
 We use di-, tri- and four-mers thus leading to a total of 2793 features ($7^2 + 7^3 + 7^4$).
-Such features have been successfully applied in prior work \cite{dyer07,me_ismb_2013}.
+Such features have been successfully applied in prior work \citep{dyer07,me_ismb_2013}.
 
 \subsection{Competing methods}
-The results upon comparing our model to various single and multitask methods are shown in Table \ref{resultsTable}: including other recent low-rank and sparse models, conventional multitask methods and prior work on HP PPI prediction.
-Wherever appropriate, we concatenated the features of the two node types into a single feature vector.
+The results upon comparing our model to various single and multitask methods are shown in Table \ref{resultsTable}.
+Our baselines include recent low-rank and sparse models, conventional multitask methods and prior work on HP PPI prediction.
+For a uniform comparison we used least squared loss in all the methods.
+The MALSAR package was used to implement some of the models.
+For the baselines wherever appropriate, we concatenated the features of the two node types into a single feature vector.
 Let $W \in \mathbb{R}^{T \times d_t}$ be the matrix with the task-specific weight vectors $\mathbf{w}_t$.
-For a uniform comparison we used least squared loss in all the methods.
-The MALSAR package was used for some of the baselines.
 
-\noindent\textbf{Single task (STL)}: Ridge regression with $\ell_2$ regularization (which performed better than $\ell_1$). \\
+\noindent\textbf{Single task (STL)}: We used ridge regression with $\ell_2$ regularization (which performed better than $\ell_1$). \\
 \noindent\textbf{MMTL}: The mean regularized multitask learning model from \cite{pontil04}. \\
-%\noindent\textbf{Dirty model} \cite{jalali2010}: This model assumes that $W = P + Q$, where $P$ enforces group sparsity and $Q$ is to control for element-wise sparsity.
+\noindent\textbf{Dirty model} \citep{jalali2010}: This model assumes that $W = P + Q$, where $P$ enforces group sparsity and $Q$ is to control for element-wise sparsity.
 It uses the regularizer $\rho_1 \|P\|_{1,\infty} + \rho_2 \| Q \|_1$.
 \\
-\noindent\textbf{Low rank model (TraceNorm)}: A low-rank structure is enforced on $W$ by minimizing the nuclear norm $\|W\|_*$. \\
-\noindent\textbf{Sparse + low-rank (SpLowRank)} \cite{chen2012}: $W$ is assumed to have the decomposition: $W = P + Q$, where $P$ is sparse and $Q$ has a low-rank structure. \\
-\noindent\textbf{IMC} \cite{prateek,nagarajan}: This is the link-prediction model from Section \ref{MCintro}, where data from all tasks is combined without incorporating any task relationships (comparable to the `union' setting from \cite{widmer}).
+\noindent\textbf{Low rank model}: A low-rank structure is enforced on $W$ by minimizing the nuclear norm $\|W\|_*$. \\
+\noindent\textbf{Sparse + low-rank} \citep{chen2012}: $W$ is assumed to have the decomposition: $W = P + Q$, where $P$ is sparse and $Q$ has a low-rank structure. \\
+\noindent\textbf{IMC} \citep{prateek,nagarajan}: This is the link-prediction model from Section \ref{MCintro}, where data from all tasks is combined without incorporating any task relationships.
 $U$ and $V$ are shared by all tasks.
@@
 \\
-\noindent\textbf{MTPL} \cite{me_ismb_2013}: A biologically inspired regularizer is used to capture task similarity.\\ \noindent\textbf{BSL-MTL}: This work: Bilinear sparse low-rank multitask learning
+\noindent\textbf{MTPL} \citep{me_ismb_2013}: A biologically inspired regularizer is used to capture task similarity.\\ \noindent\textbf{BSL-MTL}: Bilinear sparse low-rank multitask learning, the method developed in this paper.
 
@@
 The standard deviation is also shown.
-The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
-The first row is the only single-task method and all others are multitask models.}
+The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.}
 \label{resultsTable}
@@
 & \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.220$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\
-MMTL \cite{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\
+STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
+MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
 Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
-SpLowRank \cite{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\
-%Dirty model \cite{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\
-MTPL \cite{me_ismb_2013} & \textbf{0.217}$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && \textbf{0.260}$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\
-IMC \cite{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
-BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
+Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
+Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
+MTPL \citep{me_ismb_2013} & \textbf{0.217}$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && \textbf{0.260}$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
+IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
+BSL-MTL (this work) & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
 \end{tabular}
@@
 \subsection{Evaluation setup}
-We first compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data.
+We compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data.
 For the first setting we randomly split the labeled data from each task into 10\% training and 90\% test, such that the class-skew of 1:100 is maintained in both splits (i.e stratified splits).
-The second setting uses a 30\% training, 70\% test split (we use identical splits for all algorithms).
+The second setting uses a 30\% training, 70\% test split.
 In each setting we generate ten random splits and average the performance over the ten runs.
-Next, we do a standard 10-fold cross validation (CV) experiment (8 folds to train, 1 fold as held-out, 1 fold as test data).
-In this setting, each algorithm has access to a much larger training set but a significantly smaller test set.
-The two prior settings (10\% and 30\%) portray a more realistic multitask scenario where we have access to little training data from each task.
 
 We report the area under the precision recall curve (AUC-PR) along with the standard deviation. 
-AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \cite{davis2006} such as ours.
-Please refer to the supplementary for parameter tuning.
+AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \citep{davis2006} such as ours.
+
+%Cold-start: The test data can contain edges over nodes that were never seen during training.
+The number of such unseen nodes in the test data, on an average was: in the \textit{10\%} setting: 536.4 unseen nodes and in \textit{30\%}: ?????
+new nodes in test data.
+
+\subsection{Parameter tuning}
+We tune the hyper parameters using a 3 fold cross-validation (CV) on the training split. For all baseline regularization parameters %(or a grid search based on the method): 
+we tried the range: [$100, ...10^{-3}$]. To address the class-skew we 
+assign a higher weight to the positives.
+For BSL-MTL, to tune the rank parameter '$k$' we tried: 5, ...
+100 and the regularization parameter controlling the norm of $U$ and $V$ was tuned over the range $\lambda$=$\{1 ...
+10^{-3}\}$.
+For each task $t$, $\sigma_t$ was varied over the values $10^{-3}, ...
+10^{-6}$ and
+$\mu_t$ was varied over $\{0, 0.25, 0.5, 0.75, 1\}$.
+The optimal setting was:  $k=10$. %, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
+
+%For the task-specific regularization parameters, setting them all to a constant i.e $\forall t, \{\sigma_t\} = \sigma$ did not work well owing to the different sizes of the 3 tasks.
+For each task $t$, $\sigma_t$ was varied over the values $10^{-3}, 10^{-4}, 10^{-5}, 10^{-6}$.
+The optimal setting of parameters was found to be $k=10$, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
 
@@
 improvements of 39, 3 and 12 percentage points on the \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} tasks, respectively. \textit{Ebola}, the data-poorest task, benefits the most.
-10 fold CV results are in the supplementary.
+
+%\noindent\emph{Incorporating further distinct tasks}: We tried including a fourth task, \textit{HIV}-human PPI (which is very distinct from the other three, with \textit{HIV} being a retrovirus).
+Our model performs
+
+%On the Hepatitis-C task, the single-task learning method is already performing quite well indicating that this task has a lower variance.
+Hence, the improvement for this task is much lower by all the baselines.
+Yet the best performance is obtained by our methods.
+
+%Overall, we see great improvements for the two pathogens for which there is less data (Ebola) or less predictive data (Influenza) both over the single-task baseline and over other existing models.
+Results on the more predictive data (Hepatitis-C) were either as good or had smaller improvements over the baselines.
+This is consistent with the expectations for multi-task learning, where weakly-performing tasks are lifted more than strongly-performing tasks by borrowing from other related tasks.
+\\
 
@@
 \begin{floatrow}
-\ffigbox[0.65\columnwidth]{%
+\ffigbox[0.49\columnwidth]{%
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.27, trim = 0 0.8cm 0 0]{figures/fig2-eps-converted-to.pdf}
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/fig3-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0 0.8cm 0 0]{fig2.eps}
+\includegraphics[scale=0.23, trim = 0 0 0 0]{fig3.eps}
 }{%
@@
 }
-\ffigbox[0.3\columnwidth]{%
+\ffigbox[0.49\columnwidth]{%
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.2, trim = 0 0.8cm 0 0]{figures/fig4-eps-converted-to.pdf}
+\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4.eps}
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig5.eps}
 }{%
-\caption{Top tri-mer sequence motifs from virus proteins across all three viruses.}%
-\label{fig:virus_3mer_motifs}
-}
-\end{floatrow}
-\end{figure}
-
-\begin{figure}[h]
-\begin{floatrow}
-\ffigbox[0.4\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.18, trim = 0 0 0 0]{figures/fig5-eps-converted-to.pdf}
-}{%
-\caption{Sequence motif constructed from the top four-mer features of virus proteins across all three viruses.}%
-\label{fig:virus_4mer_motifs}
-}
-\ffigbox[0.4\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.18, trim = 0 0 0 0]{figures/fig6-eps-converted-to.pdf}
-}{%
-\caption{Sequence motif constructed from the top four-mer features of human proteins}%
-\label{fig:human_motifs}
+\caption{Top tri-mer and four-mer sequence motifs from virus proteins across all three viruses. These correspond to features important for interactions.}%
+\label{fig:virus_motifs}
 }
@@
 \subsubsection{Sequence motifs from virus proteins}
-In Figures \ref{fig:virus_3mer_motifs}-\ref{fig:human_motifs}, we show sequence motifs derived from the top $k$-mers that contribute to interactions.
+In Figures \ref{fig:virus_motifs}-\ref{fig:flu_motifs}, we show sequence motifs derived from the top $k$-mers that contribute to interactions.
 The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to
@@
 Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
-To summarize $I_v$, we use a popular tool Seq2Logo \cite{seq2logo} to generate a sequence motif. 
+To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif. 
 The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
@@
 with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
-analysis).
-Figures \ref{fig:virus_3mer_motifs} and \ref{fig:virus_4mer_motifs} show the motif that is common across viruses.
-%, while the other figures highlight those motifs important in a specific species.
+analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
 
 This procedure described above is used to analyze the most significant human protein features, obtained from
-the matrix $U V^T$. These motifs are shown in Figure \ref{fig:human_motifs}. We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_3mer_motifs} is dominated by 
+the matrix $U V^T$.
+These motifs are shown in Figure \ref{fig:human_motifs}.
+The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepC}$ and $S_{flu}$ respectively.
+Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
+These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and
+\ref{fig:flu_motifs}.
+
+We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by 
 hydrophilic amino acids (T, K, R, D, E). All other motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G)
@@
 In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
-Further analysis is in the supplementary.
+
+The motifs for tri-mers are shown in the supplementary.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+\begin{floatrow}
+\ffigbox[0.23\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig6.eps}
+}{%
+\caption{Sequence motif constructed from the top four-mer features of human proteins}%
+\label{fig:human_motifs}
+}
+
+\ffigbox[0.23\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0.2cm 0 0 0]{fig7.eps}
+}{%
+\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
+\label{fig:ebola_motifs}
+}
+
+\ffigbox[0.01\columnwidth]{}
+
+\ffigbox[0.23\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
+}{%
+\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
+\label{fig:hepc_motifs}
+}
+
+\ffigbox[0.23\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig9.eps}
+}{%
+\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
+\label{fig:flu_motifs}
+}
+\end{floatrow}
+\end{figure}
+
+\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
+We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \citep{iedb}.
+The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:hepc_motifs} is found in 53 epitopes.
+From the \textit{Ebola} motifs in Fig \ref{fig:ebola_motifs}, we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes.
+\texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
+Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
+is a further validation of its performance.
+Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
+features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
+Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.
+
+\noindent\emph{Phosphorylation sites}: 
+We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
+related to the amino acids being frequent targets of phosphorylation.
+Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \citep{phosphosite} are
+ repositories for known sites in human proteins. Since these are sites in human proteins, we searched for the patterns
+from the 4-mer motif in Fig \ref{fig:human_motifs} and found several to be flanking known phosphorylation sites in human proteins:
+\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site).
+This observation also supports the notion that the motifs predictive
+of interaction are biologically significant.
 
 \subsubsection{Novel interactions with Ebola proteins}
+%The correlations captured by $U V^\intercal$ and $S_t$ tell us which amino-acid $k$-mer pairs (one kmer from the host protein, one kmer from the pathogen protein) are important for interactions.
+Using these we can derive details of the putative protein interaction interfaces (i.e which parts of the molecules are in contact when the two proteins bind to each other).
 The top four Ebola-human PPI are all predictions for the Ebola envelope glycoprotein (GP) with four different human proteins (Note: GP is not in the gold standard PPIs).
-There is abundant evidence in the published literature \cite{nanbo2010} for the critical role played by GP in virus docking and fusion with the host cell.
+There is abundant evidence in the published literature \citep{nanbo2010} for the critical role played by GP in virus docking and fusion with the host cell.
+
+%Our model not only provides predictions on whether or not two proteins interact, but can also provide hypotheses 
+%as to the putative binding sites for the interaction. This is of significance for the viruses (esp. Ebola) as they have 
+%very few proteins with known 3D structures. Traditional linear models do not give us correlations between amino acid residues. 
+%We selected the top 10 Ebola-human PPI predictions (the novel interactions) from our model and analyzed their 3D structures (using the available structures in PDB). 
+%For each predicted PPI we performed protein-protein docking using the ClusPro algorithm. 
+%The sites that are within 10 the docking model also corresponded to the host-pathogen feature pairs 
+%that were responsible for that particular prediction. The 3D structure of one predicted interaction and it's binding interface is shown in Figure \ref{fig:3dstructure}.
+
+%\begin{figure}[h]
+%\caption{3D structure obtained by docking ebola virion spike glycoprotein (green) with human ubiquitin-protein ligase (cyan). The putative binding sites are shown using sticks.}%
+%\label{fig:3dstructure}
+%%trim option's parameter order: left bottom right top
+%\includegraphics[scale=0.3, trim = 0 0.8cm 0 0]{images/interaction_3d_structure.png}
+%\end{figure}
 
@@
 Finally, the model we present is general enough to be applicable on other problems such as: gene-disease relevance prediction across organisms or disease conditions.
+%As future work we intend to apply our method in some of these other settings.
 
-\scriptsize{
-\bibliographystyle{plain}
-%\bibliographystyle{natbib}
+%We envision many potential extensions to this work, two of which are: \\
+%{\em Applications:} We would like to apply our method on a multi-host (e.g.
+human, mouse, bovine, etc.) multi-pathogen graph to determine if the same multitask learning advantages accrue.
+Beyond protein interaction graphs, the next step is to apply and evaluate the method for different families of tasks, such recommendation systems data.\\
+%{\em Multiple levels:} Task hierarchies can be exploited by our model by incorporating additional components to reflect the hierarchical relationships. 
+%{\em Multiple link types:} We investigated one type of link, namely interaction links, but there can be different types of related links, such as between people and scientific articles (e.g.
+authorship, citation, has-read, dislikes, etc.).
+It will be interesting to incorporate these into our models.
+%  The multiple tasks here are predictions in different but correlated link spaces:  \textit{authorship} + \textit{citation} = \textit{self-citation}, citation is correlated with has-read, citation is anti-correlated with dislikes, but not totally, e.g.
+if contrasting one's work with a less preferred alternative.
+
+%We feel that multitask learning in link prediction is still in the early stages of research and hope this contribution will stimulate further work.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%\section*{Acknowledgement}
+
+%\paragraph{Funding\textcolon}
+
+%\input{appendix.tex}
+
+%\bibliographystyle{plainnat}
+\bibliographystyle{natbib}
 %\bibliographystyle{agsm}
-\bibliography{references}}
-%\end{scriptsize}
+\begin{scriptsize}
+\bibliography{references}
+\end{scriptsize}
 

### 2016-08-21 supplementary/supplementary.tex
first commit

--- 2016-08-21 camera_ready/supplementary.tex
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0.2cm 0 0 0]{figures/fig7-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0.2cm 0 0 0]{fig7-eps-converted-to.pdf}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/fig8-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{fig8-eps-converted-to.pdf}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/fig9-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{fig9-eps-converted-to.pdf}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top10ebola_3mers_seq2logo.eps}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10ebola_3mers_seq2logo.eps}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top10flavi_3mers_seq2logo-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flavi_3mers_seq2logo.eps}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top10flu_3mers_seq2logo-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flu_3mers_seq2logo.eps}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{figures/top20human_3mers_seq2logo-eps-converted-to.pdf}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top20human_3mers_seq2logo.eps}
 }{%

### 2016-08-22 paper.tex
some updates

@@
 \begin{footnotesize}
-\begin{tabular}{crrr} \toprule
-Task $\rightarrow$ & \textbf{\textit{Influenza A}} & \textbf{\textit{Hepatitis C}} & \textbf{\textit{Ebola}}\\ \midrule 
+\begin{tabular}{c|r|r|r} \toprule
+Task $\rightarrow$ &{\textit{Influenza A}} &\textit{Hepatitis C} & {\textit{Ebola}}\\
+& (28 strains) & (5 strains,  & (2 strains) \\
+& & many isolates) & \\ \midrule 
 number of HP PPIs (positives) & 848& 981 & 90 \\\hline
-\# of distinct virus proteins in PPIs & 54 & 151 & 2 \\\hline
-{\# of distinct human proteins in PPIs} & 362 & 385 & 88 \\\hline
-{total \# of virus proteins across strains} & 542 & 163 & 150 \\\hline
-{number of negatives} & 84800 & 98100 & 9000 \\\hline
-{density of observed graph$^\ddagger$ (as \%)} & .15 & .60 & .06 \\\hline
+number of non-homologous PPIs & 334 & 329 & 49 \\ \hline
+\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
+\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
+total \# of virus proteins in & 542 & 163 & 38 \\
+UniprotKB across strains $^\dagger$ & & & \\ \hline
+number of negatives & 84800 & 98100 & 9000 \\\hline
+density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
 \multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
-\multicolumn{4}{l}{\scriptsize{$\ddagger$: considering all proteins in the two organisms involved. }}\\
-\multicolumn{4}{l}{\scriptsize{Note: considering the total number of human proteins to be $\approx$ 100,000.}}\\
+\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The}} \\
+\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} denominator was computed considering all proteins from all strains of the virus and}}\\
+\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} $\approx$ 100,000 human proteins.}}\\
+\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniprotKB from all strains in our data}} \\
 \bottomrule
@@
 All pathogens are single stranded RNA viruses.
-Row 4 shows that each of our graphs is extremely sparse.}
+The last row shows that each of our graphs is extremely sparse.}
 \label{datasets}
@@
 of human-Ebola virus (column-3, Table \ref{datasets}) has $\approx$ 90 observed edges (equivalent to 0.06\% of the possible edges) which involve only 2 distinct virus proteins. 
-Any biologist studying virus-human interactions will be interested in knowing more about the hundreds of other virus proteins (which have yet unknown interactions).
 (3) A side-effect of having scarce data is the availability of a large number of unlabeled examples, 
@@
 
-\subsection{Competing methods}
-The results upon comparing our model to various single and multitask methods are shown in Table \ref{resultsTable}.
+\subsection{Comparison with other machine learning approaches}
+\label{mlmethods}
 Our baselines include recent low-rank and sparse models, conventional multitask methods and prior work on HP PPI prediction.
 For a uniform comparison we used least squared loss in all the methods.
-The MALSAR package was used to implement some of the models.
+The MALSAR \cite{malsar} package was used to implement some of the models.
 For the baselines wherever appropriate, we concatenated the features of the two node types into a single feature vector.
@@
 \noindent\textbf{MMTL}: The mean regularized multitask learning model from \cite{pontil04}. \\
-\noindent\textbf{Dirty model} \citep{jalali2010}: This model assumes that $W = P + Q$, where $P$ enforces group sparsity and $Q$ is to control for element-wise sparsity.
-It uses the regularizer $\rho_1 \|P\|_{1,\infty} + \rho_2 \| Q \|_1$.
-\\
-\noindent\textbf{Low rank model}: A low-rank structure is enforced on $W$ by minimizing the nuclear norm $\|W\|_*$. \\
 \noindent\textbf{Sparse + low-rank} \citep{chen2012}: $W$ is assumed to have the decomposition: $W = P + Q$, where $P$ is sparse and $Q$ has a low-rank structure. \\
@@
 
-\begin{table*}[t]\caption{Area Under the Precision-Recall curve for each task in the two settings.
-X\% training indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
+\subsection{Comparison with a homolog baseline}
+In addition to the machine learning methods presented in Section \ref{mlmethods}, we also tried a simple homology based approach that is based on protein sequence identity.
+We use the entire human-virus interactome from PHISTO as input to this method.
+Given a human protein $h$, we rank all virus proteins $v$ based on the following simple scheme.
+let $V(h)$ be a set of all virus proteins interacting with the human protein $h$ and let $H(v)$ be all human proteins interacting with the virus protein $v$.
+We score each pair $(h, v)$ from the bipartite graph as follows: $\max \{ \max $ sequence identity between $h$ and all host proteins in $H(v)$; $\max $ sequence identity between $ v $ and all virus proteins in $V(h) \}$.
+
+\subsection{Evaluation setup}
+We compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data.
+For the first setting we randomly split the labeled data from each task into 10\% training and 90\% test, such that the class-skew of 1:100 is maintained in both splits (i.e stratified splits).
+The second setting uses a 30\% training, 70\% test split.
+In each setting we generate ten random splits and average the performance over the ten runs.
+
+We report the area under the precision recall curve (AUC-PR) along with the standard deviation. 
+AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \citep{davis2006} such as ours.
+
+\begin{table*}[t]\caption{Area Under the Precision-Recall curve for each task on the PHISTO dataset without removing the homologs.
+The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
 We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
@@
 & \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & & & && & & \\
 STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
 MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
-Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
+%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
 Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
-Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
+%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
 MTPL \citep{me_ismb_2013} & \textbf{0.217}$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && \textbf{0.260}$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
@@
 \end{table*}
-
-\subsection{Evaluation setup}
-We compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data.
-For the first setting we randomly split the labeled data from each task into 10\% training and 90\% test, such that the class-skew of 1:100 is maintained in both splits (i.e stratified splits).
-The second setting uses a 30\% training, 70\% test split.
-In each setting we generate ten random splits and average the performance over the ten runs.
-
-We report the area under the precision recall curve (AUC-PR) along with the standard deviation. 
-AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \citep{davis2006} such as ours.
 
@@
 \section{Results}
+As described in Section \ref{}, the PPI data we obtain from PHISTO has homologs within a task.
+We therefore present results on two sets of data: (a) with the original PPIs from PHISTO shown in Table \ref{resultsTable} and (b) after removal of all homologs within each task in Table \ref{homresultsTable}.
+The latter comparison gives us an idea of how much the approaches benefit from the redundancy within each task.
+
 Table \ref{resultsTable} has the AUC-PR for all methods. 
@@
 
-%\noindent\emph{Incorporating further distinct tasks}: We tried including a fourth task, \textit{HIV}-human PPI (which is very distinct from the other three, with \textit{HIV} being a retrovirus).
-Our model performs
+\begin{table}[t]\caption{Area Under the Precision-Recall curve for each task upon removal of the homologs within each task, on the 10\% setting}
+\label{homresultsTable}
+\begin{small}
+\begin{center}
+\def\arraystretch{1.2}
+\begin{tabular}{c|ccc}
+\toprule
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & & & \\
+STL (Ridge Reg.)  & & & \\
+Sparse+low-rank \citep{chen2012} & & &  \\ 
+BSL-MTL (this work) & & &  \\ \bottomrule
+\end{tabular}
+\end{center}
+\end{small}
+\end{table}
 
-%On the Hepatitis-C task, the single-task learning method is already performing quite well indicating that this task has a lower variance.
-Hence, the improvement for this task is much lower by all the baselines.
-Yet the best performance is obtained by our methods.
+\subsection{Disadvantages of the Homolog method}
+In Figures \ref{fig:ebolanet}-\ref{fig:flunet} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
+The interactions mainly involve human proteins which do not have a homolog in the training data.
+This illustrates that our features capture more information than mere sequence similarity.
+Further, inclusion of interactions from other viruses in the naive manner as done by the homologs model is not helpful - our model is able to learn what to borrow from the other tasks accurately.
 
-%Overall, we see great improvements for the two pathogens for which there is less data (Ebola) or less predictive data (Influenza) both over the single-task baseline and over other existing models.
-Results on the more predictive data (Hepatitis-C) were either as good or had smaller improvements over the baselines.
-This is consistent with the expectations for multi-task learning, where weakly-performing tasks are lifted more than strongly-performing tasks by borrowing from other related tasks.
-\\
+\begin{figure*}
+\ffigbox[\textwidth]
+{
+\begin{subfloatrow}[3]
+\ffigbox[0.22\textwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{1cm}
+}
+{
+\caption{Ebola-human PPI}
+\label{fig:ebolanet}
+}
+\ffigbox[0.33\textwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.33, trim = 0.2 0.2 0.2 0.2]{flavinet.pdf}
+}
+{
+\caption{Hepatitis-human PPI}
+\label{fig:flavinet}
+}
+\ffigbox[0.34\textwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.35, trim = 0.2 0.2 0.2 0.2]{flunet.pdf}
+}
+{
+\caption{Influenza-human PPI}
+\label{fig:flunet}
+}
+\end{subfloatrow}
+}
+{
+\caption{PPI from each task that are found by our method but are missed by the homolog method}
+}
+\end{figure*}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \begin{figure}
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0.8cm 0 0]{fig2.eps}
-\includegraphics[scale=0.23, trim = 0 0 0 0]{fig3.eps}
+\includegraphics[scale=0.35, trim = 0 0.8cm 0 0]{fig2.eps}
 }{%
-\caption{Principal component analysis (PCA) of virus proteins in the original feature space (left) and the projected subspace (right).
+\caption{Principal component analysis (PCA) of virus proteins in the original feature space}%
+\label{fig:pca1}
+}
+\ffigbox[0.49\columnwidth]{%
+\includegraphics[scale=0.33, trim = 0 0 0 0]{fig3.eps}
+}{%
+\caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
 The first two principal components are shown.
 Shape of the points indicates which virus that protein comes from.}%
-\label{fig:pca}
-}
-\ffigbox[0.49\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4.eps}
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig5.eps}
-}{%
-\caption{Top tri-mer and four-mer sequence motifs from virus proteins across all three viruses. These correspond to features important for interactions.}%
-\label{fig:virus_motifs}
-}
-\end{floatrow}
-\end{figure}
-
-\subsubsection{Sequence motifs from virus proteins}
-In Figures \ref{fig:virus_motifs}-\ref{fig:flu_motifs}, we show sequence motifs derived from the top $k$-mers that contribute to interactions.
-The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to
-compute these motifs.
-The top positive-valued entries from the product $U V^T$ indicate which pairs of features: ($(f_v, f_h)$: virus protein feature, human protein feature) are important for interactions across all the virus-human PPI tasks.
-Analogously, the entries from $S_t$ give us pairs of features important to a particular virus-human task `$t$'.
-We find that most of the top entries from $U V^T$ correspond to linear virus features, whereas those from the various $S_t$
-involve bilinear features. We analyze the $k$-mers corresponding to the top 20 features from each of the matrices.
-
-Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}): 
-the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged).
-Given the set of top virus features we can obtain the corresponding set of
-amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the 
-training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
-be the set of amino-acid k-mers that appear in the training data.
-Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
-To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif. 
-The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
-want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
-with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
-analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
-
-This procedure described above is used to analyze the most significant human protein features, obtained from
-the matrix $U V^T$.
-These motifs are shown in Figure \ref{fig:human_motifs}.
-The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepC}$ and $S_{flu}$ respectively.
-Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
-These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and
-\ref{fig:flu_motifs}.
-
-We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by 
-hydrophilic amino acids (T, K, R, D, E). All other motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G)
-though S and T do appear in some motifs as well.
-%The human protein motifs are shown in \ref{fig:human_motifs}; the 3-mer motif show less of a pattern, in particular the first residue does not seem to have any dominant amino-acids.
-%However the 4-mer seems to be dominated by hydrophobic amino-acids L, P, I, F in the first three positions and hydrophilic residues S, T, Y in the fourth.
-The human protein motifs are shown in Fig \ref{fig:human_motifs}.
-In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
-
-The motifs for tri-mers are shown in the supplementary.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure}
-\begin{floatrow}
-\ffigbox[0.23\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig6.eps}
-}{%
-\caption{Sequence motif constructed from the top four-mer features of human proteins}%
-\label{fig:human_motifs}
-}
-
-\ffigbox[0.23\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0.2cm 0 0 0]{fig7.eps}
-}{%
-\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
-\label{fig:ebola_motifs}
-}
-
-\ffigbox[0.01\columnwidth]{}
-
-\ffigbox[0.23\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
-}{%
-\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
-\label{fig:hepc_motifs}
-}
-
-\ffigbox[0.23\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig9.eps}
-}{%
-\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
-\label{fig:flu_motifs}
+\label{fig:pca2}
 }
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Sequence motifs from virus proteins}
+In Figures \ref{fig:virus_motifs}-\ref{fig:flu_motifs}, we show sequence motifs derived from the top $k$-mers that contribute to interactions.
+The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to
+compute these motifs.
+The top positive-valued entries from the product $U V^T$ indicate which pairs of features: ($(f_v, f_h)$: virus protein feature, human protein feature) are important for interactions across all the virus-human PPI tasks.
+Analogously, the entries from $S_t$ give us pairs of features important to a particular virus-human task `$t$'.
+We find that most of the top entries from $U V^T$ correspond to linear virus features, whereas those from the various $S_t$
+involve bilinear features. We analyze the $k$-mers corresponding to the top 20 features from each of the matrices.
+
+Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}): 
+the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged).
+Given the set of top virus features we can obtain the corresponding set of
+amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the 
+training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
+be the set of amino-acid k-mers that appear in the training data.
+Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
+To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif. 
+The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
+want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
+with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
+analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
+
+This procedure described above is used to analyze the most significant human protein features, obtained from
+the matrix $U V^T$.
+These motifs are shown in Figure \ref{fig:human_motifs}.
+The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepC}$ and $S_{flu}$ respectively.
+Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
+These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and
+\ref{fig:flu_motifs}.
+
+We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by 
+hydrophilic amino acids (T, K, R, D, E). All other motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G)
+though S and T do appear in some motifs as well.
+%The human protein motifs are shown in \ref{fig:human_motifs}; the 3-mer motif show less of a pattern, in particular the first residue does not seem to have any dominant amino-acids.
+%However the 4-mer seems to be dominated by hydrophobic amino-acids L, P, I, F in the first three positions and hydrophilic residues S, T, Y in the fourth.
+The human protein motifs are shown in Fig \ref{fig:human_motifs}.
+In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
+
+The motifs for tri-mers are shown in the supplementary.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure}
+\begin{floatrow}
+\ffigbox[0.62\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4.eps}
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig5.eps}
+}{%
+\caption{Top tri-mer and four-mer sequence motifs from virus proteins across all three viruses. These correspond to features important for interactions.}%
+\label{fig:virus_motifs}
+}
+\ffigbox[0.31\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig6.eps}
+}{%
+\caption{Sequence motif constructed from the top four-mer features of human proteins}%
+\label{fig:human_motifs}
+}
+\end{floatrow}
+\begin{floatrow}
+\ffigbox[0.3\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0.2cm 0 0 0]{fig7.eps}
+}{%
+\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
+\label{fig:ebola_motifs}
+}
+
+\ffigbox[0.01\columnwidth]{}
+
+\ffigbox[0.3\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
+}{%
+\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
+\label{fig:hepc_motifs}
+}
+
+\ffigbox[0.3\columnwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig9.eps}
+}{%
+\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
+\label{fig:flu_motifs}
+}
+\end{floatrow}
+\end{figure}
+
 \section{Conclusions and future extensions}
@@
 Finally, the model we present is general enough to be applicable on other problems such as: gene-disease relevance prediction across organisms or disease conditions.
-%As future work we intend to apply our method in some of these other settings.
-
-%We envision many potential extensions to this work, two of which are: \\
-%{\em Applications:} We would like to apply our method on a multi-host (e.g.
-human, mouse, bovine, etc.) multi-pathogen graph to determine if the same multitask learning advantages accrue.
-Beyond protein interaction graphs, the next step is to apply and evaluate the method for different families of tasks, such recommendation systems data.\\
-%{\em Multiple levels:} Task hierarchies can be exploited by our model by incorporating additional components to reflect the hierarchical relationships. 
-%{\em Multiple link types:} We investigated one type of link, namely interaction links, but there can be different types of related links, such as between people and scientific articles (e.g.
-authorship, citation, has-read, dislikes, etc.).
-It will be interesting to incorporate these into our models.
-%  The multiple tasks here are predictions in different but correlated link spaces:  \textit{authorship} + \textit{citation} = \textit{self-citation}, citation is correlated with has-read, citation is anti-correlated with dislikes, but not totally, e.g.
-if contrasting one's work with a less preferred alternative.
-
-%We feel that multitask learning in link prediction is still in the early stages of research and hope this contribution will stimulate further work.
 

### 2016-08-31 paper.tex
old bioinfo template

@@
-\documentclass[runningheads,a4paper]{llncs}
-\usepackage[top=1.3in, bottom=1.3in, left=1in, right=1in]{geometry}
+\documentclass{bioinfo}
+\copyrightyear{2015} \pubyear{2015}
+
+%\usepackage[top=1.3in, bottom=1.3in, left=1in, right=1in]{geometry}
 \usepackage{floatrow}
@@
 \usepackage{natbib}
-\usepackage[footnotesize]{caption}
+%\usepackage[footnotesize]{caption}
 \usepackage{multirow}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newfloatcommand{capbtabbox}{table}[][11cm]
+\usepackage{url}
 
@@
 
-\usepackage{url}
-\urldef{\mailsa}\path|{mkshirsa,jgc,judithks,kmuruges}@cs.cmu.edu|
-%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newfloatcommand{capbtabbox}{table}[][11cm]
 \newcommand{\argmin}{\operatornamewithlimits{argmin}}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%
+%\setcounter{tocdepth}{3}
+%\setcounter{secnumdepth}{0}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\setcounter{tocdepth}{3}
-
-\setcounter{secnumdepth}{0}
+%\access{Advance Access Publication Date: Day Month Year}
+%\appnotes{Manuscript Category}
 
 \begin{document}
+\firstpage{1}
 
-\mainmatter
+%\subtitle{Subject Section}
 
-\title{Multitask matrix completion for learning protein interactions across diseases}
+\title[Multitask matrix completion]{Multitask matrix completion for learning protein interactions across diseases}
 
-\titlerunning{Multitask matrix completion}
-
-\author{Meghana Kshirsagar$^{1}$%
-\and Jaime G. Carbonell$^{1}$
-\and Judith Klein-Seetharaman$^{2}$
-\and Keerthiram Murugesan$^{1}$}
-
-\authorrunning{Multitask matrix completion}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-\institute{$^{1}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
-$^{2}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK\\
-\mailsa
+\author[Kshirsagar \textit{et~al}.]{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
+\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
+$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
+$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
 }
 
-\toctitle{Matrix completion for transfer learning in host-pathogen PPI networks}
-\tocauthor{Meghana Kshirsagar, Jaime G. Carbonell, Judith Klein-Seetharaman and Keerthiram Murugesan}
+%\corresp{$^\ast$To whom correspondence should be addressed.}
+
+\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
+
+\editor{Associate Editor: XXXXXXX}
+
+%\abstract{\textbf{Motivation:}
+%Disease causing pathogens such as viruses, introduce their proteins into the host cells 
+%where they interact with the host's proteins enabling the virus to replicate inside the host.
+%These interactions between pathogen and host proteins are key to understanding infectious diseases.
+%Often multiple diseases involve phylogenetically related or biologically similar pathogens.
+Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}.
+\\
+%\textbf{Results:}
+%Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
+%We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
+%We show how our model's parameters can be interpreted to reveal both general and specific
+%interaction-relevant characteristics of the viruses. \\
+%\textbf{Availability:}
+%Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
+%\textbf{Contact:}
+%\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
+%}
+
 \maketitle
@@
 \begin{abstract}
-
+\section{Motivation}
 Disease causing pathogens such as viruses, introduce their proteins into the host cells 
@@
 Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}.
+\section{Results}
 Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
-We obtain between 7 to 39 \% points improvement in predictive performance over prior state-of-the-art models. 
+We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
 We show how our model's parameters can be interpreted to reveal both general and specific
-interaction-relevant characteristics of the viruses. Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}
-\keywords{host-pathogen protein interactions, multitask learning, matrix completion}
+interaction-relevant characteristics of the viruses. \\
+\section{Availability}
+Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
+\section{Contact}
+\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
 \end{abstract}
@@
 This leads us to the following model that incorporates the above ideas.
-The interactions matrix $M_t$ of task $t$ can be written as: $M_t = \mu_t *\; \textrm{(shared component)} + (1-\mu_t) *\; \textrm{(specific component)} $, with hyperparameter $\mu_t$ allowing each task to customize its' amount of shared and specific components.
-
-\begin{figure*}
-\begin{floatrow}
-\ffigbox[0.29\textwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.33, trim = 0.2 9.2cm 4cm 0]{fig1.eps}
-}
-{
-\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
-\label{fig:hpppigraph}
-}
-\capbtabbox[0.69\textwidth]{%
-\begin{footnotesize}
-\begin{tabular}{c|r|r|r} \toprule
-Task $\rightarrow$ &{\textit{Influenza A}} &\textit{Hepatitis C} & {\textit{Ebola}}\\
-& (28 strains) & (5 strains,  & (2 strains) \\
-& & many isolates) & \\ \midrule 
-number of HP PPIs (positives) & 848& 981 & 90 \\\hline
-number of non-homologous PPIs & 334 & 329 & 49 \\ \hline
-\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
-\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
-total \# of virus proteins in & 542 & 163 & 38 \\
-UniprotKB across strains $^\dagger$ & & & \\ \hline
-number of negatives & 84800 & 98100 & 9000 \\\hline
-density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
-\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
-\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The}} \\
-\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} denominator was computed considering all proteins from all strains of the virus and}}\\
-\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} $\approx$ 100,000 human proteins.}}\\
-\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniprotKB from all strains in our data}} \\
-\bottomrule
-\end{tabular}
-\end{footnotesize}
-}
-{
-\caption{Tasks and their sizes.
-Each column corresponds to one bipartite graph between human proteins and the pathogen indicated in the column header.
-All pathogens are single stranded RNA viruses.
-The last row shows that each of our graphs is extremely sparse.}
-\label{datasets}
-}
-\end{floatrow}
-\end{figure*}
+The interactions matrix $M_t$ of task $t$ can be written as: $M_t = \mu_t *\; \textrm{(shared component)} + (1-\mu_t) *\; \textrm{(specific component)} $, with hyperparameter $\mu_t$ allowing each task to customize its amount of shared and specific components.
 
@@
 with these latent factors jointly learned over all tasks. The second component involves a task specific parameter, on which we
-additionally impose a sparsity constraint as we do not want this parameter to overfit the data. Section \ref{ourmodel} discusses
+additionally impose a sparsity constraint as we do not want this parameter to overfit the data. Section~\ref{ourmodel} discusses
 our model in detail. We trade-off the relative importance of the two components using task-specific hyperparameters.
@@
 In particular, supervised machine learning based methods use the few known interactions as training data and formulate the interaction prediction problem in a classification setting, with target classes: ``interacting'' or ``non-interacting''.
-Features are derived using various attributes of the two proteins such as: protein sequences from Uniprot \cite{uniprot}, protein structure from PDB, gene ontology from GO database \cite{GO}.
+Features are derived using various attributes of the two proteins such as: protein sequences from UniProt \cite{UniProt}, protein structure from PDB, gene ontology from GO database \cite{GO}.
 
@@
 
+\begin{figure*}
+\begin{floatrow}
+\ffigbox[0.28\textwidth]{%
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1.eps}
+}
+{
+\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
+\label{fig:hpppigraph}
+}
+\capbtabbox[0.68\textwidth]{%
+\begin{footnotesize}
+\begin{tabular}{c|r|r|r} 
+\hline
+Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
+& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
+number of HP PPIs (positives) & 848 & 981 & 90 \\\hline
+number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
+\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
+\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
+total \# of virus proteins in & 542 & 163 & 38 \\
+UniProtKB across strains $^\dagger$ & & & \\ \hline
+number of negatives & 84800 & 98100 & 9000 \\\hline
+density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
+\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
+\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed }} \\
+\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} considering all proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
+\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
+\end{tabular}
+\end{footnotesize}
+}
+{
+\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header).
+More importantly, each task represents several strains of one virus species.
+We pool together PPI from all strains as the number of interactions per strain is very small and not representative.
+PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus.
+The second row gives a conservative estimate of the number of non-homologous PPI.
+These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details).
+All pathogens are single stranded RNA viruses.
+The last row shows that each of our graphs is extremely sparse.}
+\label{datasets}
+}
+\end{floatrow}
+\end{figure*}
+
 \section{The bilinear sparse low-rank multitask model (BSL-MTL)} 
@@
 \begin{array}{ll}
-\mathcal{L}(U,V,\{S_t\}) & = \displaystyle{ \frac{1}{N} \sum_{t=1}^T \sum_{(i,j) \in \Omega_t}} c^t_{ij} \; \ell \big( M_{t_{ij}}, \mathbf{x}_{ti}^\intercal H_t \mathbf{y}_{tj} \big)^2 + \lambda ( \|U\|^2_F + \|V\|^2_F ) + \displaystyle{\sum_{t=1}^T \sigma_t \|S_t\|_1}
+\mathcal{L}(U,V,\{S_t\}) & = \displaystyle{ \frac{1}{N} \sum_{t=1}^T \sum_{(i,j) \in \Omega_t}} c^t_{ij} \; \ell \big( M_{t_{ij}}, \mathbf{x}_{ti}^\intercal H_t \mathbf{y}_{tj} \big)^2 + \\
+& \lambda ( \|U\|^2_F + \|V\|^2_F ) + \displaystyle{\sum_{t=1}^T \sigma_t \|S_t\|_1} 
 \end{array}
@@
 
-%\begin{algorithm}[h]   \caption{Alternating Least Squares}
-%   \label{alg}
-%   \begin{small}
-%\begin{algorithmic}[1]
-%\STATE \textbf{Input}: \\
-%$k$ : number of latent factors\\
-%$\Gamma$: pairs of entities for initialization\\
-%For every task $t$,\\
-%${\{\mathbf{x}_{ti}\}, \{\mathbf{y}_{tj}\}}$: feature vectors for pathogen and host proteins resp.\\
-%$\Omega_t$: the observed entries of the matrix $M_t$\\
-%\STATE \textbf{Initialization}: \\
-%\STATE An iteration $r$, let $\mathbf{S}^r$ represent $\{S^r_t\}_{t=1}^T$
-%\STATE $\mathbf{S}^0 \leftarrow 0$
-%\STATE $U^0 \leftarrow$ top-$k$ left singular vectors and $V^0 \leftarrow$ top-$k$ right singular vectors from the SVD of $\displaystyle{\sum_{(i,j) \in \Gamma}} \mathbf{x}_i \mathbf{y}_j^\intercal$\\
-%\STATE $\mathcal{L}^0$ : initial loss
-%\REPEAT
-%  \STATE $U^{r+1} \leftarrow \displaystyle{\argmin_{U}} \; \mathcal{L}(U, V^r, \mathbf{S}^r)$  \STATE $V^{r+1} \leftarrow \displaystyle{\argmin_{V}} \; \mathcal{L}(U^{r+1}, V, \mathbf{S}^r) $  \STATE For each task $t$\\
-%  \hspace{0.5cm} $S^{r+1}_t \leftarrow \displaystyle{\argmin_{S_t}} \; \mathcal{L}(U^{r+1}, V^{r+1}, \mathbf{S}^r_{-t}) $
-%  \STATE Compute $\mathcal{L}^{r+1}$ and let $\delta \leftarrow (\mathcal{L}^r - \mathcal{L}^{r+1})/\mathcal{L}^r$
-%\UNTIL {$\delta < \tau$}%\While
-%\end{algorithmic}\end{small}\end{algorithm}
+\begin{algorithm}[!t]   \caption{Alternating Least Squares}
+   \label{alg}
+   %\begin{minipage}{0.9\textwidth}
+   \begin{small}
+\begin{algorithmic}[1]
+\STATE \textbf{Input}: \\
+$k$ : number of latent factors\\
+$\Gamma$: pairs of entities for initialization\\
+For every task $t$,\\
+${\{\mathbf{x}_{ti}\}, \{\mathbf{y}_{tj}\}}$: feature vectors for pathogen and host proteins resp.\\
+$\Omega_t$: the observed entries of the matrix $M_t$\\
+\STATE \textbf{Initialization}: \\
+\STATE An iteration $r$, let $\mathbf{S}^r$ represent $\{S^r_t\}_{t=1}^T$
+\STATE $\mathbf{S}^0 \leftarrow 0$
+\STATE $U^0 \leftarrow$ top-$k$ left singular vectors and $V^0 \leftarrow$ top-$k$ right singular vectors from the SVD of $\displaystyle{\sum_{(i,j) \in \Gamma}} \mathbf{x}_i \mathbf{y}_j^\intercal$\\
+\STATE $\mathcal{L}^0$ : initial loss
+\REPEAT
+  \STATE $U^{r+1} \leftarrow \displaystyle{\argmin_{U}} \; \mathcal{L}(U, V^r, \mathbf{S}^r)$  \STATE $V^{r+1} \leftarrow \displaystyle{\argmin_{V}} \; \mathcal{L}(U^{r+1}, V, \mathbf{S}^r) $  \STATE For each task $t$\\
+  \hspace{0.5cm} $S^{r+1}_t \leftarrow \displaystyle{\argmin_{S_t}} \; \mathcal{L}(U^{r+1}, V^{r+1}, \mathbf{S}^r_{-t}) $
+  \STATE Compute $\mathcal{L}^{r+1}$ and let $\delta \leftarrow (\mathcal{L}^r - \mathcal{L}^{r+1})/\mathcal{L}^r$
+\UNTIL {$\delta < \tau$}%\While
+\end{algorithmic}\end{small}
+%\end{minipage}
+\end{algorithm}
 
@@
 and a block coordinate descent method called alternating least squares (ALS) is commonly used to optimize such functions. 
-To speed up convergence we use an adaptive step size. %The detailed optimization procedure is shown in Algorithm \ref{alg}.
+To speed up convergence we use an adaptive step size. The detailed optimization procedure is shown in Algorithm \ref{alg}.
 
@@
 
-\section{Experimental setup}
-\subsection{Datasets and features}
+\section{Dataset}
 \label{sec:datasets}
 We use three human-virus PPI datasets from the PHISTO \citep{phisto} database (version from 2014), the characteristics of which are summarized in Table \ref{datasets}. 
-The \textit{Influenza A} task includes various strains of flu: H1N1, H3N2, H5N1, H7N3. 
-Similarly, the \textit{Hepatitis} task includes various strains of the virus. 
+The \textit{Influenza A} task includes various strains of flu such as: H1N1 (Puerto Rico, Texas strains etc), H3N2 (England strain etc), H5N1, H7N3.
+Similarly, the \textit{Hepatitis} task includes strains such as of the virus. 
 All three are single-strand RNA viruses, with \textit{Hepatitis} being a positive-strand ssRNA whereas \textit{Influenza} and \textit{Ebola} are negative-strand viruses.
-Note that we use UniprotKB identifiers to refer to
-the proteins; hence if an identical protein appears in two strains it will be treated differently as it has a different
-UniprotKB id. The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (row-4 in Table \ref{datasets}). \\
+The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (row-4 in Table \ref{datasets}).
+\\
+
+\subsection{Addressing the homologs in PHISTO}
+\label{sec:phisto}
+The interactions data reported in PHISTO database is aggregated from several host-pathogen interactions data sources such as Virus-Mint, PHI-base etc.
+We found that many of the interactions (within a task) involve homologous PPI - virus protein homologs from two strains interacting with either the same human protein or with homologous human proteins.
+Such PPI usually come from different publications/ studies, but they introduce a lot of redundancy in the dataset.
+Removing such PPI is challenging as there are several criteria to consider, the most important being: \textit{what threshold of sequence similarity constitutes a homolog}?
+The very existence of sequence similarity between proteins is what makes it possible for biologists to predict unknown interactions and most computational methods rely on it for PPI prediction.
+
+In row-2 of Table \ref{datasets}, we give a conservative estimate of the number of non-homologous PPI.
+For this, we identify homologous PPI and retain only one of them in the dataset.
+Homologs were obtained using BLAST sequence alignment (blastp) using an e-value cut-off threshold of $1e^{-10}$.
+This is a relaxed cut-off since there are cases where two functionally different proteins (either virus-virus or human-human) with low query coverage and low identity end up as `homologs'.
+For example, human proteins Q8WV28 (gene BLNK, B-cell linker protein) and O00459 (gene PIK3R2, Phosphatidylinositol 3-kinase) are considered homologs by this cut-off.
+The number of non-homologous PPI should therefore be considered a lower limit.
+
+In our experiments section, we thus evaluate our models in two settings - (1) on the original dataset (2) by removing homologous PPI.
+
+%and thus interologs (within a task), the presence of which may not give a clear picture of a computational model.
+However these are reported by different publications/studies and were found using different experimental methods, so we do not exclude them and instead report performance on
+
 \noindent\textbf{Features}: 
@@
 Such features have been successfully applied in prior work \citep{dyer07,me_ismb_2013}.
+
+\section{Experimental setup}
 
@@
 We score each pair $(h, v)$ from the bipartite graph as follows: $\max \{ \max $ sequence identity between $h$ and all host proteins in $H(v)$; $\max $ sequence identity between $ v $ and all virus proteins in $V(h) \}$.
+This baseline is shown in the first row as \textit{Homolog} in Table \ref{resultsTable}.
 
@@
 The standard deviation is also shown.
-The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.}
+The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
+The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
+This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
 \label{resultsTable}
@@
 \toprule
-& \multicolumn{3}{c}{\textbf{10\% training}} && \multicolumn{3}{c}{\textbf{30\% training}} \\ \cline{2-4}\cline{6-8}
+& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
 & \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & & & && & & \\
+Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
 STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
@@
 %Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
-MTPL \citep{me_ismb_2013} & \textbf{0.217}$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && \textbf{0.260}$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
+MTPL \citep{me_ismb_2013} & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
 IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
@@
 
-\section{Results}
-As described in Section \ref{}, the PPI data we obtain from PHISTO has homologs within a task.
-We therefore present results on two sets of data: (a) with the original PPIs from PHISTO shown in Table \ref{resultsTable} and (b) after removal of all homologs within each task in Table \ref{homresultsTable}.
-The latter comparison gives us an idea of how much the approaches benefit from the redundancy within each task.
-
-Table \ref{resultsTable} has the AUC-PR for all methods. 
-Note that the AUC-PR of a random classifier model is $\approx 0.01$.
-The first row (STL) is the single-task baseline and all others are multitask models. In general, we notice that multitask learning benefits
-all tasks. The first three columns show the results in the 10\% setting. 
-%The number of training positive examples from each task are 8 for Ebola, 85 for Influenza and 98 for Hepatitis-C. 
-Our model (last row) has significant gains for Influenza (1.4 times better than the next best) and modest improvements for the other tasks.
-The variance in the performance is high for the Ebola task (column 1) owing to the small number of positives 
-in the training splits (8 positives).
-The most benefits for our model are seen in the 30\% setting for all tasks, with 
-improvements of 39, 3 and 12 percentage points on the \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} tasks, respectively. \textit{Ebola}, the data-poorest task, benefits the most.
-
-\begin{table}[t]\caption{Area Under the Precision-Recall curve for each task upon removal of the homologs within each task, on the 10\% setting}
+\begin{table}\caption{Area Under the Precision-Recall curve for each task on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
+Only the most representative baselines are retained for this experiment.
+Our model does significantly better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
 \label{homresultsTable}
@@
 \toprule
+\multicolumn{4}{c}{\textbf{10 \% training, 90\% test on Non-homologous PPI}} \\
 & \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & & & \\
-STL (Ridge Reg.)  & & & \\
-Sparse+low-rank \citep{chen2012} & & &  \\ 
-BSL-MTL (this work) & & &  \\ \bottomrule
+Homolog & 0.207 $\pm$.07 & 0.113 $\pm$.01 & 0.109 $\pm$ .01 \\
+STL  & 0.210 $\pm$.14 & 0.474 $\pm$ .06 & 0.302 $\pm$.03 \\
+Sparse+LR & 0.216 $\pm$.13 & 0.482 $\pm$.05 & 0.312 $\pm$.04 \\ 
+BSL-MTL & 0.295 $\pm$.10 & 0.544 $\pm$.06 & 0.505 $\pm$.05 \\ \bottomrule
 \end{tabular}
@@
 
+\section{Results}
+The PPI data we obtain from PHISTO has homologous PPI within a task.
+Note that the presence of sequence similarity across virus proteins is a strong criterion for computational methods to work, however orthologs or paralogs of proteins within a task lead to redundancy in the data and some readers may find it harder to judge the contribution of the methods.
+We therefore present results on two sets of data:
+\begin{enumerate}[a]
+\item with the original PPIs from PHISTO shown in Table \ref{resultsTable}
+\item after removal of all homologous PPI.
+Here we ensure that the test data does not contain any homologs of PPI observed in the training data.
+A BLAST sequence similarity cut-off of $1e^{-10}$ was used to find homologs (see Section \ref{sec:phisto} for details).
+Note that the number of negatives in the test data is still 100 times the number of positives.
+Table \ref{homresultsTable} shows these results.
+\end{enumerate}
+
+Table \ref{resultsTable} has the AUC-PR for all methods. 
+Note that the AUC-PR of a random classifier model is $\approx 0.01$.
+The first row is the Homolog heuristic based approach, STL is the single-task baseline and all others are multitask models. In general, we notice that multitask learning benefits
+all tasks. The first three columns show the results in the 10\% setting. 
+%The number of training positive examples from each task are 8 for Ebola, 85 for Influenza and 98 for Hepatitis-C. 
+Our model (last row) has significant gains for Influenza (1.4 times better than the next best) and modest improvements for the other tasks.
+The variance in the AUC is high for the Ebola task (column 1) owing to the small number of positives 
+in the training splits (8 positives).
+The most benefits for our model are seen in the 30\% setting for all tasks.
+Notably, Homolog performs comparably to our model.
+In addition to the training data seen by all other methods, this baseline also uses additional data in the form of the virus-host interactome from several other viruses.
+Results on the no-homolog part of our dataset sheds further light on this.
+
+In Table \ref{homresultsTable} we show AUC for all the representative baselines in the 10\% setting.
+Our method outperforms all others by large margins on the non-homologous portion of the test data, by intelligently sharing information with other tasks.
+There are thus two components to the prediction performance - (1) the presence of sequence-similar proteins within a task (2) non-linear similarities between the interactions within a task and across tasks.
+We are able to capture the latter well via the shared subspace structure of our model.
+
+%Homolog & 0.010 $\pm$.03 & 0.079 $\pm$.01 & 0.135 $\pm$.02 \\  
+%STL (Ridge Reg.)  & 0.054 $\pm$.03 &  0.400 $\pm$.06 & 0.192 $\pm$.03 \\
+%BSL-MTL (this work) & 0.071 $\pm$.03 & \textbf{0.501} $\pm$.05 & \textbf{0.331} $\pm$.03  \\ \bottomrule
+
 \subsection{Disadvantages of the Homolog method}
-In Figures \ref{fig:ebolanet}-\ref{fig:flunet} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
+In Figure \ref{fig:network} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
 The interactions mainly involve human proteins which do not have a homolog in the training data.
 This illustrates that our features capture more information than mere sequence similarity.
-Further, inclusion of interactions from other viruses in the naive manner as done by the homologs model is not helpful - our model is able to learn what to borrow from the other tasks accurately.
+Further, inclusion of interactions from other viruses in the na\"ive manner as done by the homologs in fact hurts the performance (results in too many false positives) - our model is able to learn what to borrow from the other tasks accurately.
 
-\begin{figure*}
-\ffigbox[\textwidth]
-{
-\begin{subfloatrow}[3]
-\ffigbox[0.22\textwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{1cm}
-}
-{
-\caption{Ebola-human PPI}
-\label{fig:ebolanet}
-}
-\ffigbox[0.33\textwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.33, trim = 0.2 0.2 0.2 0.2]{flavinet.pdf}
-}
-{
-\caption{Hepatitis-human PPI}
-\label{fig:flavinet}
-}
-\ffigbox[0.34\textwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.35, trim = 0.2 0.2 0.2 0.2]{flunet.pdf}
-}
-{
-\caption{Influenza-human PPI}
-\label{fig:flunet}
-}
-\end{subfloatrow}
-}
-{
-\caption{PPI from each task that are found by our method but are missed by the homolog method}
-}
-\end{figure*}
+\begin{table}\caption{Area Under the Precision-Recall curve for each task on the \textbf{homologous PPI} within each task, in the 10\% setting.
+The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks.
+The Homolog method has access to a wider virus-human PPI graph involving several other viruses and is able to infer homologous PPI for \textit{Ebola} better (note that some of the homologous PPI are not redundant examples but involve proteins with very small regions of sequence similarity.}
+\label{homresultsTable}
+\begin{small}
+\begin{center}
+\def\arraystretch{1.2}
+\begin{tabular}{c|ccc}
+\toprule
+\multicolumn{4}{c}{\textbf{10 \% training, test on Homologous PPI }} \\
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & 0.399 $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
+STL   & 0.192 $\pm$.15 & 0.504 $\pm$.10 & 0.305 $\pm$.03 \\
+Sparse+LR & 0.184 $\pm$.13 & 0.512 $\pm$.10 & 0.315 $\pm$.04 \\ 
+BSL-MTL & 0.130 $\pm$.10 & 0.815 $\pm$.02 & 0.475 $\pm$.02 \\ \bottomrule
+\end{tabular}
+\end{center}
+\end{small}
+\end{table}
 
@@
 Firstly, the projected data has a much better separation than the original data.
-Secondly, Fig \ref{fig:pca} (right) tells us that Hepatitis-C and Influenza have many proteins with similar binding tendencies, and that these behave differently than most Ebola virus proteins.
-This observation is not obvious in the PCA of the original feature space (Fig \ref{fig:pca} left), where proteins with similar sequences cluster together.
-These clusters of proteins can be analyzed further for enrichment of Gene Ontology (GO) annotations.
+Secondly, Fig \ref{fig:pca2} tells us that Hepatitis-C and Influenza have many proteins with similar binding tendencies, and that these behave differently than most Ebola virus proteins.
+This observation is not obvious in the PCA of the original feature space (Fig \ref{fig:pca1}), where proteins with similar sequences group together.
+We analyze the projected data further by looking at the clusters of proteins for enrichment of Gene Ontology (GO) annotations (proteins were first clustered in this lower dimensional space using k-means and setting the number of clusters to 6).
+Of particular interest are the highlighted three clusters which contain either proteins projected far from others (such as cluster-1) and proteins from different viruses projected close together (cluster-2 and cluster-3).
+For enrichment analysis, we use the FuncAssociate 3.0 \citep{funcAsso} tool and GO annotations for the three viruses from UniProtKB.
 
 \begin{figure}
-\begin{floatrow}
-\ffigbox[0.49\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.35, trim = 0 0.8cm 0 0]{fig2.eps}
-}{%
-\caption{Principal component analysis (PCA) of virus proteins in the original feature space}%
+\begin{tabular}{c}
+%\begin{minipage}{0.3\textwidth}
+%\begin{center}
+%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%\end{center}
+%\end{minipage}
+%&
+\begin{minipage}{\textwidth}
+\begin{center}
+\includegraphics[scale=0.33, trim = 0.2 0.2 0.2 0.2]{flavinet.pdf}
+\end{center}
+\end{minipage}
+\\
+\begin{minipage}{\textwidth}
+\begin{center}
+\includegraphics[scale=0.35, trim = 0.2 0.2 0.2 0.2]{flunet.pdf}
+\end{center}
+\end{minipage}
+\end{tabular}
+\caption{PPI from each task that are found by our method but are missed by the homolog method.
+To obtain these, we chose a random train/test split and picked examples from the test data on which the homolog method fails - these are ones for which sequence-similar proteins are not found.
+(a) shows \textit{Ebola}-human PPI, (b) \textit{Hepatitis-C} human PPI and (c) \textit{Influenza}-human PPI.}
+\label{fig:network}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[scale=0.4, trim = 0 0.2cm 0 0.5cm]{fig2.eps}
+\caption{Principal component analysis (PCA) of virus proteins in the original feature space.
+The first two principal components are shown.
+Shape and color of the points indicates which virus that protein comes from.}%
 \label{fig:pca1}
-}
-\ffigbox[0.49\columnwidth]{%
-\includegraphics[scale=0.33, trim = 0 0 0 0]{fig3.eps}
-}{%
+\end{figure}
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.27, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
 \caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
 The first two principal components are shown.
-Shape of the points indicates which virus that protein comes from.}%
+Shape of the points indicates which virus that protein comes from.
+We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
 \label{fig:pca2}
-}
-\end{floatrow}
 \end{figure}
 
-\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
-We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \citep{iedb}.
-The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:hepc_motifs} is found in 53 epitopes.
-From the \textit{Ebola} motifs in Fig \ref{fig:ebola_motifs}, we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes.
-\texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
-Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
-is a further validation of its performance.
-Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
-features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
-Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.
-
-\noindent\emph{Phosphorylation sites}: 
-We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
-related to the amino acids being frequent targets of phosphorylation.
-Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \citep{phosphosite} are
- repositories for known sites in human proteins. Since these are sites in human proteins, we searched for the patterns
-from the 4-mer motif in Fig \ref{fig:human_motifs} and found several to be flanking known phosphorylation sites in human proteins:
-\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site).
-This observation also supports the notion that the motifs predictive
-of interaction are biologically significant.
+\begin{table}[bht]
+\label{goterms}
+\begin{scriptsize}
+\begin{center}
+\def\arraystretch{1.1}
+\begin{tabular}{|c|c|}
+\multicolumn{2}{c}{Gene Ontology terms for clusters from Figure \ref{fig:pca2}} \\ \hline
+cluster 1 & mRNA (guanine-N7-)-methyltransferase  \\
+(ebola only)& activity; RNA-directed RNA   \\
+& polymerase activity; ATP binding \\ \hline
+cluster 2 & host cell endoplasmic reticulum membrane; \\ 
+(hepatitis \& & host cell lipid particle; induction by virus \\
+influenza & of host autophagy; host cell mitochondrial membrane; \\
+proteins) & suppression by virus of host STAT1 activity \\ \hline
+cluster 3 & fusion of virus membrane with \\
+(all) & host plasma membrane \\ \hline
+\end{tabular}
+\caption{GO terms (for process, function, component) enriched in the highlighted clusters from Figure \ref{fig:pca2}.
+FuncAssociate GO enrichment tool was used and terms with p-value < 0.001 are shown.
+The first column also indicates the composition of each cluster.}
+\end{center}
+\end{scriptsize}
+\end{table}
 
@@
 \subsubsection{Sequence motifs from virus proteins}
-In Figures \ref{fig:virus_motifs}-\ref{fig:flu_motifs}, we show sequence motifs derived from the top $k$-mers that contribute to interactions.
-The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to
+We analyze sequence motifs derived from the top $k$-mers that contribute to interactions. The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to 
 compute these motifs.
@@
 analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
+We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by 
+hydrophilic amino acids (T, K, R, D, E).
 
@@
 the matrix $U V^T$.
-These motifs are shown in Figure \ref{fig:human_motifs}.
-The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepC}$ and $S_{flu}$ respectively.
-Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
-These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and
-\ref{fig:flu_motifs}.
-
-We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by 
-hydrophilic amino acids (T, K, R, D, E). All other motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G)
-though S and T do appear in some motifs as well.
+These motifs are shown in the supplementary.
+The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepc}$ and $S_{flu}$ respectively.
+The motif for Hepatitis is shown in Figure \ref{fig:virus_motifs}(right); the rest can be found in the supplementary.
+%Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
+These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and \ref{fig:flu_motifs}.
+The virus-specific motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G) though S and T do appear in some motifs as well.
 %The human protein motifs are shown in \ref{fig:human_motifs}; the 3-mer motif show less of a pattern, in particular the first residue does not seem to have any dominant amino-acids.
 %However the 4-mer seems to be dominated by hydrophobic amino-acids L, P, I, F in the first three positions and hydrophilic residues S, T, Y in the fourth.
-The human protein motifs are shown in Fig \ref{fig:human_motifs}.
+%The human protein motifs are shown in Fig \ref{fig:human_motifs}.
 In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
 
-The motifs for tri-mers are shown in the supplementary.
+%The motifs for tri-mers are shown in the supplementary.
+
+\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
+We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \citep{iedb}.
+The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:virus_motifs}(right) is found in 53 epitopes.
+From the \textit{Ebola} motif (supplementary), we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes.
+\texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
+Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
+is a further validation of its performance.
+Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
+features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
+Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.
+
+\noindent\emph{Phosphorylation sites}: 
+We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
+related to the amino acids being frequent targets of phosphorylation.
+Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \citep{phosphosite} are
+ repositories for known sites in human proteins.
+Since these are sites in human proteins, we searched for the patterns from the 4-mer motif of human proteins (supplementary) and found several to be flanking known phosphorylation sites in human proteins:
+\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site).
+This observation also supports the notion that the motifs predictive
+of interaction are biologically significant.
 
@@
 
-\begin{figure}
-\begin{floatrow}
-\ffigbox[0.62\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4.eps}
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig5.eps}
-}{%
-\caption{Top tri-mer and four-mer sequence motifs from virus proteins across all three viruses. These correspond to features important for interactions.}%
-\label{fig:virus_motifs}
-}
-\ffigbox[0.31\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig6.eps}
-}{%
-\caption{Sequence motif constructed from the top four-mer features of human proteins}%
-\label{fig:human_motifs}
-}
-\end{floatrow}
-\begin{floatrow}
-\ffigbox[0.3\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0.2cm 0 0 0]{fig7.eps}
-}{%
-\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
-\label{fig:ebola_motifs}
-}
+%}
+%\ffigbox[0.31\columnwidth]{%
+%%trim option's parameter order: left bottom right top
+%\includegraphics[scale=0.165, trim = 0 0 0 0]{fig6.eps}
+%}{%
+%\caption{Sequence motif constructed from the top four-mer features of human proteins}%
+%\label{fig:human_motifs}
+%}
+%\end{floatrow}
+%\begin{floatrow}
+%\ffigbox[0.3\columnwidth]{%
+%%trim option's parameter order: left bottom right top
+%\includegraphics[scale=0.165, trim = 0.2cm 0 0 0]{fig7.eps}
+%}{%
+%\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
+%\label{fig:ebola_motifs}
+%}
+%
+%\ffigbox[0.01\columnwidth]{}
+%
+%\ffigbox[0.3\columnwidth]{%
+%%trim option's parameter order: left bottom right top
+%\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
+%}{%
+%\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
+%\label{fig:hepc_motifs}
+%}
+%
+%\ffigbox[0.3\columnwidth]{%
+%%trim option's parameter order: left bottom right top
+%\includegraphics[scale=0.165, trim = 0 0 0 0]{fig9.eps}
+%}{%
+%\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
+%\label{fig:flu_motifs}
+%}
+%\end{floatrow}
+%\end{figure*}
 
-\ffigbox[0.01\columnwidth]{}
+\section{Results on HIV data}
+We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
+This experiment tests how our multitask model behaves in the presence of a relatively outlier task.
+The interactions data is from XXXX database and has XXXX human-HIV PPI.
+In Table \ref{hivresults} we show the AUC-PR for our method and the representative baselines.
 
-\ffigbox[0.3\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
-}{%
-\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
-\label{fig:hepc_motifs}
-}
-
-\ffigbox[0.3\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig9.eps}
-}{%
-\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
-\label{fig:flu_motifs}
-}
-\end{floatrow}
-\end{figure}
+\begin{table}[h]\caption{Area Under the Precision-Recall curve for each task upon addition of a new task - HIV, in the 10\% setting.
+Note that the performance of STL and the homolog method on the previous three tasks is not influenced by the addition of a new task.}
+\label{hivresults}
+\begin{small}
+\begin{center}
+\def\arraystretch{1.2}
+\begin{tabular}{c|cccc}
+\toprule
+& \multicolumn{4}{c}{10 \% training } \\
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
+Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
+STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
+Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
+BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
+\end{tabular}
+\end{center}
+\end{small}
+\end{table}
 
@@
 
+\begin{figure}
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4.eps}
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
+\caption{(\textit{Left}) Top tri-mer sequence motifs from virus proteins across all three viruses.
+These correspond to features important for interactions and is derived from our model parameter $H = U V^{T}$.
+(\textit{Right}) Motif corresponding to enriched 4-mer patterns found by our model (parameter: $S_{hepc}$) for the Hepatitis-C virus task.}%
+\label{fig:virus_motifs}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-%\bibliographystyle{plainnat}
 \bibliographystyle{natbib}
-%\bibliographystyle{agsm}
 \begin{scriptsize}

### 2016-08-31 paper.tex

@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1.eps}
+\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
 }
@@
 \centering
-\includegraphics[scale=0.4, trim = 0 0.2cm 0 0.5cm]{fig2.eps}
+\includegraphics[scale=0.4, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
 \caption{Principal component analysis (PCA) of virus proteins in the original feature space.
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4.eps}
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
+\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4-eps-converted-to.pdf}
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8-eps-converted-to.pdf}
 \caption{(\textit{Left}) Top tri-mer sequence motifs from virus proteins across all three viruses.

### 2016-08-31 paper.tex
new pdf

@@
 
-%\access{Advance Access Publication Date: Day Month Year}
-%\appnotes{Manuscript Category}
+\access{Advance Access Publication Date: Day Month Year}
+\appnotes{Manuscript Category}
 
@@
 
-%\subtitle{Subject Section}
+\subtitle{Subject Section}
 
@@
 
-%\corresp{$^\ast$To whom correspondence should be addressed.}
+\corresp{$^\ast$To whom correspondence should be addressed.}
 
@@
 
-%\abstract{\textbf{Motivation:}
+\abstract{\textbf{Motivation:}
+Disease causing pathogens such as viruses, introduce their proteins into the host cells 
+where they interact with the host's proteins enabling the virus to replicate inside the host.
+These interactions between pathogen and host proteins are key to understanding infectious diseases.
+Often multiple diseases involve phylogenetically related or biologically similar pathogens.
+Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}.
+\\
+\textbf{Results:}
+Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
+We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
+We show how our model's parameters can be interpreted to reveal both general and specific
+interaction-relevant characteristics of the viruses. \\
+\textbf{Availability:}
+Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl\_mtl.tgz}\\
+\textbf{Contact:}
+\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
+}
+
+\maketitle
+
+%\begin{abstract}
+%\section{Motivation}
 %Disease causing pathogens such as viruses, introduce their proteins into the host cells 
@@
 Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}.
-\\
-%\textbf{Results:}
+%\section{Results}
 %Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
@@
 %interaction-relevant characteristics of the viruses. \\
-%\textbf{Availability:}
+%\section{Availability}
 %Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
-%\textbf{Contact:}
+%\section{Contact}
 %\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
-%}
-
-\maketitle
-
-\begin{abstract}
-\section{Motivation}
-Disease causing pathogens such as viruses, introduce their proteins into the host cells 
-where they interact with the host's proteins enabling the virus to replicate inside the host.
-These interactions between pathogen and host proteins are key to understanding infectious diseases.
-Often multiple diseases involve phylogenetically related or biologically similar pathogens.
-Here we present a multitask learning method to jointly model interactions between human proteins and three different, but related viruses: \textit{Hepatitis C}, \textit{Ebola} virus and \textit{Influenza A}.
-\section{Results}
-Our multitask matrix completion based model uses a shared low-rank structure in addition to a task-specific sparse structure to incorporate the various interactions.
-We obtain between 7 to 39 percentage points improvement in predictive performance over prior state-of-the-art models. 
-We show how our model's parameters can be interpreted to reveal both general and specific
-interaction-relevant characteristics of the viruses. \\
-\section{Availability}
-Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
-\section{Contact}
-\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
-\end{abstract}
+%\end{abstract}
 

### 2016-09-01 paper.tex
citations and results

@@
 %The matrix factorization based methods have been more popular than the other methods. 
-There have also been multitask approaches on link prediction \citep{zhang12,cao_icml2010,li09,cmf}. 
+There have also been multitask approaches on link prediction \citep{zhang12,cao_icml2010,li09,cmf} however many of those models do not incorporate features of the nodes in the graph. 
 %\cite{li11} presents a survey on multitask/transfer methods in collaborative filtering. 
@@
 %has been proposed for single-graph link prediction problems. 
-\cite{menon_ecml11} propose a single-graph model that combines linear 
-and bilinear features, latent parameters on the nodes and several other parameters into a function that minimizes a ranking loss. 
+\cite{menon_ecml11} combine linear 
+and bilinear features, latent parameters on the nodes and several other parameters into a function that minimizes a ranking loss for link prediction in a single graph. 
 %Probabilistic models to combine network structure with node features have also been proposed. 
@@
 Chen et al \cite{chen2012} have developed efficient algorithms for sparse and low-rank multitask learning models for classification.
+Another related line of work is co-embedding, where the goal is to embed different types of objects in the same shared lower-dimensional space \citep{mirzazadeh:2015}.
 
@@
 For this, we identify homologous PPI and retain only one of them in the dataset.
-Homologs were obtained using BLAST sequence alignment (blastp) using an e-value cut-off threshold of $1e^{-10}$.
+Homologs were obtained using BLAST sequence alignment (blastp) using an e-value cut-off threshold of $1e^{-5}$.
 This is a relaxed cut-off since there are cases where two functionally different proteins (either virus-virus or human-human) with low query coverage and low identity end up as `homologs'.
@@
 \subsection{Comparison with a homolog baseline}
-In addition to the machine learning methods presented in Section \ref{mlmethods}, we also tried a simple homology based approach that is based on protein sequence identity.
+In addition to the machine learning methods presented in Section \ref{mlmethods}, we also tried a simple homology based approach.
+As per this method, a virus protein $v$ will interact with a human protein $h$ if it interacts with any other human protein $h'$ similar to $h$.
+The converse holds too.
+Such homology-motivated heuristic approaches have been traditionally popular in the inter-organism PPI prediction literature \citep{vidal:2001,matthews:2001}.
+\cite{mika:2006} offers a commentary on the the success of such approaches in prediction PPI within an organism and the transferability across organisms.
+Modifications of such approaches have been published recently \citep{chen:2009,lin:2013,murakami:2014}.
+
+The virus-virus and human-human protein similarity graphs used in this baseline were computed using protein sequence similarity.
+We ran BLAST sequence alignment (\texttt{blastp}) with an e-value cut-off of 0.1 and the negative logarithm of the e-value was used as a measure of similarity (we use a somewhat relaxed cut-off as we are comparing across organisms).
 We use the entire human-virus interactome from PHISTO as input to this method.
 Given a human protein $h$, we rank all virus proteins $v$ based on the following simple scheme.
-let $V(h)$ be a set of all virus proteins interacting with the human protein $h$ and let $H(v)$ be all human proteins interacting with the virus protein $v$.
-We score each pair $(h, v)$ from the bipartite graph as follows: $\max \{ \max $ sequence identity between $h$ and all host proteins in $H(v)$; $\max $ sequence identity between $ v $ and all virus proteins in $V(h) \}$.
+Let $V(h)$ be a set of all virus proteins interacting with the human protein $h$ and let $H(v)$ be all human proteins interacting with the virus protein $v$.
+We score each pair $(h, v)$ from the bipartite graph as follows: $\max \{ \max $ similarity between $h$ and all human proteins in $H(v)$; $\max $ similarity between $v$ and all virus proteins in $V(h) \}$.
 This baseline is shown in the first row as \textit{Homolog} in Table \ref{resultsTable}.
@@
 
-\begin{table*}[t]\caption{Area Under the Precision-Recall curve for each task on the PHISTO dataset without removing the homologs.
+\begin{table*}[h]\caption{Area Under the Precision-Recall curve for each task on the PHISTO dataset without removing the homologs.
 The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
@@
 
+\section{Results}
+The PPI data we obtain from PHISTO has homologous PPI within a task.
+Note that the presence of sequence similarity across virus proteins is a strong criterion for computational methods to work, however orthologs or paralogs of proteins within a task lead to redundancy in the data and some readers may find it harder to judge the contribution of the methods.
+We therefore present results on three sets:
+\begin{enumerate}[a]
+\item the original PPIs from PHISTO. AUC-PR is shown in Table \ref{resultsTable}
+\item \textbf{Non-homologous partition}: after removal of all homologous PPI from the test data.
+Here we ensure that the test data does not contain homologs of PPI observed in the training data (this ensures there are no redundant examples).
+Further, every set of homologous PPIs in the test data is replaced with a single PPI.
+A BLAST sequence similarity cut-off of $1e^{-5}$ was used to find homologs.
+The results are in Table \ref{nohomresultsTable}.
+\item \textbf{Homologous PPI partition} : only homologous PPI part of the test data. This is essentially the set of PPI removed in the set-(b) above. Table \ref{homresultsTable} shows these results. 
+\end{enumerate}
+Note that the partitions in (b) and (c) together comprise our test data.
+Also, the negatives in the test data are randomly split into the `no-homolog' and 'homolog' partitions while maintaining the number of negatives in each partition to be 100 times the number of positives.
+
+\noindent\textit{Overall performance}
+For reference, the AUC-PR of a random classifier model is $\approx 0.01$ due to the class imbalanced nature of the data (class skew = 1:100).
+The first row is the Homolog heuristic based approach, STL is the single-task baseline and all others are multitask models. In general, we notice that multitask learning benefits
+all tasks. The first three columns show the results in the 10\% setting. 
+%The number of training positive examples from each task are 8 for Ebola, 85 for Influenza and 98 for Hepatitis-C. 
+Our model (last row) has large gains for Influenza (1.4 times better than the next best) and modest improvements for the other tasks.
+The variance in the AUC is high for the Ebola task (column 1) owing to the small number of positives 
+in the training splits (8 positives).
+The most benefits for our model are seen in the 30\% setting for all tasks.
+Notably, Homolog performs comparably to our model.
+In addition to the training data seen by all other methods, this baseline also uses additional data in the form of the virus-virus, human-human sequence similarities and a virus-human interactome from several other viruses (besides the three tasks).
+Results on the homolog part of our dataset sheds further light on this.
+
+\noindent\textit{Performance on the non-homologous PPI partition}:
+In Table \ref{nohomresultsTable} we show AUC for all the representative baselines in the 10\% setting. 
+Overall, our method performs best on this partition over all tasks.
+Comparison with the AUC in Table \ref{resultsTable} shows that AUC for the \textit{Ebola} task is higher on all baselines except Homolog, indicating that
+a bigger fraction of the overall performance comes from the non-homologous partition (this is also marginally the case for \textit{Influenza}). 
+This effect is the most pronounced for our method (BSL-MTL)
+and demonstrates that we are able to predict such PPI by intelligently sharing information with other tasks. 
+There are thus two components to the prediction performance - (1) the presence of sequence-similar proteins within a task (2) non-linear similarities between the interactions within a task and across tasks.
+We are able to capture the latter well via the shared subspace structure of our model.
+
+%Homolog & 0.010 $\pm$.03 & 0.079 $\pm$.01 & 0.135 $\pm$.02 \\  
+%STL (Ridge Reg.)  & 0.054 $\pm$.03 &  0.400 $\pm$.06 & 0.192 $\pm$.03 \\
+%BSL-MTL (this work) & 0.071 $\pm$.03 & \textbf{0.501} $\pm$.05 & \textbf{0.331} $\pm$.03  \\ \bottomrule
+
 \begin{table}\caption{Area Under the Precision-Recall curve for each task on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
+To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}.
 Only the most representative baselines are retained for this experiment.
-Our model does significantly better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
-\label{homresultsTable}
+Our model performs better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
+\label{nohomresultsTable}
 \begin{small}
@@
 Sparse+LR & 0.216 $\pm$.13 & 0.482 $\pm$.05 & 0.312 $\pm$.04 \\ 
-BSL-MTL & 0.295 $\pm$.10 & 0.544 $\pm$.06 & 0.505 $\pm$.05 \\ \bottomrule
+BSL-MTL & \textbf{0.295} $\pm$.10 & \textbf{0.544} $\pm$.06 & \textbf{0.505} $\pm$.05 \\ \bottomrule
 \end{tabular}
@@
 
-\section{Results}
-The PPI data we obtain from PHISTO has homologous PPI within a task.
-Note that the presence of sequence similarity across virus proteins is a strong criterion for computational methods to work, however orthologs or paralogs of proteins within a task lead to redundancy in the data and some readers may find it harder to judge the contribution of the methods.
-We therefore present results on two sets of data:
-\begin{enumerate}[a]
-\item with the original PPIs from PHISTO shown in Table \ref{resultsTable}
-\item after removal of all homologous PPI.
-Here we ensure that the test data does not contain any homologs of PPI observed in the training data.
-A BLAST sequence similarity cut-off of $1e^{-10}$ was used to find homologs (see Section \ref{sec:phisto} for details).
-Note that the number of negatives in the test data is still 100 times the number of positives.
-Table \ref{homresultsTable} shows these results.
-\end{enumerate}
-
-Table \ref{resultsTable} has the AUC-PR for all methods. 
-Note that the AUC-PR of a random classifier model is $\approx 0.01$.
-The first row is the Homolog heuristic based approach, STL is the single-task baseline and all others are multitask models. In general, we notice that multitask learning benefits
-all tasks. The first three columns show the results in the 10\% setting. 
-%The number of training positive examples from each task are 8 for Ebola, 85 for Influenza and 98 for Hepatitis-C. 
-Our model (last row) has significant gains for Influenza (1.4 times better than the next best) and modest improvements for the other tasks.
-The variance in the AUC is high for the Ebola task (column 1) owing to the small number of positives 
-in the training splits (8 positives).
-The most benefits for our model are seen in the 30\% setting for all tasks.
-Notably, Homolog performs comparably to our model.
-In addition to the training data seen by all other methods, this baseline also uses additional data in the form of the virus-host interactome from several other viruses.
-Results on the no-homolog part of our dataset sheds further light on this.
-
-In Table \ref{homresultsTable} we show AUC for all the representative baselines in the 10\% setting.
-Our method outperforms all others by large margins on the non-homologous portion of the test data, by intelligently sharing information with other tasks.
-There are thus two components to the prediction performance - (1) the presence of sequence-similar proteins within a task (2) non-linear similarities between the interactions within a task and across tasks.
-We are able to capture the latter well via the shared subspace structure of our model.
-
-%Homolog & 0.010 $\pm$.03 & 0.079 $\pm$.01 & 0.135 $\pm$.02 \\  
-%STL (Ridge Reg.)  & 0.054 $\pm$.03 &  0.400 $\pm$.06 & 0.192 $\pm$.03 \\
-%BSL-MTL (this work) & 0.071 $\pm$.03 & \textbf{0.501} $\pm$.05 & \textbf{0.331} $\pm$.03  \\ \bottomrule
+\noindent\textit{Performance on the homologous PPI partition}:
+Except the \textit{Ebola} task, BSL-MTL has the highest AUC.
+The Homolog baseline best captures the homologous PPI in \textit{Ebola} (with an AUC twice the next best).
+However, since there are far fewer homologous PPI in Ebola than in the other tasks (row-2 of Table \ref{datasets}), this contributes less to the overall performance of the Homolog baseline.
+\textit{Hepatitis-C} has the highest fraction of homologous PPI and BSL-MTL does significantly better on these than the other baselines which also leads to a higher overall AUC.
+The same can be said of the \textit{Influenza} task.
 
@@
 \toprule
-\multicolumn{4}{c}{\textbf{10 \% training, test on Homologous PPI }} \\
+\multicolumn{4}{c}{\textbf{10 \% training, test only Homologous PPI }} \\
 & \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
@@
 (i.e computing $X U^\intercal$ and $Y V^\intercal$ to get projections of the virus and human proteins respectively).  
-The principal component analysis (PCA) of this lower dimensional representation is compared with PCA in the original feature space (protein sequence features) in Fig \ref{fig:pca}. 
+The principal component analysis (PCA) of this lower dimensional representation is compared with PCA in the original feature space (protein sequence features) in Figures \ref{fig:pca1} and \ref{fig:pca2}.
 Firstly, the projected data has a much better separation than the original data.
@@
 This observation also supports the notion that the motifs predictive
-of interaction are biologically significant.
+of interaction are biologically interesting.
 
@@
 %and showed significant increases in prediction performance. 
-The method was evaluated in the host-pathogen protein interaction domain for three pathogens (three tasks) and exhibited significant increases in prediction accuracy.
+The method was evaluated in the host-pathogen protein interaction domain for three pathogens (three tasks) and exhibited large increases in prediction accuracy.
 The model parameters provide several avenues for

### 2016-09-01 paper.tex
citations and results

@@
 \textbf{Availability:}
-Our code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl\_mtl.tgz}\\
+Code and data is available at: \url{http://www.cs.cmu.edu/~mkshirsa/bsl_mtl.tgz}\\
 \textbf{Contact:}
@@
 The \textit{Influenza A} task includes various strains of flu such as: H1N1 (Puerto Rico, Texas strains etc), H3N2 (England strain etc), H5N1, H7N3.
-Similarly, the \textit{Hepatitis} task includes strains such as of the virus. 
+The \textit{Hepatitis} task has several subtypes (1a, 1b), isolates such as H77, HC-J6 etc and \textit{Ebola} has the Zaire ebolavirus strain Mayinga and strain 1995.
 All three are single-strand RNA viruses, with \textit{Hepatitis} being a positive-strand ssRNA whereas \textit{Influenza} and \textit{Ebola} are negative-strand viruses.
-The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (row-4 in Table \ref{datasets}).
+The density of the known interactions is quite small when considering the entire proteome (i.e all known proteins) of the host and pathogen species (last row in Table \ref{datasets}).
 \\
@@
 \label{sec:phisto}
-The interactions data reported in PHISTO database is aggregated from several host-pathogen interactions data sources such as Virus-Mint, PHI-base etc.
+The interactions data reported in PHISTO database is aggregated from several host-pathogen interactions data sources such as MINT, IntAct, DIP etc.
 We found that many of the interactions (within a task) involve homologous PPI - virus protein homologs from two strains interacting with either the same human protein or with homologous human proteins.
@@
 
-In our experiments section, we thus evaluate our models in two settings - (1) on the original dataset (2) by removing homologous PPI.
+In our experiments section, we thus evaluate our models in two settings - (1) on the original dataset (2) by creating partitions: the first without any homologous PPI and the other with only the homologous PPI.
 

### 2016-09-01 paper.tex
citations and results
minor

@@
 \begin{array}{ll}
-\mathcal{L}(U,V,\{S_t\}) & = \displaystyle{ \frac{1}{N} \sum_{t=1}^T \sum_{(i,j) \in \Omega_t}} c^t_{ij} \; \ell \big( M_{t_{ij}}, \mathbf{x}_{ti}^\intercal H_t \mathbf{y}_{tj} \big)^2 + \\
+\mathcal{L}(U,V,\{S_t\}) & = \displaystyle{ \frac{1}{N} \sum_{t=1}^T \sum_{(i,j) \in \Omega_t}} c^t_{ij} \; \ell \big( M_{t_{ij}}, \mathbf{x}_{ti}^\intercal H_t \mathbf{y}_{tj} \big) + \\
 & \lambda ( \|U\|^2_F + \|V\|^2_F ) + \displaystyle{\sum_{t=1}^T \sigma_t \|S_t\|_1} 
@@
 
-\noindent\textbf{Features}: 
+\subsection{Features}
 Since the sequence of a protein determines its structure and consequently its function, it may be possible to predict PPIs using the amino acid sequence of a protein pair.
@@
 \begin{center}
-\includegraphics[scale=0.33, trim = 0.2 0.2 0.2 0.2]{flavinet.pdf}
+\includegraphics[scale=0.28, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flavinet.pdf}
 \end{center}
@@
 \multicolumn{2}{c}{Gene Ontology terms for clusters from Figure \ref{fig:pca2}} \\ \hline
-cluster 1 & mRNA (guanine-N7-)-methyltransferase  \\
-(ebola only)& activity; RNA-directed RNA   \\
-& polymerase activity; ATP binding \\ \hline
+cluster 1 & mRNA (guanine-N7-)-methyltransferase activity; RNA-directed \\
+(ebola only)& RNA polymerase activity; ATP binding \\ \hline
 cluster 2 & host cell endoplasmic reticulum membrane; \\ 
@@
 proteins) & suppression by virus of host STAT1 activity \\ \hline
-cluster 3 & fusion of virus membrane with \\
-(all) & host plasma membrane \\ \hline
+cluster 3 (all) & fusion of virus membrane with host plasma membrane \\ \hline
 \end{tabular}
@@
 FuncAssociate GO enrichment tool was used and terms with p-value < 0.001 are shown.
-The first column also indicates the composition of each cluster.}
+The first column also indicates the composition of each cluster (i.e which task's proteins appear in the cluster).}
 \end{center}
@@
 This experiment tests how our multitask model behaves in the presence of a relatively outlier task.
-The interactions data is from XXXX database and has XXXX human-HIV PPI.
-In Table \ref{hivresults} we show the AUC-PR for our method and the representative baselines.
+The interactions data has $\approx$4000 human-HIV PPI across several strains of which 1320 are non-homologous PPI.
+This experiment particularly tests the robustness of all the multitask learning approaches by the addition of an unrelated task.
+Table \ref{hivresults} shows the results.
 
 \begin{table}[h]\caption{Area Under the Precision-Recall curve for each task upon addition of a new task - HIV, in the 10\% setting.
-Note that the performance of STL and the homolog method on the previous three tasks is not influenced by the addition of a new task.}
+Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table \ref{resultsTable}.
+Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
 \label{hivresults}

### 2016-09-01 paper.tex
new pdf

@@
 
-\subsection{Comparison with a homolog baseline}
-In addition to the machine learning methods presented in Section \ref{mlmethods}, we also tried a simple homology based approach.
-As per this method, a virus protein $v$ will interact with a human protein $h$ if it interacts with any other human protein $h'$ similar to $h$.
-The converse holds too.
-Such homology-motivated heuristic approaches have been traditionally popular in the inter-organism PPI prediction literature \citep{vidal:2001,matthews:2001}.
-\cite{mika:2006} offers a commentary on the the success of such approaches in prediction PPI within an organism and the transferability across organisms.
-Modifications of such approaches have been published recently \citep{chen:2009,lin:2013,murakami:2014}.
-
-The virus-virus and human-human protein similarity graphs used in this baseline were computed using protein sequence similarity.
-We ran BLAST sequence alignment (\texttt{blastp}) with an e-value cut-off of 0.1 and the negative logarithm of the e-value was used as a measure of similarity (we use a somewhat relaxed cut-off as we are comparing across organisms).
-We use the entire human-virus interactome from PHISTO as input to this method.
-Given a human protein $h$, we rank all virus proteins $v$ based on the following simple scheme.
-Let $V(h)$ be a set of all virus proteins interacting with the human protein $h$ and let $H(v)$ be all human proteins interacting with the virus protein $v$.
-We score each pair $(h, v)$ from the bipartite graph as follows: $\max \{ \max $ similarity between $h$ and all human proteins in $H(v)$; $\max $ similarity between $v$ and all virus proteins in $V(h) \}$.
-This baseline is shown in the first row as \textit{Homolog} in Table \ref{resultsTable}.
-
-\subsection{Evaluation setup}
-We compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data.
-For the first setting we randomly split the labeled data from each task into 10\% training and 90\% test, such that the class-skew of 1:100 is maintained in both splits (i.e stratified splits).
-The second setting uses a 30\% training, 70\% test split.
-In each setting we generate ten random splits and average the performance over the ten runs.
-
-We report the area under the precision recall curve (AUC-PR) along with the standard deviation. 
-AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \citep{davis2006} such as ours.
-
-\begin{table*}[h]\caption{Area Under the Precision-Recall curve for each task on the PHISTO dataset without removing the homologs.
+\begin{table*}[!h]\caption{Area Under the Precision-Recall curve for each task on the PHISTO dataset without removing the homologs.
 The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
@@
 \end{table*}
+
+\subsection{Comparison with a homolog baseline}
+In addition to the machine learning methods presented in Section \ref{mlmethods}, we also tried a simple homology based approach.
+As per this method, a virus protein $v$ will interact with a human protein $h$ if it interacts with any other human protein $h'$ similar to $h$.
+The converse holds too.
+Such homology-motivated heuristic approaches have been traditionally popular in the inter-organism PPI prediction literature \citep{vidal:2001,matthews:2001}.
+\cite{mika:2006} offers a commentary on the the success of such approaches in prediction PPI within an organism and the transferability across organisms.
+Modifications of such approaches have been published recently \citep{chen:2009,lin:2013,murakami:2014}.
+
+The virus-virus and human-human protein similarity graphs used in this baseline were computed using protein sequence similarity.
+We ran BLAST sequence alignment (\texttt{blastp}) with an e-value cut-off of 0.1 and the negative logarithm of the e-value was used as a measure of similarity (we use a somewhat relaxed cut-off as we are comparing across organisms).
+We use the entire human-virus interactome from PHISTO as input to this method.
+Given a human protein $h$, we rank all virus proteins $v$ based on the following simple scheme.
+Let $V(h)$ be a set of all virus proteins interacting with the human protein $h$ and let $H(v)$ be all human proteins interacting with the virus protein $v$.
+We score each pair $(h, v)$ from the bipartite graph as follows: $\max \{ \max $ similarity between $h$ and all human proteins in $H(v)$; $\max $ similarity between $v$ and all virus proteins in $V(h) \}$.
+This baseline is shown in the first row as \textit{Homolog} in Table \ref{resultsTable}.
+
+\subsection{Evaluation setup}
+We compare all the methods in two settings, where a small proportion of the available labeled data is randomly sampled and used to train a model which is then evaluated on the remaining data.
+For the first setting we randomly split the labeled data from each task into 10\% training and 90\% test, such that the class-skew of 1:100 is maintained in both splits (i.e stratified splits).
+The second setting uses a 30\% training, 70\% test split.
+In each setting we generate ten random splits and average the performance over the ten runs.
+
+We report the area under the precision recall curve (AUC-PR) along with the standard deviation. 
+AUC-PR has been shown to give a more informative picture of an algorithm's performance than ROC curves in high class imbalance datasets \citep{davis2006} such as ours.
 
@@
 \noindent\textit{Overall performance}
-For reference, the AUC-PR of a random classifier model is $\approx 0.01$ due to the class imbalanced nature of the data (class skew = 1:100).
-The first row is the Homolog heuristic based approach, STL is the single-task baseline and all others are multitask models. In general, we notice that multitask learning benefits
+For reference, the AUC-PR of a random classifier model is $\approx 0.01$ due to the class imbalanced nature of the data.
+In general, we notice that multitask learning benefits
 all tasks. The first three columns show the results in the 10\% setting. 
@@
 The variance in the AUC is high for the Ebola task (column 1) owing to the small number of positives 
-in the training splits (8 positives).
+in the training splits.
 The most benefits for our model are seen in the 30\% setting for all tasks.
@@
 In addition to the training data seen by all other methods, this baseline also uses additional data in the form of the virus-virus, human-human sequence similarities and a virus-human interactome from several other viruses (besides the three tasks).
-Results on the homolog part of our dataset sheds further light on this.
+Results on the partitions shed further light on this.
 
 \noindent\textit{Performance on the non-homologous PPI partition}:
-In Table \ref{nohomresultsTable} we show AUC for all the representative baselines in the 10\% setting. 
-Overall, our method performs best on this partition over all tasks.
+Table \ref{nohomresultsTable} has these results. Overall, our method performs best on this partition over all tasks.
 Comparison with the AUC in Table \ref{resultsTable} shows that AUC for the \textit{Ebola} task is higher on all baselines except Homolog, indicating that
@@
 This effect is the most pronounced for our method (BSL-MTL)
-and demonstrates that we are able to predict such PPI by intelligently sharing information with other tasks. 
+and demonstrates that we are able to predict such PPI by sharing information with other tasks. 
 There are thus two components to the prediction performance - (1) the presence of sequence-similar proteins within a task (2) non-linear similarities between the interactions within a task and across tasks.
@@
 \begin{center}
-\includegraphics[scale=0.28, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flavinet.pdf}
+\includegraphics[scale=0.3, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[scale=0.35, trim = 0.2 0.2 0.2 0.2]{flunet.pdf}
+\includegraphics[scale=0.35, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
 \end{center}
@@
 %Multitask link prediction is an important area with many applications ranging from recommendation systems to biomedical host-pathogen interactions.  
-This work developed and tested a new method based on low-rank matrix completion for sharing information across tasks
+This work developed and tested a new multitask learning method for host-pathogen PPI prediction based on low-rank matrix completion for sharing information across tasks.
+Our method, BSL-MTL exhibited large increases in prediction accuracy compared to a variety of baselines.
 %and showed significant increases in prediction performance. 
-The method was evaluated in the host-pathogen protein interaction domain for three pathogens (three tasks) and exhibited large increases in prediction accuracy.
-The model parameters provide several avenues for
-studying host-pathogen interactions for biologists that can lead to interesting observations and insights.
+%The method was evaluated on host-pathogen protein interaction domain for three pathogens (three tasks) and exhibited large increases in prediction accurac
+The model parameters provide several avenues for further studying host-pathogen PPI that can lead to interesting observations and insights.
 Finally, the model we present is general enough to be applicable on other problems such as: gene-disease relevance prediction across organisms or disease conditions.

### 2016-09-01 paper.tex
new pdf

@@
 
-\begin{table}\caption{Area Under the Precision-Recall curve for each task on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
+\begin{table}[!h]
+\caption{Area Under the Precision-Recall curve for each task on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
 To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}.
@@
 
-\begin{table}\caption{Area Under the Precision-Recall curve for each task on the \textbf{homologous PPI} within each task, in the 10\% setting.
+\begin{table}[!h]
+\caption{Area Under the Precision-Recall curve for each task on the \textbf{homologous PPI} within each task, in the 10\% setting.
 The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks.

### 2016-09-01 paper.tex
new pdf

@@
 
-\subsection{Disadvantages of the Homolog method}
-In Figure \ref{fig:network} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
+%\subsection{Disadvantages of the Homolog method}
+%In Figure \ref{fig:network} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
 The interactions mainly involve human proteins which do not have a homolog in the training data.
@@
 
-\begin{figure}
-\begin{tabular}{c}
-%\begin{minipage}{0.3\textwidth}
+%\begin{figure}
+%\begin{tabular}{c}
+%%\begin{minipage}{0.3\textwidth}
+%%\begin{center}
+%%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%%\end{center}
+%%\end{minipage}
+%%&
+%\begin{minipage}{\textwidth}
 %\begin{center}
-%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
 %\end{center}
 %\end{minipage}
-%&
-\begin{minipage}{\textwidth}
-\begin{center}
-\includegraphics[scale=0.3, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
-\end{center}
-\end{minipage}
-\\
-\begin{minipage}{\textwidth}
-\begin{center}
-\includegraphics[scale=0.35, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
-\end{center}
-\end{minipage}
-\end{tabular}
-\caption{PPI from each task that are found by our method but are missed by the homolog method.
+%\\
+%\begin{minipage}{\textwidth}
+%\begin{center}
+%\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
+%\end{center}
+%\end{minipage}
+%\end{tabular}
+%\caption{PPI from each task that are found by our method but are missed by the homolog method.
 To obtain these, we chose a random train/test split and picked examples from the test data on which the homolog method fails - these are ones for which sequence-similar proteins are not found.
 (a) shows \textit{Ebola}-human PPI, (b) \textit{Hepatitis-C} human PPI and (c) \textit{Influenza}-human PPI.}
-\label{fig:network}
-\end{figure}
+%\label{fig:network}
+%\end{figure}
 

### 2016-09-01 paper.tex
new pdf

@@
 
+\begin{table*}[!h]\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
+The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
+We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
+The standard deviation is also shown.
+The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
+The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
+This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
+\label{resultsTable}
+\begin{small}
+\begin{center}
+%\def\arraystretch{1.2}
+\begin{tabular}{c|ccccccc}
+\toprule
+& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
+STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
+MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
+%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
+Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
+%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
+MTPL \citep{me_ismb_2013} & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
+IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
+BSL-MTL (this work) & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
+\end{tabular}
+\end{center}
+\end{small}
+\end{table*}
+
 \section{Experimental setup}
@@
 \noindent\textbf{MTPL} \citep{me_ismb_2013}: A biologically inspired regularizer is used to capture task similarity.\\ \noindent\textbf{BSL-MTL}: Bilinear sparse low-rank multitask learning, the method developed in this paper.
-
-\begin{table*}[!h]\caption{Area Under the Precision-Recall curve for each task on the PHISTO dataset without removing the homologs.
-The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
-We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
-The standard deviation is also shown.
-The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
-The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
-This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
-\label{resultsTable}
-\begin{small}
-\begin{center}
-\def\arraystretch{1.2}
-\begin{tabular}{c|ccccccc}
-\toprule
-& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
-STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
-MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
-%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
-Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
-%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
-MTPL \citep{me_ismb_2013} & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
-IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
-BSL-MTL (this work) & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-\end{table*}
 
@@
 assign a higher weight to the positives.
-For BSL-MTL, to tune the rank parameter '$k$' we tried: 5, ...
+For BSL-MTL, we tried $k$ = 5, ...
 100 and the regularization parameter controlling the norm of $U$ and $V$ was tuned over the range $\lambda$=$\{1 ...
@@
 $\mu_t$ was varied over $\{0, 0.25, 0.5, 0.75, 1\}$.
-The optimal setting was:  $k=10$. %, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
+%The optimal setting was:  $k=10$. %, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
 
@@
 \begin{table}[!h]
-\caption{Area Under the Precision-Recall curve for each task on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
+\caption{AUC-PR on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
 To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}.
@@
 \begin{center}
-\def\arraystretch{1.2}
+%\def\arraystretch{1.2}
 \begin{tabular}{c|ccc}
@@
 \begin{table}[!h]
-\caption{Area Under the Precision-Recall curve for each task on the \textbf{homologous PPI} within each task, in the 10\% setting.
+\caption{AUC-PR on the \textbf{homologous PPI} within each task, in the 10\% setting.
 The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks.
@@
 \centering
-\includegraphics[scale=0.4, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
+\includegraphics[scale=0.35, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
 \caption{Principal component analysis (PCA) of virus proteins in the original feature space.
@@
 \centering
-\includegraphics[scale=0.27, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
+\includegraphics[scale=0.23, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
 \caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
@@
 \begin{center}
-\def\arraystretch{1.1}
+%\def\arraystretch{1.1}
 \begin{tabular}{|c|c|}
@@
 
-\begin{table}[h]\caption{Area Under the Precision-Recall curve for each task upon addition of a new task - HIV, in the 10\% setting.
+\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
 Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table \ref{resultsTable}.
@@
 \begin{center}
-\def\arraystretch{1.2}
+%\def\arraystretch{1.2}
 \begin{tabular}{c|cccc}

### 2016-09-01 paper.tex
new pdf

@@
 
-\begin{algorithm}[!t]   \caption{Alternating Least Squares}
-   \label{alg}
-   %\begin{minipage}{0.9\textwidth}
-   \begin{small}
-\begin{algorithmic}[1]
-\STATE \textbf{Input}: \\
-$k$ : number of latent factors\\
-$\Gamma$: pairs of entities for initialization\\
-For every task $t$,\\
-${\{\mathbf{x}_{ti}\}, \{\mathbf{y}_{tj}\}}$: feature vectors for pathogen and host proteins resp.\\
-$\Omega_t$: the observed entries of the matrix $M_t$\\
-\STATE \textbf{Initialization}: \\
-\STATE An iteration $r$, let $\mathbf{S}^r$ represent $\{S^r_t\}_{t=1}^T$
-\STATE $\mathbf{S}^0 \leftarrow 0$
-\STATE $U^0 \leftarrow$ top-$k$ left singular vectors and $V^0 \leftarrow$ top-$k$ right singular vectors from the SVD of $\displaystyle{\sum_{(i,j) \in \Gamma}} \mathbf{x}_i \mathbf{y}_j^\intercal$\\
-\STATE $\mathcal{L}^0$ : initial loss
-\REPEAT
-  \STATE $U^{r+1} \leftarrow \displaystyle{\argmin_{U}} \; \mathcal{L}(U, V^r, \mathbf{S}^r)$  \STATE $V^{r+1} \leftarrow \displaystyle{\argmin_{V}} \; \mathcal{L}(U^{r+1}, V, \mathbf{S}^r) $  \STATE For each task $t$\\
-  \hspace{0.5cm} $S^{r+1}_t \leftarrow \displaystyle{\argmin_{S_t}} \; \mathcal{L}(U^{r+1}, V^{r+1}, \mathbf{S}^r_{-t}) $
-  \STATE Compute $\mathcal{L}^{r+1}$ and let $\delta \leftarrow (\mathcal{L}^r - \mathcal{L}^{r+1})/\mathcal{L}^r$
-\UNTIL {$\delta < \tau$}%\While
-\end{algorithmic}\end{small}
-%\end{minipage}
-\end{algorithm}
-
 \noindent \textbf{Optimization}:
@@
 and a block coordinate descent method called alternating least squares (ALS) is commonly used to optimize such functions. 
-To speed up convergence we use an adaptive step size. The detailed optimization procedure is shown in Algorithm \ref{alg}.
+To speed up convergence we use an adaptive step size. The detailed optimization procedure is shown in the supplementary.
+%Algorithm \ref{alg}.
 
@@
 
-\begin{table*}[!h]\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
+\begin{table*}[!b]\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
 The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
@@
 
-\subsection{Parameter tuning}
-We tune the hyper parameters using a 3 fold cross-validation (CV) on the training split. For all baseline regularization parameters %(or a grid search based on the method): 
-we tried the range: [$100, ...10^{-3}$]. To address the class-skew we 
-assign a higher weight to the positives.
-For BSL-MTL, we tried $k$ = 5, ...
-100 and the regularization parameter controlling the norm of $U$ and $V$ was tuned over the range $\lambda$=$\{1 ...
-10^{-3}\}$.
-For each task $t$, $\sigma_t$ was varied over the values $10^{-3}, ...
-10^{-6}$ and
-$\mu_t$ was varied over $\{0, 0.25, 0.5, 0.75, 1\}$.
-%The optimal setting was:  $k=10$. %, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
-
-%For the task-specific regularization parameters, setting them all to a constant i.e $\forall t, \{\sigma_t\} = \sigma$ did not work well owing to the different sizes of the 3 tasks.
-For each task $t$, $\sigma_t$ was varied over the values $10^{-3}, 10^{-4}, 10^{-5}, 10^{-6}$.
-The optimal setting of parameters was found to be $k=10$, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
+\textbf{Parameter tuning}: Please refer to the supplementary.
 
@@
 We find that most of the top entries from $U V^T$ correspond to linear virus features, whereas those from the various $S_t$
-involve bilinear features. We analyze the $k$-mers corresponding to the top 20 features from each of the matrices.
+involve bilinear features. These results are in the supplementary.
 
-Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}): 
-the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged).
-Given the set of top virus features we can obtain the corresponding set of
-amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the 
-training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
-be the set of amino-acid k-mers that appear in the training data.
-Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
-To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif. 
-The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
-want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
-with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
-analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
-We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by 
-hydrophilic amino acids (T, K, R, D, E).
-
-This procedure described above is used to analyze the most significant human protein features, obtained from
-the matrix $U V^T$.
-These motifs are shown in the supplementary.
-The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepc}$ and $S_{flu}$ respectively.
-The motif for Hepatitis is shown in Figure \ref{fig:virus_motifs}(right); the rest can be found in the supplementary.
-%Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
-These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and \ref{fig:flu_motifs}.
-The virus-specific motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G) though S and T do appear in some motifs as well.
-%The human protein motifs are shown in \ref{fig:human_motifs}; the 3-mer motif show less of a pattern, in particular the first residue does not seem to have any dominant amino-acids.
-%However the 4-mer seems to be dominated by hydrophobic amino-acids L, P, I, F in the first three positions and hydrophilic residues S, T, Y in the fourth.
-%The human protein motifs are shown in Fig \ref{fig:human_motifs}.
-In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
-
-%The motifs for tri-mers are shown in the supplementary.
-
-\noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
-We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \citep{iedb}.
-The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:virus_motifs}(right) is found in 53 epitopes.
-From the \textit{Ebola} motif (supplementary), we find that \texttt{TLAT} is part of six different epitopes, \texttt{SLTT} appears in three epitopes.
-\texttt{PLIK}, \texttt{SLLL} from the \textit{Influenza} motif are also found in many epitopes.
-Finding that the virus k-mer patterns predicted by our method are recognized by human antibodies
-is a further validation of its performance.
-Further, using higher dimensional $k$-mers (where $k$=7, 8, 9) as
-features in our model will give motifs from which complete epitopes can be derived. % which can then be verified by biochemical peptide interaction assays.
-Our model thus has applications in epitope prediction as well, where conventional methods consist of scanning all possible $k$-mers from protein sequences to identify likely epitopes.
-
-\noindent\emph{Phosphorylation sites}: 
-We found the frequent occurrence of S and T and sometimes Y in the motifs striking and suspected this may be
-related to the amino acids being frequent targets of phosphorylation.
-Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \citep{phosphosite} are
- repositories for known sites in human proteins.
-Since these are sites in human proteins, we searched for the patterns from the 4-mer motif of human proteins (supplementary) and found several to be flanking known phosphorylation sites in human proteins:
-\texttt{LLLs}, \texttt{LLLt}, \texttt{ILLs}, \texttt{PPPs}, \texttt{PIPs}, \texttt{PIPt}, \texttt{LIPs}, \texttt{PLLt} (lower-case indicates the putative phosphorylation site).
-This observation also supports the notion that the motifs predictive
-of interaction are biologically interesting.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%}
-%\ffigbox[0.31\columnwidth]{%
-%%trim option's parameter order: left bottom right top
-%\includegraphics[scale=0.165, trim = 0 0 0 0]{fig6.eps}
-%}{%
-%\caption{Sequence motif constructed from the top four-mer features of human proteins}%
-%\label{fig:human_motifs}
-%}
-%\end{floatrow}
-%\begin{floatrow}
-%\ffigbox[0.3\columnwidth]{%
-%%trim option's parameter order: left bottom right top
-%\includegraphics[scale=0.165, trim = 0.2cm 0 0 0]{fig7.eps}
-%}{%
-%\caption{Sequence motif from top four-mer features specific to \textit{Ebola} proteins}%
-%\label{fig:ebola_motifs}
-%}
-%
-%\ffigbox[0.01\columnwidth]{}
-%
-%\ffigbox[0.3\columnwidth]{%
-%%trim option's parameter order: left bottom right top
-%\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8.eps}
-%}{%
-%\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
-%\label{fig:hepc_motifs}
-%}
-%
-%\ffigbox[0.3\columnwidth]{%
-%%trim option's parameter order: left bottom right top
-%\includegraphics[scale=0.165, trim = 0 0 0 0]{fig9.eps}
-%}{%
-%\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
-%\label{fig:flu_motifs}
-%}
-%\end{floatrow}
-%\end{figure*}
-
-\section{Results on HIV data}
+\subsection{Results on HIV data}
 We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
@@
 The interactions data has $\approx$4000 human-HIV PPI across several strains of which 1320 are non-homologous PPI.
-This experiment particularly tests the robustness of all the multitask learning approaches by the addition of an unrelated task.
-Table \ref{hivresults} shows the results.
-
-\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
-Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table \ref{resultsTable}.
-Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
-\label{hivresults}
-\begin{small}
-\begin{center}
-%\def\arraystretch{1.2}
-\begin{tabular}{c|cccc}
-\toprule
-& \multicolumn{4}{c}{10 \% training } \\
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
-Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
-STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
-Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
-BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-\end{table}
+Please refer to the supplementary for the results.
 

### 2016-09-01 supplementary/supplementary.tex
new pdf

@@
 
+\section{Optimization algorithm}
+
+\begin{algorithm}[!t]   \caption{Alternating Least Squares}
+   \label{alg}
+   %\begin{minipage}{0.9\textwidth}
+   \begin{small}
+\begin{algorithmic}[1]
+\STATE \textbf{Input}: \\
+$k$ : number of latent factors\\
+$\Gamma$: pairs of entities for initialization\\
+For every task $t$,\\
+${\{\mathbf{x}_{ti}\}, \{\mathbf{y}_{tj}\}}$: feature vectors for pathogen and host proteins resp.\\
+$\Omega_t$: the observed entries of the matrix $M_t$\\
+\STATE \textbf{Initialization}: \\
+\STATE An iteration $r$, let $\mathbf{S}^r$ represent $\{S^r_t\}_{t=1}^T$
+\STATE $\mathbf{S}^0 \leftarrow 0$
+\STATE $U^0 \leftarrow$ top-$k$ left singular vectors and $V^0 \leftarrow$ top-$k$ right singular vectors from the SVD of $\displaystyle{\sum_{(i,j) \in \Gamma}} \mathbf{x}_i \mathbf{y}_j^\intercal$\\
+\STATE $\mathcal{L}^0$ : initial loss
+\REPEAT
+  \STATE $U^{r+1} \leftarrow \displaystyle{\argmin_{U}} \; \mathcal{L}(U, V^r, \mathbf{S}^r)$  \STATE $V^{r+1} \leftarrow \displaystyle{\argmin_{V}} \; \mathcal{L}(U^{r+1}, V, \mathbf{S}^r) $  \STATE For each task $t$\\
+  \hspace{0.5cm} $S^{r+1}_t \leftarrow \displaystyle{\argmin_{S_t}} \; \mathcal{L}(U^{r+1}, V^{r+1}, \mathbf{S}^r_{-t}) $
+  \STATE Compute $\mathcal{L}^{r+1}$ and let $\delta \leftarrow (\mathcal{L}^r - \mathcal{L}^{r+1})/\mathcal{L}^r$
+\UNTIL {$\delta < \tau$}%\While
+\end{algorithmic}\end{small}
+%\end{minipage}
+\end{algorithm}
+
+\section{Results on HIV data}
+We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
+This experiment tests how our multitask model behaves in the presence of a relatively outlier task.
+The interactions data has $\approx$4000 human-HIV PPI across several strains of which 1320 are non-homologous PPI.
+Please refer to the supplementary for the results.
+
+This experiment particularly tests the robustness of all the multitask learning approaches by the addition of an unrelated task.
+Table \ref{hivresults} shows the results.
+
+\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
+Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table \ref{resultsTable}.
+Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
+\label{hivresults}
+\begin{small}
+\begin{center}
+%\def\arraystretch{1.2}
+\begin{tabular}{c|cccc}
+\toprule
+& \multicolumn{4}{c}{10 \% training } \\
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
+Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
+STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
+Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
+BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
+\end{tabular}
+\end{center}
+\end{small}
+\end{table}
+
 \section{Parameter tuning}
@@
 
-\section{10 fold cross validation}
-In the 10 fold cross-validation (CV) experiments, a much larger training set is available.
-So the single task baseline becomes much harder to beat as it can independently tune parameters for each task.
-All methods have a higher variance in their performance (we think this is due to the smaller and hence more variable test data in a 10 fold CV setting).
-Our method improves on only the \textit{Hep-C} task.
+\section{Biological significance of interactions}
 
-\begin{table*}[h]\caption{Area Under the Precision-Recall curve for 10 fold CV. The first row is the only single-task method and all others are multitask models.}
-\label{resultsTable}
-\begin{center}
-\def\arraystretch{1.2}
-\begin{tabular}{c|ccc}
-\toprule
-& \multicolumn{3}{c}{\textbf{10 fold CV}} \\
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
- STL     & 0.28$\pm$.11 & 0.74$\pm$.05 & 0.66$\pm$.08  \\ 
- MMTL    & 0.28$\pm$.11 & 0.68$\pm$.04 & 0.40$\pm$.03  \\ 
- Norm    & 0.20$\pm$.09 & 0.70$\pm$.06 & 0.44$\pm$.05  \\ 
- Rank    & 0.18$\pm$.08 & 0.70$\pm$.06 & 0.45$\pm$.04  \\ 
- MTPL    & 0.27$\pm$.08 & 0.67$\pm$.05 & 0.48$\pm$.06  \\ 
- IMC     & 0.16$\pm$.08 & 0.72$\pm$.06 & 0.45$\pm$.06  \\ 
- BSL-MTL & 0.28$\pm$.11 & 0.82$\pm$.05 & 0.62$\pm$.02  \\ 
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table*}
+We analyze sequence motifs derived from the top $k$-mers that contribute to interactions. The significant entries of the model parameters $U$, $V$ and $\{S_t\}$ were used to
+compute these motifs.
+The top positive-valued entries from the product $U V^T$ indicate which pairs of features: ($(f_v, f_h)$: virus protein feature, human protein feature) are important for interactions across all the virus-human PPI tasks.
+Analogously, the entries from $S_t$ give us pairs of features important to a particular virus-human task `$t$'.
+We find that most of the top entries from $U V^T$ correspond to linear virus features, whereas those from the various $S_t$
+involve bilinear features.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We analyze the $k$-mers corresponding to the top 20 features from each of the matrices.
 
-\section{Homologous interactions}
-Since we use data from several strains for each task, the PPI data contains some interactions that are interologs.
-We observed this for two of the tasks: \textit{Hepatitis-C} and \textit{Influenza A}.
-Note that we did not find any interologs \textit{across} tasks.
-These homologous interactions in the various strains are reported by different experimental studies and we believe their presence suggests the confidence of such interactions.
-Since we use only PPI derived from experimental methods (and not electronically inferred), we do not exclude any homologous interactions from our training.
-\\
-\textbf{Removing homologs from the evaluation}:\\
-\noindent The number of distinct PPI for which interologs or homologs exist is very small ($\approx$20), but there are several homologs for each such PPI.
-Note that any benefits achieved from the presence of homologous interactions will be available to all the methods (since the models were built using identical train/test data).
-Here we present results on the 10\% setting by removing all homologs from the test data (note: the \textit{Ebola} task did not have any homologs).
-For lack of space we only mention the trend for our method (BSL-MTL) - \textit{Hep-C}: 0.85 and \textit{Influenza:} 0.45.
-The other methods show a similar trend; our method continues to outperform by significant margins.
+Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}):
+the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged).
+Given the set of top virus features we can obtain the corresponding set of
+amino-acid $k$-mers, say $AA_v$, by reversing the feature-generation step. However most of the possible $k$-mers do not appear in the
+training data (ex: out of the 160,000 (=$20^4$) possible 4-mers $\approx$24,000 appear). Let $AA_{tr}$
+be the set of amino-acid k-mers that appear in the training data.
+Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
+To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif.
+The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
+want to summarize, we use the Shannon logo type (which does not consider any background amino-acid distribution)
+with the following settings: clustering-method=None, weight-on-prior=1 (pseudo-counts do not make sense in our
+analysis). Figure \ref{fig:virus_motifs} shows the motif that is common across viruses. %, while the other figures highlight those motifs important in a specific species.
+We observe that the shared tri-mer motif for virus proteins in Figure \ref{fig:virus_motifs} is dominated by
+hydrophilic amino acids (T, K, R, D, E).
 
-\section{Biological significance of interactions} 
+This procedure described above is used to analyze the most significant human protein features, obtained from
+the matrix $U V^T$.
+These motifs are shown in the supplementary.
+The task-specific features i.e significant $k$-mers from virus-proteins of \textit{Ebola}, \textit{Hepatitis} and \textit{Influenza} are obtained from the matrices $S_{ebola}$, $S_{hepc}$ and $S_{flu}$ respectively.
+The motif for Hepatitis is shown in Figure \ref{fig:virus_motifs}(right); the rest can be found in the supplementary.
+%Knowledge of these pathogen-specific $k$-mers can help in design of drugs to target specific pathogens.
+These are shown in Figures \ref{fig:ebola_motifs}, \ref{fig:hepc_motifs} and \ref{fig:flu_motifs}.
+The virus-specific motifs seem to be dominated by hydrophobic residues (I, P, L, V, A, G) though S and T do appear in some motifs as well.
+%The human protein motifs are shown in \ref{fig:human_motifs}; the 3-mer motif show less of a pattern, in particular the first residue does not seem to have any dominant amino-acids.
+%However the 4-mer seems to be dominated by hydrophobic amino-acids L, P, I, F in the first three positions and hydrophilic residues S, T, Y in the fourth.
+%The human protein motifs are shown in Fig \ref{fig:human_motifs}.
+In most cases, the first position of the tri-mer was less significant than the second and third, while for the tetramer all four positions show clear preferences.
+
+%The motifs for tri-mers are shown in the supplementary.
+
  \noindent\emph{Phosphorylation sites}: 

### 2016-09-01 paper.tex
new pdf

@@
 The second setting uses a 30\% training, 70\% test split.
-In each setting we generate ten random splits and average the performance over the ten runs.
+In each setting we generate ten random splits and average the performance over the ten runs (parameter tuning details are in the supplementary).
 
@@
 new nodes in test data.
-
-\textbf{Parameter tuning}: Please refer to the supplementary.
 
@@
 Analogously, the entries from $S_t$ give us pairs of features important to a particular virus-human task `$t$'.
-We find that most of the top entries from $U V^T$ correspond to linear virus features, whereas those from the various $S_t$
-involve bilinear features. These results are in the supplementary.
+These results are in the supplementary.
 
-\subsection{Results on HIV data}
+\noindent\textbf{Results on HIV data}: 
 We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).

### 2016-09-01 paper.tex
new pdf

@@
 Analogously, the entries from $S_t$ give us pairs of features important to a particular virus-human task `$t$'.
-These results are in the supplementary.
+These results are in the supplementary.\\
 
@@
 
-\begin{figure}
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4-eps-converted-to.pdf}
-\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8-eps-converted-to.pdf}
-\caption{(\textit{Left}) Top tri-mer sequence motifs from virus proteins across all three viruses.
-These correspond to features important for interactions and is derived from our model parameter $H = U V^{T}$.
-(\textit{Right}) Motif corresponding to enriched 4-mer patterns found by our model (parameter: $S_{hepc}$) for the Hepatitis-C virus task.}%
-\label{fig:virus_motifs}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-09-01 supplementary/supplementary.tex
new pdf

@@
 
+\begin{figure}
+%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.165, trim = 0 0.8cm 0 0]{fig4-eps-converted-to.pdf}
+\includegraphics[scale=0.165, trim = 0 0 0 0]{fig8-eps-converted-to.pdf}
+\caption{(\textit{Left}) Top tri-mer sequence motifs from virus proteins across all three viruses.
+These correspond to features important for interactions and is derived from our model parameter $H = U V^{T}$.
+(\textit{Right}) Motif corresponding to enriched 4-mer patterns found by our model (parameter: $S_{hepc}$) for the Hepatitis-C virus task.}%
+\label{fig:virus_motifs}
+\end{figure}
+
 \begin{figure}[h]
 \begin{floatrow}
-\ffigbox[0.4\columnwidth]{%
+\ffigbox[0.3\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 
-\ffigbox[0.4\columnwidth]{%
+\ffigbox[0.3\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 
-\end{floatrow}
-\end{figure}
-
-\begin{figure}[h]
-\begin{floatrow}
-\ffigbox[0.4\columnwidth]{%
+\ffigbox[0.3\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 }
+
 \end{floatrow}
 \end{figure}
-
-\newpage
 

### 2016-09-01 paper.tex
new pdf

@@
 \begin{figure*}
-\begin{floatrow}
-\ffigbox[0.28\textwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
-}
-{
-\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
-\label{fig:hpppigraph}
-}
 \capbtabbox[0.68\textwidth]{%
@@
 }
+
+\begin{floatrow}
+\ffigbox[0.28\textwidth]{%
+%trim option's parameter order: left bottom right top
+%\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
+\includegraphics[scale=0.35, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}}
+{
+\caption{Principal component analysis (PCA) of virus proteins in the original feature space.
+The first two principal components are shown.
+Shape and color of the points indicates which virus that protein comes from.}%
+\label{fig:pca1}
+%\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
+%\label{fig:hpppigraph}
+}
+
 \end{floatrow}
@@
 
-\begin{table*}[!b]\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
-The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
-We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
-The standard deviation is also shown.
-The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
-The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
-This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
-\label{resultsTable}
+\begin{figure*}
+\capbtabbox[0.68\textwidth]{%
 \begin{small}
@@
 \end{small}
-\end{table*}
+}
+{
+\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
+The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
+We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
+The standard deviation is also shown.
+The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
+The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
+This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
+\label{resultsTable}
+}
+
+\begin{floatrow}
+\ffigbox[0.28\textwidth]{%
+\includegraphics[scale=0.23, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
+}
+{
+\caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
+The first two principal components are shown.
+Shape of the points indicates which virus that protein comes from.
+We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
+\label{fig:pca2}
+}
+\end{floatrow}
+\end{figure*}
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[scale=0.35, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
-\caption{Principal component analysis (PCA) of virus proteins in the original feature space.
-The first two principal components are shown.
-Shape and color of the points indicates which virus that protein comes from.}%
-\label{fig:pca1}
-\end{figure}
-
-\begin{figure}[h]
-\centering
-\includegraphics[scale=0.23, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
-\caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
-The first two principal components are shown.
-Shape of the points indicates which virus that protein comes from.
-We have highlighted three clusters for which we show GO term enrichment analysis results in Table \ref{goterms}.}%
-\label{fig:pca2}
-\end{figure}
-
 \begin{table}[bht]

### 2016-09-01 paper.tex
new pdf

@@
 An elegant way to formulate the interaction prediction problem is via a graph completion based framework, where we have several bipartite 
-graphs over multiple hosts and pathogens as illustrated in Figure \ref{fig:hpppigraph}.
+graphs over multiple hosts and pathogens. % as illustrated in Figure \ref{fig:hpppigraph}.
 Nodes in the graphs represent host proteins (circles) and pathogen proteins (triangles), with edges between them representing interactions 
@@
 \begin{figure*}
+\begin{floatrow}
 \capbtabbox[0.68\textwidth]{%
@@
 
-\begin{floatrow}
-\ffigbox[0.28\textwidth]{%
+\ffigbox[0.3\textwidth]{%
 %trim option's parameter order: left bottom right top
 %\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
-\includegraphics[scale=0.35, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}}
+\includegraphics[scale=0.32, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
+}
 {
@@
 }
-
 \end{floatrow}
@@
 STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
-MMTL \citep{pontil04} & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
+MMTL & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
 %Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
-Sparse+low-rank \citep{chen2012} & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
+Sparse+low-rank & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
 %Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
-MTPL \citep{me_ismb_2013} & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
-IMC \citep{nagarajan} & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
-BSL-MTL (this work) & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
+MTPL & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
+IMC & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
+BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
 \end{tabular}
@@
 
-\begin{floatrow}
 \ffigbox[0.28\textwidth]{%
-\includegraphics[scale=0.23, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
+\includegraphics[scale=0.21, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
 }

### 2016-09-01 paper.tex
new pdf

@@
 }
-
-\ffigbox[0.3\textwidth]{%
+\ffigbox[0.28\textwidth]{%
 %trim option's parameter order: left bottom right top
@@
 & \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & 0.399 $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
+Homolog & \textbf{0.399} $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
 STL   & 0.192 $\pm$.15 & 0.504 $\pm$.10 & 0.305 $\pm$.03 \\
 Sparse+LR & 0.184 $\pm$.13 & 0.512 $\pm$.10 & 0.315 $\pm$.04 \\ 
-BSL-MTL & 0.130 $\pm$.10 & 0.815 $\pm$.02 & 0.475 $\pm$.02 \\ \bottomrule
+BSL-MTL & 0.130 $\pm$.10 & \textbf{0.815} $\pm$.02 & \textbf{0.475} $\pm$.02 \\ \bottomrule
 \end{tabular}

### 2016-09-01 supplementary/supplementary.tex
new pdf

@@
 
+\usepackage{floatrow}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{subfigure} 
-\usepackage{floatrow}
+\usepackage{amssymb}
+\usepackage{amsmath}
 \usepackage{booktabs}
@@
 \usepackage{url}
+
+\newcommand{\argmin}{\operatornamewithlimits{argmin}}
 
@@
 \begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
-Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table \ref{resultsTable}.
+Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2.
 Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
@@
 
-Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section \ref{sec:datasets}):
+Note that our features do not directly correspond to a unique amino-acid $k$-mer (see Section 4.2):
 the virus feature $f_v$ will map to several amino-acid sequences (for instance KKCC, KRCC, RRCC etc all map to a single feature due to the molecular similarity between the amino acids K and R being both positively charged).
@@
 Then, the intersection $I_v = AA_v \cap AA_{tr}$ gives us the important amino-acid $k$-mers from virus proteins w.r.t interaction prediction.
-To summarize $I_v$, we use a popular tool Seq2Logo \citep{seq2logo} to generate a sequence motif.
+To summarize $I_v$, we use a popular tool Seq2Logo to generate a sequence motif.
 The logos for the two-, three-, four-mers from $I_v$ are generated independently. Since we only
@@
  related to the amino acids being frequent targets of phosphorylation.
- Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite \cite{phosphosite} are
+ Phosphorylated sites often serve as PPI sites, and databases such as Phosphosite are
   repositories for known sites in human proteins. Since these are sites in human proteins, we searched for the patterns
@@
 \noindent\emph{Evidence in IEDB}\footnote{\url{www.iedb.org}}:
-We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB) \cite{iedb}.
+We found experimental evidence for the significance for the virus motifs in the Immune Epitope\footnote{An epitope is a very short sequence from the virus that binds to human antibodies} database (IEDB).
 The pattern \texttt{IVGG} from the \textit{Hepatitis-C} motif in Fig. \ref{fig:hepc_motifs} is found in 53 epitopes.
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{top10ebola_3mers_seq2logo.eps}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10ebola_3mers_seq2logo.pdf}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flavi_3mers_seq2logo.eps}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flavi_3mers_seq2logo.pdf}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flu_3mers_seq2logo.eps}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flu_3mers_seq2logo.pdf}
 }{%
@@
 %trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{top20human_3mers_seq2logo.eps}
+\includegraphics[scale=0.25, trim = 0 0 0 0]{top10human_3mers_seq2logo.pdf}
 }{%

### 2016-09-01 supplementary/supplementary.tex
new pdf

@@
 \documentclass[a4paper,11pt]{article}
+
+\usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
 
@@
 
-\title{Supplementary Information}
+\noindent{\Huge Supplementary Information}
 
@@
 
-\begin{algorithm}[!t]   \caption{Alternating Least Squares}
+\begin{algorithm}[!h]   \caption{Alternating Least Squares}
    \label{alg}
@@
 \begin{floatrow}
-\ffigbox[0.3\columnwidth]{%
+\ffigbox[0.45\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 
-\ffigbox[0.3\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{fig8-eps-converted-to.pdf}
-}{%
-\caption{Sequence motifs specific to \textit{Hepatitic-C} proteins that are also important to interactions.}%
-\label{fig:hepc_motifs}
-}
-
-\ffigbox[0.3\columnwidth]{%
+\ffigbox[0.45\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 \begin{floatrow}
-\ffigbox[0.4\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{top10ebola_3mers_seq2logo.pdf}
-}{%
-\caption{Sequence motifs specific to \textit{Ebola} proteins that are also important to interactions.}%
-\label{fig:hepc_motifs}
-}
-
-\ffigbox[0.4\columnwidth]{%
+\ffigbox[0.45\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 }
-\end{floatrow}
-\end{figure}
-
-\begin{figure}[h]
-\begin{floatrow}
-\ffigbox[0.4\columnwidth]{%
-%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.25, trim = 0 0 0 0]{top10flu_3mers_seq2logo.pdf}
-}{%
-\caption{Sequence motifs specific to \textit{Influenza} proteins that are also important to interactions.}%
-\label{fig:flu_motifs}
-}
-\ffigbox[0.4\columnwidth]{%
+\ffigbox[0.45\columnwidth]{%
 %trim option's parameter order: left bottom right top
@@
 
-\bibliographystyle{plain}
-\bibliography{references}
+%\bibliographystyle{plain}
+%\bibliography{references}
 

### 2016-09-01 paper.tex
new pdf

@@
 \usepackage{amsmath}
-\usepackage{algorithm}
-\usepackage{algorithmic}
 \usepackage{booktabs}
@@
 The experimental discovery of host-pathogen protein interactions involves biochemical and biophysical methods such as co-immunoprecipitation (co-IP), yeast two-hybrid (Y2H) assays, co-crystalization.
-The host-pathogen protein interactions from several small-scale and high throughput experiments are aggregated by databases such as virus-MINT\cite{virusmint}, HPIDB \cite{hpidb}, PHISTO \cite{phisto} etc by literature curation.
+The host-pathogen protein interactions from several small-scale and high throughput experiments are aggregated by databases such as MINT \citep{virusmint}, HPIDB \citep{hpidb}, PHISTO \citep{phisto} etc by literature curation.
 These databases are valuable sources of information to bioinformaticians for developing models.
@@
 In particular, supervised machine learning based methods use the few known interactions as training data and formulate the interaction prediction problem in a classification setting, with target classes: ``interacting'' or ``non-interacting''.
-Features are derived using various attributes of the two proteins such as: protein sequences from UniProt \cite{UniProt}, protein structure from PDB, gene ontology from GO database \cite{GO}.
+Features are derived using various attributes of the two proteins such as: protein sequences from UniProt \citep{UniProt}, protein structure from PDB, gene ontology from GO database \citep{GO}.
 
@@
 For a uniform comparison we used least squared loss in all the methods.
-The MALSAR \cite{malsar} package was used to implement some of the models.
+The MALSAR \citep{malsar} package was used to implement some of the models.
 For the baselines wherever appropriate, we concatenated the features of the two node types into a single feature vector.
@@
 
-\noindent\textit{Performance on the homologous PPI partition}:
+\noindent\textit{Performance on the homologous PPI partition} (Table \ref{homresultsTable}):
 Except the \textit{Ebola} task, BSL-MTL has the highest AUC.
@@
 \begin{table}[bht]
-\label{goterms}
 \begin{scriptsize}
@@
 The first column also indicates the composition of each cluster (i.e which task's proteins appear in the cluster).}
+\label{goterms}
 \end{center}

### 2016-09-01 paper.tex
with eps figures, latex compilation

@@
 %\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
-\includegraphics[scale=0.32, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
+\includegraphics[scale=0.32, trim = 0 0.2cm 0 0.5cm]{pca_original_feats.eps}
 }
@@
 \ffigbox[0.28\textwidth]{%
-\includegraphics[scale=0.21, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
+\includegraphics[scale=0.21, trim = 0 5.6cm 0 6cm]{pca_bslmtl_with_clusters.eps}
 }

### 2016-09-01 paper.tex
with eps figures, latex compilation

--- 2016-09-01 paper.tex
@@
 %\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
-\includegraphics[scale=0.32, trim = 0 0.2cm 0 0.5cm]{fig2-eps-converted-to.pdf}
+\includegraphics[scale=0.32, trim = 0 0 0 0]{pca_original_feats.eps}
 }
@@
 \ffigbox[0.28\textwidth]{%
-\includegraphics[scale=0.21, trim = 0 5.6cm 0 6cm]{pca_with_clusters.pdf}
+\includegraphics[scale=0.21, trim = 0 0 0 0]{pca_bslmtl_with_clusters.eps}
 }

### 2016-09-01 paper.tex
with eps figures, latex compilation

@@
 %\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
-\includegraphics[scale=0.32, trim = 0 0 0 0]{pca_original_feats.eps}
+\includegraphics[scale=0.34, trim = 0 0 0 0]{pca_original_feats.eps}
 }
@@
 \ffigbox[0.28\textwidth]{%
-\includegraphics[scale=0.21, trim = 0 0 0 0]{pca_bslmtl_with_clusters.eps}
+%%trim option's parameter order: left bottom right top
+\includegraphics[scale=0.28, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
 }

### 2016-09-01 paper.tex
with eps figures, latex compilation

@@
 
-%\subsection{Disadvantages of the Homolog method}
-%In Figure \ref{fig:network} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
-The interactions mainly involve human proteins which do not have a homolog in the training data.
-This illustrates that our features capture more information than mere sequence similarity.
-Further, inclusion of interactions from other viruses in the na\"ive manner as done by the homologs in fact hurts the performance (results in too many false positives) - our model is able to learn what to borrow from the other tasks accurately.
-
 \begin{table}[!h]
@@
 For enrichment analysis, we use the FuncAssociate 3.0 \citep{funcAsso} tool and GO annotations for the three viruses from UniProtKB.
-
-%\begin{figure}
-%\begin{tabular}{c}
-%%\begin{minipage}{0.3\textwidth}
-%%\begin{center}
-%%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
-%%\end{center}
-%%\end{minipage}
-%%&
-%\begin{minipage}{\textwidth}
-%\begin{center}
-%\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
-%\end{center}
-%\end{minipage}
-%\\
-%\begin{minipage}{\textwidth}
-%\begin{center}
-%\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
-%\end{center}
-%\end{minipage}
-%\end{tabular}
-%\caption{PPI from each task that are found by our method but are missed by the homolog method.
-To obtain these, we chose a random train/test split and picked examples from the test data on which the homolog method fails - these are ones for which sequence-similar proteins are not found.
-(a) shows \textit{Ebola}-human PPI, (b) \textit{Hepatitis-C} human PPI and (c) \textit{Influenza}-human PPI.}
-%\label{fig:network}
-%\end{figure}
 
@@
 
-\noindent\textbf{Results on HIV data}: 
+\section{Results on HIV data}
 We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
@@
 Please refer to the supplementary for the results.
+
+This experiment particularly tests the robustness of all the multitask learning approaches by the addition of an unrelated task.
+Table \ref{hivresults} shows the results.
+
+\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
+Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2.
+Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
+\label{hivresults}
+\begin{small}
+\begin{center}
+%\def\arraystretch{1.2}
+\begin{tabular}{c|cccc}
+\toprule
+& \multicolumn{4}{c}{10 \% training } \\
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
+Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
+STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
+Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
+BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
+\end{tabular}
+\end{center}
+\end{small}
+\end{table}
 

### 2016-09-01 supplementary/supplementary.tex
with eps figures, latex compilation

@@
 
-\section{Results on HIV data}
-We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
-This experiment tests how our multitask model behaves in the presence of a relatively outlier task.
-The interactions data has $\approx$4000 human-HIV PPI across several strains of which 1320 are non-homologous PPI.
-Please refer to the supplementary for the results.
-
-This experiment particularly tests the robustness of all the multitask learning approaches by the addition of an unrelated task.
-Table \ref{hivresults} shows the results.
-
-\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
-Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2.
-Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
-\label{hivresults}
-\begin{small}
-\begin{center}
-%\def\arraystretch{1.2}
-\begin{tabular}{c|cccc}
-\toprule
-& \multicolumn{4}{c}{10 \% training } \\
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
-Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
-STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
-Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
-BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-\end{table}
-
 \section{Parameter tuning}
@@
 The optimal setting was:  $k=10$. %, $\lambda = 0.01$, $\sigma_{ebola}=10^{-5}$, $\sigma_{flu}=\sigma_{hepc}=10^{-6}$.
+
+\subsection{Disadvantages of the Homolog method}
+In Figure \ref{fig:network} we show for each task, the protein interactions in the test data (for one train-test split) that were not found by the homologs model which relies purely on homology-based sequence similarity.
+The interactions mainly involve human proteins which do not have a homolog in the training data.
+This illustrates that our features capture more information than mere sequence similarity.
+Further, inclusion of interactions from other viruses in the na\"ive manner as done by the homologs in fact hurts the performance (results in too many false positives) - our model is able to learn what to borrow from the other tasks accurately.
+
+\begin{figure}
+\begin{tabular}{c}
+%\begin{minipage}{0.3\textwidth}
+%\begin{center}
+%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%\end{center}
+%\end{minipage}
+%&
+\begin{minipage}{0.4\textwidth}
+\begin{center}
+\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
+\end{center}
+\end{minipage}
+\\
+\begin{minipage}{0.4\textwidth}
+\begin{center}
+\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
+\end{center}
+\end{minipage}
+\end{tabular}
+\caption{PPI from each task that are found by our method but are missed by the homolog method.
+To obtain these, we chose a random train/test split and picked examples from the test data on which the homolog method fails - these are ones for which sequence-similar proteins are not found.
+(a) shows \textit{Ebola}-human PPI, (b) \textit{Hepatitis-C} human PPI and (c) \textit{Influenza}-human PPI.}
+\label{fig:network}
+\end{figure}
 

### 2016-09-01 paper.tex
with eps figures, latex compilation

@@
 We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
-This experiment tests how our multitask model behaves in the presence of a relatively outlier task.
-The interactions data has $\approx$4000 human-HIV PPI across several strains of which 1320 are non-homologous PPI.
-Please refer to the supplementary for the results.
-
-This experiment particularly tests the robustness of all the multitask learning approaches by the addition of an unrelated task.
+The interactions data (obtained from PHISTO and NIAID databases) has $\approx$4000 human-HIV PPI across several strains, of which 1320 PPI are non-homologous.
+This experiment particularly tests the robustness of the multitask learning approaches by the addition of a relatively unrelated task.
 Table \ref{hivresults} shows the results.
@@
 Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2.
-Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of an unrelated task such as HIV.}
+Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of a somewhat dissimilar task such as HIV.}
 \label{hivresults}

### 2016-09-01 paper.tex
with eps figures, latex compilation

@@
 This observation is not obvious in the PCA of the original feature space (Fig \ref{fig:pca1}), where proteins with similar sequences group together.
-We analyze the projected data further by looking at the clusters of proteins for enrichment of Gene Ontology (GO) annotations (proteins were first clustered in this lower dimensional space using k-means and setting the number of clusters to 6).
+We analyze the projected data further by looking at the clusters of proteins for enrichment of Gene Ontology (GO) \footnote{\url{http://www.geneontology.org/}} \citep{GO} annotations (proteins were first clustered in this lower dimensional space using k-means and setting the number of clusters to 6).
 Of particular interest are the highlighted three clusters which contain either proteins projected far from others (such as cluster-1) and proteins from different viruses projected close together (cluster-2 and cluster-3).
-For enrichment analysis, we use the FuncAssociate 3.0 \citep{funcAsso} tool and GO annotations for the three viruses from UniProtKB.
+For enrichment analysis, we use the FuncAssociate 3.0 \footnote{\url{http://llama.mshri.on.ca/funcassociate/}} \citep{funcAsso} tool and GO annotations for the three viruses from UniProtKB.
 

### 2016-09-02 paper.tex
adding eps figures

@@
 
+\noindent\textit{Disadvantages of the Homolog method}:
+In Figure \ref{fig:network} we show for two tasks, the protein interactions in the test data (for one train-test split) that were not found by the Homolog model which relies purely on homology-based sequence similarity.
+%The interactions mainly involve human proteins which do not have a homolog in the training data.
+This illustrates that our features capture more information than mere sequence similarity.
+Further, inclusion of interactions from other viruses in the na\"ive manner as done by the homologs in fact hurts the performance (results in too many false positives) - our model is able to learn what to borrow from the other tasks accurately.
+
+\begin{figure}
+\begin{tabular}{c}
+%\begin{minipage}{0.3\textwidth}
+%\begin{center}
+%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%\end{center}
+%\end{minipage}
+%&
+\begin{minipage}{\textwidth}
+\begin{center}
+\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.eps}
+\end{center}
+\end{minipage}
+\\
+\begin{minipage}{\textwidth}
+\begin{center}
+\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.eps}
+\end{center}
+\end{minipage}
+\end{tabular}
+\caption{PPI from each task that are found by our method but are missed by the homolog method.
+These are ones for which sequence-similar proteins were not found.
+(a) \textit{Hepatitis-C} human PPI and (b) \textit{Influenza}-human PPI.}
+\label{fig:network}
+\end{figure}
+
 \subsection{Biological significance of the model}

### 2016-09-02 supplementary/supplementary.tex
adding eps figures

@@
 
-\begin{figure}
-\begin{tabular}{c}
-%\begin{minipage}{0.3\textwidth}
+%\begin{figure}
+%\begin{tabular}{c}
+%%\begin{minipage}{0.3\textwidth}
+%%\begin{center}
+%%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%%\end{center}
+%%\end{minipage}
+%%&
+%\begin{minipage}{0.4\textwidth}
 %\begin{center}
-%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
+%\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
 %\end{center}
 %\end{minipage}
-%&
-\begin{minipage}{0.4\textwidth}
-\begin{center}
-\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.pdf}
-\end{center}
-\end{minipage}
-\\
-\begin{minipage}{0.4\textwidth}
-\begin{center}
-\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
-\end{center}
-\end{minipage}
-\end{tabular}
-\caption{PPI from each task that are found by our method but are missed by the homolog method.
+%\\
+%\begin{minipage}{0.4\textwidth}
+%\begin{center}
+%\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.pdf}
+%\end{center}
+%\end{minipage}
+%\end{tabular}
+%\caption{PPI from each task that are found by our method but are missed by the homolog method.
 To obtain these, we chose a random train/test split and picked examples from the test data on which the homolog method fails - these are ones for which sequence-similar proteins are not found.
 (a) shows \textit{Ebola}-human PPI, (b) \textit{Hepatitis-C} human PPI and (c) \textit{Influenza}-human PPI.}
-\label{fig:network}
-\end{figure}
+%\label{fig:network}
+%\end{figure}
 

### 2016-11-22 paper.tex
adding eps figures
new stuff

@@
-\documentclass{bioinfo}
-\copyrightyear{2015} \pubyear{2015}
+\documentclass[journal,12pt,onecolumn,draftclsnofoot,]{IEEEtran}
 
@@
 \usepackage{natbib}
+\usepackage{hyperref}
 %\usepackage[footnotesize]{caption}
@@
 \usepackage{url}
+
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd{\@makecaption}
+  {\scshape}
+  {}
+  {}
+  {}
+\makeatletter
+\patchcmd{\@makecaption}
+  {\\}
+  {.\ }
+  {}
+  {}
+\makeatother
+\def\tablename{Table}
+\renewcommand{\thetable}{\arabic{table}}
 
@@
 
-\access{Advance Access Publication Date: Day Month Year}
-\appnotes{Manuscript Category}
+\begin{document}
 
-\begin{document}
-\firstpage{1}
+\title{Multitask matrix completion for learning protein interactions across diseases}
 
-\subtitle{Subject Section}
+\author{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
+%\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
+%$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
+%$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
+%}
 
-\title[Multitask matrix completion]{Multitask matrix completion for learning protein interactions across diseases}
-
-\author[Kshirsagar \textit{et~al}.]{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
-\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
-$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
-$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
-}
-
-\corresp{$^\ast$To whom correspondence should be addressed.}
-
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
-
-\editor{Associate Editor: XXXXXXX}
-
-\abstract{\textbf{Motivation:}
+\begin{abstract}
+\textbf{Motivation:}
 Disease causing pathogens such as viruses, introduce their proteins into the host cells 
@@
 \textbf{Contact:}
-\href{mkshirsa@cs.cmu.edu}{mkshirsa@cs.cmu.edu}
-}
+\href{mkshirsa@cs.cmu.edu}{meghana.ksagar@gmail.com}
+\end{abstract}
 
@@
 
-\begin{figure*}
-\begin{floatrow}
-\capbtabbox[0.68\textwidth]{%
-\begin{footnotesize}
-\begin{tabular}{c|r|r|r} 
-\hline
-Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
-& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
-number of HP PPIs (positives) & 848 & 981 & 90 \\\hline
-number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
-\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
-\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
-total \# of virus proteins in & 542 & 163 & 38 \\
-UniProtKB across strains $^\dagger$ & & & \\ \hline
-number of negatives & 84800 & 98100 & 9000 \\\hline
-density of observed graph$^\ddagger$ (as \%) & .15 & .60 & .20 \\\hline
-\multicolumn{4}{l}{\scriptsize{HP PPI: host-pathogen protein protein interactions}}\\
-\multicolumn{4}{l}{\scriptsize{$\ddagger$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed }} \\
-\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} considering all proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
-\multicolumn{4}{l}{\scriptsize{$\dagger$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
-\end{tabular}
-\end{footnotesize}
-}
-{
-\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header).
-More importantly, each task represents several strains of one virus species.
-We pool together PPI from all strains as the number of interactions per strain is very small and not representative.
-PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus.
-The second row gives a conservative estimate of the number of non-homologous PPI.
-These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details).
-All pathogens are single stranded RNA viruses.
-The last row shows that each of our graphs is extremely sparse.}
-\label{datasets}
-}
-\ffigbox[0.28\textwidth]{%
+\begin{figure}
 %trim option's parameter order: left bottom right top
 %\includegraphics[scale=0.35, trim = 0.2 9.2cm 4cm 0]{fig1-eps-converted-to.pdf}
-\includegraphics[scale=0.34, trim = 0 0 0 0]{pca_original_feats.eps}
-}
-{
+\includegraphics[scale=0.34, trim = 0 0 0 0]{Kshirsagar.Fig_1.eps}
 \caption{Principal component analysis (PCA) of virus proteins in the original feature space.
@@
 \label{fig:pca1}
-%\caption{Multiple bipartite graphs: on the left are host proteins and on the right are proteins from different virus species. Edges represent protein interactions.}
-%\label{fig:hpppigraph}
-}
-\end{floatrow}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}
-\begin{floatrow}
-\capbtabbox[0.68\textwidth]{%
-\begin{small}
-\begin{center}
-%\def\arraystretch{1.2}
-\begin{tabular}{c|ccccccc}
-\toprule
-& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
-STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
-MMTL & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
-%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
-Sparse+low-rank & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
-%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
-MTPL & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
-IMC & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
-BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-}
-{
-\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
-The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
-We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
-The standard deviation is also shown.
-The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
-The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
-This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
-\label{resultsTable}
-}
-
-\ffigbox[0.28\textwidth]{%
-%%trim option's parameter order: left bottom right top
-\includegraphics[scale=0.28, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
-}
-{
+\begin{figure}
+\includegraphics[scale=0.25, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
 \caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
@@
 \label{fig:pca2}
-}
-\end{floatrow}
-\end{figure*}
+\end{figure}
 
@@
 We therefore present results on three sets:
-\begin{enumerate}[a]
+\begin{enumerate}
 \item the original PPIs from PHISTO. AUC-PR is shown in Table \ref{resultsTable}
@@
 
-\begin{table}[!h]
-\caption{AUC-PR on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
-To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}.
-Only the most representative baselines are retained for this experiment.
-Our model performs better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
-\label{nohomresultsTable}
-\begin{small}
-\begin{center}
-%\def\arraystretch{1.2}
-\begin{tabular}{c|ccc}
-\toprule
-\multicolumn{4}{c}{\textbf{10 \% training, 90\% test on Non-homologous PPI}} \\
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & 0.207 $\pm$.07 & 0.113 $\pm$.01 & 0.109 $\pm$ .01 \\
-STL  & 0.210 $\pm$.14 & 0.474 $\pm$ .06 & 0.302 $\pm$.03 \\
-Sparse+LR & 0.216 $\pm$.13 & 0.482 $\pm$.05 & 0.312 $\pm$.04 \\ 
-BSL-MTL & \textbf{0.295} $\pm$.10 & \textbf{0.544} $\pm$.06 & \textbf{0.505} $\pm$.05 \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-\end{table}
-
 \noindent\textit{Performance on the homologous PPI partition} (Table \ref{homresultsTable}):
@@
 The same can be said of the \textit{Influenza} task.
-
-\begin{table}[!h]
-\caption{AUC-PR on the \textbf{homologous PPI} within each task, in the 10\% setting.
-The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks.
-The Homolog method has access to a wider virus-human PPI graph involving several other viruses and is able to infer homologous PPI for \textit{Ebola} better (note that some of the homologous PPI are not redundant examples but involve proteins with very small regions of sequence similarity.}
-\label{homresultsTable}
-\begin{small}
-\begin{center}
-%\def\arraystretch{1.2}
-\begin{tabular}{c|ccc}
-\toprule
-\multicolumn{4}{c}{\textbf{10 \% training, test only Homologous PPI }} \\
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
-Homolog & \textbf{0.399} $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
-STL   & 0.192 $\pm$.15 & 0.504 $\pm$.10 & 0.305 $\pm$.03 \\
-Sparse+LR & 0.184 $\pm$.13 & 0.512 $\pm$.10 & 0.315 $\pm$.04 \\ 
-BSL-MTL & 0.130 $\pm$.10 & \textbf{0.815} $\pm$.02 & \textbf{0.475} $\pm$.02 \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-\end{table}
 
@@
 \begin{figure}
-\begin{tabular}{c}
-%\begin{minipage}{0.3\textwidth}
-%\begin{center}
-%\includegraphics[scale=0.23, trim = 0.3 0 0.2 0.2]{ebolanet.pdf}\vspace{0.6cm}
-%\end{center}
-%\end{minipage}
-%&
-\begin{minipage}{\textwidth}
+\begin{tabular}{cc}
+\begin{minipage}{0.4\textwidth}
 \begin{center}
-\includegraphics[scale=0.25, trim = 0.7cm 0.7cm 0.7cm 0.7cm]{flavinet.eps}
+\includegraphics[scale=0.25, trim = 2cm 1.5cm 1cm 1.5cm]{Kshirsagar.Fig_3a.eps}
 \end{center}
 \end{minipage}
-\\
-\begin{minipage}{\textwidth}
+&
+\begin{minipage}{0.55\textwidth}
 \begin{center}
-\includegraphics[scale=0.25, trim = 0.5cm 0.5cm 0.5cm 0.5cm]{flunet.eps}
+\includegraphics[scale=0.35, trim = 2cm 1.5cm 1cm 1.5cm]{Kshirsagar.Fig_3b.eps}
 \end{center}
@@
 For enrichment analysis, we use the FuncAssociate 3.0 \footnote{\url{http://llama.mshri.on.ca/funcassociate/}} \citep{funcAsso} tool and GO annotations for the three viruses from UniProtKB.
-
-\begin{table}[bht]
-\begin{scriptsize}
-\begin{center}
-%\def\arraystretch{1.1}
-\begin{tabular}{|c|c|}
-\multicolumn{2}{c}{Gene Ontology terms for clusters from Figure \ref{fig:pca2}} \\ \hline
-cluster 1 & mRNA (guanine-N7-)-methyltransferase activity; RNA-directed \\
-(ebola only)& RNA polymerase activity; ATP binding \\ \hline
-cluster 2 & host cell endoplasmic reticulum membrane; \\ 
-(hepatitis \& & host cell lipid particle; induction by virus \\
-influenza & of host autophagy; host cell mitochondrial membrane; \\
-proteins) & suppression by virus of host STAT1 activity \\ \hline
-cluster 3 (all) & fusion of virus membrane with host plasma membrane \\ \hline
-\end{tabular}
-\caption{GO terms (for process, function, component) enriched in the highlighted clusters from Figure \ref{fig:pca2}.
-FuncAssociate GO enrichment tool was used and terms with p-value < 0.001 are shown.
-The first column also indicates the composition of each cluster (i.e which task's proteins appear in the cluster).}
-\label{goterms}
-\end{center}
-\end{scriptsize}
-\end{table}
 
@@
 
-\section{Results on HIV data}
-We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
+%\section{Results on HIV data}
+%We incorporate an additional task: HIV (human immunodeficiency virus), which is quite dissimilar to our original setting involving three viruses (HIV is a retrovirus).
 The interactions data (obtained from PHISTO and NIAID databases) has $\approx$4000 human-HIV PPI across several strains, of which 1320 PPI are non-homologous.
@@
 Table \ref{hivresults} shows the results.
-
-\begin{table}[h]\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
-Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2.
-Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of a somewhat dissimilar task such as HIV.}
-\label{hivresults}
-\begin{small}
-\begin{center}
-%\def\arraystretch{1.2}
-\begin{tabular}{c|cccc}
-\toprule
-& \multicolumn{4}{c}{10 \% training } \\
-& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
-Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
-STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
-Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
-BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
-\end{tabular}
-\end{center}
-\end{small}
-\end{table}
 
@@
 \bibliographystyle{natbib}
+\begin{normalsize}
+\bibliography{references}
+\end{normalsize}
+
+\newpage
+
+\begin{table}
+\begin{footnotesize}
+\caption{Tasks and their sizes (a column corresponds to one bipartite graph between human proteins and the virus indicated in the column header).
+More importantly, each task represents several strains of one virus species.
+We pool together PPI from all strains as the number of interactions per strain is very small and not representative.
+PPI data from different strains often involves homologous proteins (virus proteins or/and human proteins) as these were curated from different publications/studies involving the virus.
+The second row gives a conservative estimate of the number of non-homologous PPI.
+These were obtained by computing BLAST sequence similarity (see Section \ref{sec:phisto} for details).
+All pathogens are single stranded RNA viruses.
+The last row shows that each of our graphs is extremely sparse.}
+\label{datasets}
+\begin{tabular}{c|r|r|r} 
+\hline
+Task $\rightarrow$ &{\textit{Influenza A} } &\textit{Hepatitis C} & {\textit{Ebola}}\\
+& (28 strains) & (5 strains, many isolates)  & (3 strains) \\
+number of HP PPIs$^a$ (positives) & 848 & 981 & 90 \\\hline
+number of non-homologous PPIs & 274 & 270 & 75 \\ \hline
+\# of unique virus proteins in dataset & 54 & 151 & 2 \\\hline
+\# of unique human proteins in dataset & 362 & 385 & 88 \\\hline
+total \# of virus proteins in & 542 & 163 & 38 \\
+UniProtKB across strains$^b$ & & & \\ \hline
+number of negatives & 84800 & 98100 & 9000 \\\hline
+density of observed graph$^c$ (as \%) & .15 & .60 & .20 \\\hline
+\multicolumn{4}{l}{\scriptsize{$^a$ HP PPI: host-pathogen protein protein interactions}}\\
+\multicolumn{4}{l}{\scriptsize{$^c$ Density = (no. of positives + no. of negatives)/ total number of possible edges. The denominator was computed considering all}} \\
+\multicolumn{4}{l}{\scriptsize{\hspace{0.1cm} proteins from all strains of the virus and $\approx$ 100,000 human proteins.}}\\
+\multicolumn{4}{l}{\scriptsize{$^b$ this is the number of `reviewed' proteins on UniProtKB from all strains in our data}} \\ \hline
+\end{tabular}
+\end{footnotesize}
+\end{table}
+
+\clearpage
+
+\begin{table}[!t]
+\caption{Area Under the Precision-Recall curve (AUC-PR) on the PHISTO dataset.
+The column header `X\% training' indicates the fraction of the labeled data used for training and tuning the model with the rest (100-X)\% used as test data.
+We report the average AUC-PR over 10 random train-test splits (stratified splits that maintain the class-skew of 1:100).
+The standard deviation is also shown.
+The performance of the best baseline and the overall best method (BSL-MTL) is highlighted in bold.
+The Homolog baseline performs comparably to our BSL-MTL method on the \textit{Ebola} task.
+This is explained by the virus-human PPI graph, from several other viruses, that the Homolog baseline has access to.}
+\label{resultsTable}
+\begin{tabular}{c|ccccccc}
+\toprule
+& \multicolumn{3}{c}{\textbf{10\% training, 90\% test}} && \multicolumn{3}{c}{\textbf{30\% training, 70\% test}} \\ \cline{2-4}\cline{6-8}
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} && \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & \textbf{0.230}$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 && \textbf{0.311}$\pm$.03 & 0.180$\pm$.01 & 0.198$\pm$.01 \\
+STL (Ridge Reg.)  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02 && 0.130$\pm$.03 & 0.802$\pm$.03 & 0.428$\pm$.03 \\ 
+MMTL & 0.113$\pm$.04 & 0.767$\pm$.03 & 0.321$\pm$.02 && 0.129$\pm$.02 & 0.802$\pm$.04 & 0.430$\pm$.03  \\ 
+%Trace-norm  & 0.199$\pm$.11 & 0.767$\pm$.03 & 0.318$\pm$.02 && 0.207$\pm$.02 & 0.808$\pm$.02 & 0.409$\pm$.03 \\ 
+Sparse+low-rank & 0.144$\pm$.07 & 0.767$\pm$.02 & 0.318$\pm$.02 && 0.153$\pm$.02 & \textbf{0.814}$\pm$.01 & 0.414$\pm$.03  \\ 
+%Dirty model \citep{jalali2010} & 0.074$\pm$.03 & 0.767$\pm$.04 & 0.324$\pm$.02 && 0.165$\pm$.02 & 0.813$\pm$.03 & 0.412$\pm$.03  \\ 
+MTPL & 0.217$\pm$.08 & 0.695$\pm$.02 & 0.345$\pm$.02 && 0.260$\pm$.05 & 0.713$\pm$.01 & \textbf{0.496}$\pm$.03 \\ 
+IMC & 0.087$\pm$.04 & \textbf{0.779}$\pm$.02 & \textbf{0.362}$\pm$.01 && 0.122$\pm$.02 & 0.801$\pm$.01 & 0.410$\pm$.03  \\ \midrule
+BSL-MTL & \textbf{0.233}$\pm$.10 & \textbf{0.807}$\pm$.02 & \textbf{0.486}$\pm$.02 && \textbf{0.361}$\pm$.03 & \textbf{0.842}$\pm$.01 & \textbf{0.560}$\pm$.02  \\ \bottomrule
+\end{tabular}
+\end{table}
+
+\clearpage
+
+\begin{table}[!t]
+\caption{AUC-PR on the \textbf{non-homologous PPI within each task}, in the 10\% setting.
+To get an estimate of the relative size of the non-homologous partition for each task, see Table \ref{datasets}.
+Only the most representative baselines are retained for this experiment.
+Our model performs better than the others on the non-homologous part of the test data, as it captures non-linear similarities among the interactions within a task and across tasks very well.}
+\label{nohomresultsTable}
+\begin{small}
+%\def\arraystretch{1.2}
+\begin{tabular}{c|ccc}
+\toprule
+\multicolumn{4}{c}{\textbf{10 \% training, 90\% test on Non-homologous PPI}} \\
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & 0.207 $\pm$.07 & 0.113 $\pm$.01 & 0.109 $\pm$ .01 \\
+STL  & 0.210 $\pm$.14 & 0.474 $\pm$ .06 & 0.302 $\pm$.03 \\
+Sparse+LR & 0.216 $\pm$.13 & 0.482 $\pm$.05 & 0.312 $\pm$.04 \\ 
+BSL-MTL & \textbf{0.295} $\pm$.10 & \textbf{0.544} $\pm$.06 & \textbf{0.505} $\pm$.05 \\ \bottomrule
+\end{tabular}
+\end{small}
+\end{table}
+
+\clearpage
+
+\begin{table}[!t]
+\caption{AUC-PR on the \textbf{homologous PPI} within each task, in the 10\% setting.
+The Homolog baseline outperforms all on the \textit{Ebola} task and BSL-MTL beats all other methods on the other tasks.
+The Homolog method has access to a wider virus-human PPI graph involving several other viruses and is able to infer homologous PPI for \textit{Ebola} better (note that some of the homologous PPI are not redundant examples but involve proteins with very small regions of sequence similarity.}
+\label{homresultsTable}
+\begin{small}
+%\def\arraystretch{1.2}
+\begin{tabular}{c|ccc}
+\toprule
+\multicolumn{4}{c}{\textbf{10 \% training, test only Homologous PPI }} \\
+& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} \\ \midrule
+Homolog & \textbf{0.399} $\pm$.07 & 0.161 $\pm$.01 & 0.180 $\pm$ .01 \\
+STL   & 0.192 $\pm$.15 & 0.504 $\pm$.10 & 0.305 $\pm$.03 \\
+Sparse+LR & 0.184 $\pm$.13 & 0.512 $\pm$.10 & 0.315 $\pm$.04 \\ 
+BSL-MTL & 0.130 $\pm$.10 & \textbf{0.815} $\pm$.02 & \textbf{0.475} $\pm$.02 \\ \bottomrule
+\end{tabular}
+\end{small}
+\end{table}
+
+\clearpage
+
+\begin{table}[!t]
 \begin{scriptsize}
-\bibliography{references}
+\begin{center}
+%\def\arraystretch{1.1}
+\caption{GO terms (for process, function, component) enriched in the highlighted clusters from Figure \ref{fig:pca2}.
+FuncAssociate GO enrichment tool was used and terms with p-value $<$ 0.001 are shown.
+The first column also indicates the composition of each cluster (i.e which task's proteins appear in the cluster).}
+\label{goterms}
+\begin{tabular}{|c|c|}
+\multicolumn{2}{c}{Gene Ontology terms for clusters from Figure \ref{fig:pca2}} \\ \hline
+cluster 1 (ebola only) & mRNA (guanine-N7-)-methyltransferase activity; RNA-directed RNA polymerase activity; ATP binding \\ \hline
+cluster 2 (hepatitis \& & host cell endoplasmic reticulum membrane; host cell lipid particle; induction by virus \\ 
+influenza proteins) &  of host autophagy; host cell mitochondrial membrane; suppression by virus of host STAT1 activity \\ \hline
+cluster 3 (all) & fusion of virus membrane with host plasma membrane \\ \hline
+\end{tabular}
+\end{center}
 \end{scriptsize}
+\end{table}
+
+%\clearpage
+
+%
+%\begin{table}[!t]
+%\caption{AUC-PR upon addition of a new task - HIV, in the 10\% setting.
+Note that STL and the homolog method build independent models so their AUC numbers are the same as in Table 2.
+Among the multitask learning methods, BSL-MTL's performance is not hurt by the addition of a somewhat dissimilar task such as HIV.}
+%\label{hivresults}
+%\begin{small}
+%%\def\arraystretch{1.2}
+%\begin{tabular}{c|cccc}
+%\toprule
+%& \multicolumn{4}{c}{10 \% training } \\
+%& \textit{Ebola} & \textit{Hep-C} & \textit{Influenza} & \textit{HIV} \\ \midrule
+%Homolog & 0.230$\pm$.06 & 0.178$\pm$.01 & 0.158$\pm$.01 & 0.170 $\pm$.01 \\
+%STL  & 0.189$\pm$.09 & 0.702$\pm$.08 & 0.286$\pm$.02  & 0.208 $\pm$.01 \\ 
+%Sparse+LR & 0.135 $\pm$.06 & 0.724 $\pm$.05 & 0.280 $\pm$.01 & 0.205 $\pm$.03  \\ 
+%BSL-MTL & \textbf{0.241} $\pm$.01 & \textbf{0.793} $\pm$.06 & \textbf{0.459} $\pm$.01 & \textbf{0.354} $\pm$.01  \\ \bottomrule
+%\end{tabular}
+%\end{small}
+%\end{table}
 

### 2016-11-22 make.sh

@@ -1,0 +1,4 @@
+latex paper.tex
+bibtex paper
+latex paper.tex
+dvipdf paper

### 2016-11-22 paper.tex
new stuff

@@
 
-\author{Meghana Kshirsagar$^{1}$, Keerthiram Murugesan$^{2}$, Jaime G. Carbonell$^{2}$ and Judith Klein-Seetharaman$^{3}$}
-%\address{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021\\
-%$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA\\
-%$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK
-%}
+\author{Meghana Kshirsagar$^{1*}$, Keerthiram Murugesan$^{2}$,\\  Jaime G. Carbonell$^{2}$, Judith Klein-Seetharaman$^{3}$
+\thanks{$^{1}$Memorial Sloan Kettering Cancer Center, 1275 York Ave., NY 10021}
+\thanks{$^{2}$Language Technologies Institute, Carnegie Mellon Univ., 5000 Forbes Ave., Pittsburgh PA 15213, USA}
+\thanks{$^{3}$Metabolic \& Vascular Health, Warwick Medical School, Univ. of Warwick, Coventry, UK}
+}
 
@@
 \textbf{Contact:}
-\href{mkshirsa@cs.cmu.edu}{meghana.ksagar@gmail.com}
+\href{mkshirsa@cs.cmu.edu}{kshirsam@mskcc.org}
 \end{abstract}
@@
 \begin{figure}
-\includegraphics[scale=0.25, trim = 1.1cm 0.9cm 0 0]{pca_bslmtl_with_clusters.eps}
+\includegraphics[scale=0.25, trim = 1.1cm 0.9cm 0 0]{Kshirsagar.Fig_2.eps}
 \caption{Principal component analysis (PCA) of virus proteins in the projected subspace.
@@
 \bibliographystyle{natbib}
-\begin{normalsize}
 \bibliography{references}
-\end{normalsize}
 
